---
title: "2nd_input_data_processing.Rmd"
author: "Johannes Piipponen"
date: "`r Sys.Date()`"
output: html_document
---


Abbreviations
gl = grazing lands




# MCD
More spesifically MCD12Q1.061 (version 6.1)

Includes IGBP classes: 


See: 
https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MCD12Q1#description
or


```{r}
# Function to get MCD raster files for a specific year
rastlist_mcd_year <- function(year) {
  list.files(path = here("Data", "MODIS", "MCD", "v061_from_GEE"), 
             pattern= paste0("MCDv061_extent_in_", year), 
             full.names = TRUE) }

rastlist_mcd_year("2001") #check 



# Function to create a raster based on a raster list and merge tiles to cover the globe
f_merge_year_mcd <- function(year) {
  lapply(rastlist_mcd_year(year), rast) %>% # convert list to raster
  sprc() %>% # Create a SpatRasterCollection from the raster list
  merge(.,   # Merge rasters into a single raster
        filename = paste0(here("Data", "MODIS", "MCD", "v061_processed"),
                          "/mcd_merged_",
                        year ,".tif"),
        names = paste0("mcd", year),
        datatype = "INT1U") 
}


# Run the merge function for each year (2001-2020)
tic()
sapply(timestep_2001_2020 , f_merge_year_mcd) # 2h total
toc() 



# Create a raster stack from the merged TIF files
mymergedfiles_mcd <- list.files(path = here("Data", "MODIS", "MCD", "v061_processed"),
                            pattern="mcd_merged_",
                            full.names=TRUE)
mymergedfiles_mcd

mcd_merged <- rast(mymergedfiles_mcd) # no any kind of mask here!



# Create a reclassification matrix to change values 6, 7, 8, 9, 10 and 16 to 1 and all other values to NA
rclmat_mcd_igbp6to10or16 <- 
  matrix(c(6, 10, 1,# Values from 6 to 10 will become 1
           16, 16, 1),# Values of 16 will become 1
           ncol = 3, byrow = TRUE)


# Classify the mcd raster using the reclassification matrix
mcd_igbp6to10or16_gets_val_1_otherwise_NA <-
  classify(mcd_merged,
           rclmat_mcd_igbp6to10or16, 
           right = NA, # close both sides with right=NA. Only meaningful with "from-to-becomes" classification with integers.
           others = NA, # all values that are not matched are set to this value 
           filename = here("Data", "MODIS", "MCD", "v061_processed",
                           "mcd_igbp_6to10or16_gets_val_1_otherwiseNA.tif"),
           overwrite = T) # 30min




# Replace 0 values with NA in the raster stack. Then, find cells where there is at least one NA during the time period



# Create a raster mask with NA values where any year has an NA value. In other words:
# calculate sum of mcd_igbp_6to10_gets_val1_otherwiseNA using na.rm = F. This returns a raster where cell get NA value if any of the years has NA value in that cell. This mask is used to find cells that belong to some IGBP class 6-10 every year. E.g. cell 5,6,7,NA,6,6...,7 becomes NA

 # this retrieves 1 layer we use for masking
# mcd_gl_every_year_mask_1_or_NA <- 
#   sum(mcd_igbp_6to10_gets_val_1_otherwise_NA, na.rm = FALSE,
#       filename = here("Data", "MODIS", "MCD", "v061_processed",
#                       "mcd_sum_2001_2020_na.rm_FALSE.tif"), 
#       overwrite = T) # 10min


mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask <- 
  sum(mcd_igbp6to10or16_gets_val_1_otherwise_NA, na.rm = FALSE,
      filename = here("Data", "MODIS", "MCD", "v061_output",
                      "mcd_sum_igbp6to10or16_2001_2020_na.rm_FALSE.tif"), 
      overwrite = T) # 10min


#################################### get the file
mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask <- 
   here("Data", "MODIS", "MCD", "v061_output",
        "mcd_sum_igbp6to10or16_2001_2020_na.rm_FALSE.tif") %>%  
  rast() # actualy gets value 20 or NA (sum of 20 years)

mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask

plot(mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask,
     main = "1lyr showing where there is igbp6to10or16 every year 2001-2020")
#################################### get the file


```


# Fraction of grazing land within each 5 arc-minutes cell ---------------------------- poista turhat jutut täältä

The raster we created above gets value 1 if there are some of the IGPB classes 6to10 or 16 every year 2001-2020. Otherwise it gets value NA.
No we convert NA to 0 and aggregate and resample the raster to 5arcmin. This reveals how large fraction of cell is covered by grazing lands

```{r}


# Aggregate and resample the binary grazing land raster to 5 arc-minutes resolution
# This results in a raster representing the fraction of grazing land within each 5 arc-minutes cell
# Create a reclassification matrix for values greater than 1 to become 1
rclmat1toInf1 <- matrix(c(1, Inf, 1), ncol = 3, byrow = TRUE)
# Classify the mcd_gl_every_year_mask raster using the reclassification matrix (ifel would work too)
# same as mcd0_1_sum <- mcd0_to_na_sum; mcd0_1_sum[mcd0_1_sum > 1] <- 1; mcd0_1_sum <- classify(mcd0_1_sum, cbind(NA, 0))



mcd_0or1_sum_igbp6to10or16 <- 
  classify(mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask, 
           rclmat1toInf1, 
           include.lowest = TRUE, 
           others = 0)

mcd_0or1_sum_igbp6to10or16 <- # just in case the previous one leaves NAs anyways
  classify(mcd_0or1_sum_igbp6to10or16, cbind(NA, 0))




r_fraction_gl_0toNA_igbp6to10or16  <- 
  aggregate(mcd_0or1_sum_igbp6to10or16, fact = 18, na.rm = T) %>%
  resample(., template_rast_5arcmin) %>% 
  classify(., cbind(0, NA), 
           filename = here("Data", "MODIS", "MCD", "v061_output",
                          "mcd_fraction_of_gl0toNA_in_cell_where_igbp6to10or16_2001_2020_5arcmin.tif"),
                          overwrite = T)


# Calculate land area in 2010
# r_of_areas <- cellSize(mcd0_to_na$mcd2010, unit = "km") 
# mcd_area_2010 <- global(r_of_areas, fun = "sum", na.rm = T) 
# mcd_area_2010  / (148940000 - 14200000)  # XX% of total land area (excluding antarctica)
# 
# # Calculate the total land area when land usage is always IGBP 6to10or16 
r_of_study_area <- cellSize(mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask, unit = "km") *
  mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask
mcd_area_always_gl <-  global(r_of_study_area, fun = "sum", na.rm = T) #  M km2
mcd_area_always_gl  / (148940000 - 14200000)  # % of total land area (excluding antarctica)



#################################### get the file
r_fraction_gl_0toNA_igbp6to10or16 <- 
  here("Data", "MODIS", "MCD", "v061_output",
       "mcd_fraction_of_gl0toNA_in_cell_where_igbp6to10or16_2001_2020_5arcmin.tif") %>% 
  rast()
#################################### get the file


```





# Net Primary Productivity (NPP) 
More spesifically MOD17A3HGF.061 (version 6.1)

See:
https://developers.google.com/earth-engine/datasets/catalog/MODIS_061_MOD17A3HGF?hl=en

```{r}
# Function that get rasterlist for different years
rastlist_npp_year <- function(year) {
  list.files(path = here("Data", "MODIS", "NPP", "v061_from_GEE"),
             pattern= paste0("NPP", year), 
             full.names = TRUE) }

rastlist_npp_year("2001") #check 


f_merge_year_npp <- function(year) {
  lapply(rastlist_npp_year(year), rast) %>% # convert list to raster
  sprc() %>% # make a SpatRasterCollection from a list
  merge(.,
        filename = paste0(here("Data", "MODIS", "NPP", "v061_processed"),
                          "/npp_merged_",
                        year ,".tif"),
        names = paste0("npp", year)) #merge
}


##f_merge_year_npp("2001") # test for 2001
tic()
sapply(timestep_2001_2020 , f_merge_year_npp) # this takes a lot of time
toc() #90min



mymergedfiles_npp <- list.files(path =
                                  here("Data", "MODIS", "NPP", "v061_processed"),
                            pattern="npp_merged_",
                            full.names=TRUE)
mymergedfiles_npp
npp_nomask <- rast(mymergedfiles_npp) 




# writeRaster(npp_nomask, 
#             filename = here("Data", "MODIS", "NPP", "v061_output",
#                             "npp_kgC_ha_2001_2020_nomask.tif"),
#             overwrite = T)





 # mask
# ! Important note: The NPP data downloaded from GEE had a scaling factor of 0.0001 and the unit was kg*C/m^2.
# Therefore, it should be multiplied by 0.0001 to get the unscaled NPP. However, as we want NPP in the unit kg*C/ha, we should again multiply by 10000 to convert from m^2 to ha. 
# In this case, the two factors cancel each other out, as 0.0001 * 10000 = 1. Thus, we don't need to do anything to the GEE-downloaded data, as NPP in kg*C/m^2 with a 0.0001 scale is the same as NPP in kg*C/ha.


npp <- 
  mask(npp_nomask, 
       mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask, 
       filename = here("Data", "MODIS", "NPP", "v061_output",
              "npp_kgC_ha_where_igbp6to10or16_every_year2001_2020_res000449.tif"),
       overwrite = T)


#################################### get the file
 # allocated where grazing land every year 2001-2020 (some of the IGBP classes 6to10or16 every year)
npp <- here("Data", "MODIS", "NPP", "v061_output",
              "npp_kgC_ha_where_igbp6to10or16_every_year2001_2020_res000449.tif") %>% 
  rast()

plot(npp$npp2019)
#################################### get the file
```



# Tree Cover Multiplier (TCM)
More spesifically, the input data for TCM is MOD44B.006

See:
https://developers.google.com/earth-engine/datasets/catalog/MODIS_006_MOD44B


```{r}
# function that get rasterlist for different years
rastlist_treecover_year <- function(year) {
  list.files(path = here("Data", "MODIS", "MOD44B", "v06_from_GEE"),
             pattern= paste0("MOD44B", year), 
             full.names = TRUE) }



# rastlist_treecover_year("2001")[1] %>% rast()




# function for merging
f_merge_year_tc <- function(year) {
  lapply(rastlist_treecover_year(year), rast) %>% # convert list to raster
  sprc() %>% # make a SpatRasterCollection from a list
  merge(., # merge tiles
        filename = paste0(here("Data", "MODIS", "MOD44B", "v06_processed"),
                          "/tc_merged_",
                        year ,".tif"),
        names = paste0("tc", year, c("bot","med", "up"), collapse = ",")) 
}


tic()
sapply(timestep_2001_2020 , f_merge_year_tc) # this takes a lot of time
toc() #90min for a layer



# combine merged tiffs 
mymergedfiles_treecover <- 
  list.files(path = here("Data", "MODIS", "MOD44B", "v06_processed"),
             pattern="tc_merged_",
             full.names=TRUE)

mymergedfiles_treecover
treecover_nomask <- rast(mymergedfiles_treecover) # 3 curves for every year 2001-2020 
treecover_nomask$tc2020med %>% plot()


# mask 
# tic()
# treecover <- 
#   mask(treecover_nomask, mcd_gl_every_year_mask,
#        filename = here("Data", "MODIS", "MOD44B", "v06_output",
#                        "treecover2001_2020_where_gl_every_year2001_2020_res000449.tif"),
#        overwrite = T)
# toc() # 2.7h


# or
treecover <- 
  mask(treecover_nomask, mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask,
       filename = here("Data", "MODIS", "MOD44B", "v06_output",
               "treecover2001_2020_where_igbp6to10or16_every_year2001_2020_res000449.tif"),
       overwrite = T)

# ################################### get the file (allocated where pastureland 2001-2020)
treecover <- 
  here("Data", "MODIS", "MOD44B", "v06_output",
       "treecover2001_2020_where_igbp6to10or16_every_year2001_2020_res000449.tif") %>%
  rast()
treecover$tc2020med %>% plot() # looks ok: near deserts etc no trees --> tc close to 1
# ################################### get the file (masked)
```


# MAP 

Mean Annual Precipitation (MAP) based on climate normals of 1991-2020 (30yrs)
v061_processed
Data from
https://climate.northwestknowledge.net/TERRACLIMATE/index_directDownloads.php 

!! possible to get via GEE -- update code later if time
https://developers.google.com/earth-engine/datasets/catalog/IDAHO_EPSCOR_TERRACLIMATE

```{r}
# get precipitation files based on a year
rastlist_terraclim_ppt <- function(year) {
  list.files(path = here("Data", "TerraClimate", "directdownloads"),
             pattern= paste0("TerraClimate_ppt_", year), 
             full.names = TRUE) }

rastlist_terraclim_ppt(2001)
rastlist_terraclim_ppt(2001)%>% rast() #12 layers for a year, monthly, ppt in mm 



# function to get annual sums
f_sum_ppt <- function(year) { 
  sum(rastlist_terraclim_ppt(year) %>% rast(),
  na.rm = T)} # cumulative precipitation

# test1 <- f_sum_ppt(2001);plot(test1); summary(test1) #  Median : 322.0, Mean   : 537.0 


# Get precipitation sum for each year
tic()
ppt_1991_2020 <- sapply(timestep_climnorm, f_sum_ppt) %>% 
  rast() # 30min. 30 layers
toc()


# derive MAP 
MAP_2.5arcmin <- mean(ppt_1991_2020, na.rm = T)
plot(MAP_2.5arcmin); summary(MAP_2.5arcmin) # Median : 339.6,  Mean   : 546.4, Max.   :7105.0


 # save
writeRaster(MAP_2.5arcmin, 
            here("Data", "TerraClimate", "processed", "MAP_2.5arcmin.tif"),
            names = "MAP", overwrite = T)




MAP_5arcmin_nomask <- 
  aggregate(MAP_2.5arcmin, fact = 2, na.rm = T) %>% 
  resample(., template_rast_5arcmin) %>% 
  project(., template_rast_5arcmin, # not necessary to resample and project!
          filename = 
             here("Data", "TerraClimate", "output",
                  "MAP_nomask_5arcmin.tif"),
           overwrite = T)



MAP_5arcmin_nomask <-
  here("Data", "TerraClimate", "output",
       "MAP_nomask_5arcmin.tif") %>% 
  rast()


# MAP_5arcmin <- 
#   mask(MAP_5arcmin_nomask, r_fraction_gl_0toNA, filename = 
#          here("Data", "TerraClimate", "output",
#               "MAP_where_gl_every_year2001_2020_5arcmin.tif"),
#        overwrite =T)

# or
MAP_5arcmin <- 
  mask(MAP_5arcmin_nomask, 
       r_fraction_gl_0toNA_igbp6to10or16, 
       filename = here("Data", "TerraClimate", "output",
              "MAP_where_igbp6to10or16gl_every_year2001_2020_5arcmin.tif"),
       overwrite =T)



#disaggregate 
# tic()
# MAP_disagg <-  
#   disagg(here("Data", "TerraClimate", "processed", "MAP_2.5arcmin.tif") %>% rast(),
#                fact = 9,
#                filename = here("Data", "TerraClimate", "processed", 
#                                "MAP_disagg_res000463.tif"),
#          names = "MAP", overwrite = T) 
# toc() # 2min
# 
# 
# 
# # resample to MODIS NPP (500m or 0.004491576 deg) using mcd rast as template
# tic()
# MAP_res_nomask_000449 <- 
#   resample(MAP_disagg, mcd_gl_every_year_mask,
#            filename = here("Data", "TerraClimate", "output",
#                            "nomask_MAP_1991_2020_res000449.tif"),
#            overwrite = T)
# toc()


# MAP_res_000449 <- 
#   mask(MAP_res_nomask_000449, mcd_gl_every_year_mask,
#        filename = here("Data","TerraClimate", "output",
#                        "MAP_1991_2020_where_gl_every_year2001_2020_res000449.tif"),
#        overwrite = T)


# get the file
MAP_5arcmin <- 
  here("Data", "TerraClimate", "output",
       "MAP_where_igbp6to10or16gl_every_year2001_2020_5arcmin.tif") %>% 
  rast()
plot(MAP_5arcmin)
```

# MAT

Mean Annual Temperature (MAT) based on climate normals of 1991-2020 (30yrs).
Unit for MAT: Celsius´degrees

Data from
https://climate.northwestknowledge.net/TERRACLIMATE/index_directDownloads.php 

```{r}
# function for retrieving the temperature files for each year (tmin and tmax)
rastlist_terraclim_temperature <- function(min_or_max, year) {
  list.files(path = here("Data", "TerraClimate", "directdownloads"),
             pattern= paste0(min_or_max, "_", year), 
             full.names = TRUE) }

rastlist_terraclim_temperature("tmin", "2001") # check


# create a function that creates a raster based on a list, takes mean for each month (mean of min and max) and then derives annual mean temperature for every year
f_mean_temp <- function(year) { mean( # first mean function has arguments tmin, tmax
  rastlist_terraclim_temperature("tmin", year) %>% rast(), 
  rastlist_terraclim_temperature("tmax", year) %>% rast(),
  na.rm = T) %>% 
  mean(., na.rm = T)  } # annual average temperature

# test1 <- f_mean_temp("2001");plot(test1) # test



# Get average temperature for each year
avgtemp <- sapply(timestep_climnorm, f_mean_temp) %>% 
  rast() # 30min. Names as paste0("avgtemp_", timestep_climnorm)



# derive MAT from average annual temperatures
MAT_2.5arcmin <- mean(avgtemp, na.rm = T)
plot(MAT_2.5arcmin)
 # save
writeRaster(MAT_2.5arcmin, here("Data", "TerraClimate", "processed",
                                "MAT_2.5arcmin.tif"),
            names = "MAT", overwrite = T)


MAT_2.5arcmin <- 
  here("Data", "TerraClimate", "processed",
                                "MAT_2.5arcmin.tif") %>% 
  rast()

MAT_5arcmin_nomask <- 
  aggregate(MAT_2.5arcmin, fact = 2, na.rm = T) %>% 
  resample(., template_rast_5arcmin) %>% 
  project(., template_rast_5arcmin, filename = 
             here("Data", "TerraClimate", "output",
                  "MAT_nomask_5arcmin.tif"),
           overwrite = T)


MAT_5arcmin_nomask <- 
  here("Data", "TerraClimate", "output","MAT_nomask_5arcmin.tif") %>% 
  rast()

# MAT_5arcmin <- 
#   mask(MAT_5arcmin_nomask, r_fraction_gl_0toNA, filename = 
#          here("Data", "TerraClimate", "output",
#               "MAT_where_gl_every_year2001_2020_5arcmin.tif"),
#        overwrite =T)
# or
MAT_5arcmin <- 
  mask(MAT_5arcmin_nomask, 
       r_fraction_gl_0toNA_igbp6to10or16, filename = 
         here("Data", "TerraClimate", "output",
              "MAT_where_igbp6to10or16gl_every_year2001_2020_5arcmin.tif"),
       overwrite =T)

# disaggregate
# MAT_disagg <-  
#   disagg(here("Data", "TerraClimate", "processed", "MAT_2.5arcmin.tif") %>% rast(),
#                fact = 9,
#                filename = here("Data","TerraClimate", "processed",
#                                "MAT_disagg_res000463.tif"),
#          overwrite = T) 
# 
# MAT_disagg
# 
# 
# # resample to MODIS NPP (500m or 0.004491576 deg) using mcd rast as template
# MAT_res_000449_nomask <- 
#   resample(MAT_disagg, mcd_gl_every_year_mask,
#            filename = here("Data","TerraClimate", "output",
#                            "nomask_MAT_1991_2020_res000449.tif"),
#            overwrite =T)


# MAT_res_000449 <- mask(MAT_res_000449_nomask, mcd_gl_every_year_mask,
#             filename = here("Data","TerraClimate", "output",
#                             "MAT_1991_2020_where_gl_every_year2001_2020_res000449.tif"),
#             overwrite =T)

# ################################### get the file (masked)
MAT_5arcmin <- 
  here("Data", "TerraClimate", "output",
       "MAT_where_igbp6to10or16gl_every_year2001_2020_5arcmin.tif") %>% 
  rast()
MAT_5arcmin %>% plot()
################################### get the file
```


# Terrain slope

Data from: https://www.nature.com/articles/s41597-020-0479-6
Expressed originally in degrees*100 but we must convert them to slope%
Ref: 
Amatulli, G., McInerney, D., Sethi, T., Strobl, P., & Domisch, S. (2020). Geomorpho90m, empirical evaluation and accuracy assessment of global high-resolution geomorphometric layers. Scientific Data, 7(1), 162. https://doi.org/10.1038/s41597-020-0479-6


```{r}
slope_input <- 
  here("Data", "Slope", "rawdata",
       "dtm_slope_merit.dem_m_250m_s0..0cm_2018_v1.0.tif") %>%
  rast()


# aggregate and change to percents. Slopes originally degree * 100 but we need slope% instead 
slope_agg_percent <- aggregate(slope_input, fact = 2, na.rm = TRUE) %>% 
  app(., function (x) {100 * tan((x/100)*pi/180  )})



# Reclassify slope raster, which is now expressed in percents
m_slopes <- c(0, 10, 1,
              10, 30, 0.7,
              30, 60, 0.4,
              60, Inf, 0) # George and Lyle 2009
rclmat_slopes <- matrix(m_slopes, ncol=3, byrow=TRUE)

slope_agg_percent_reclassified <-
  classify(slope_agg_percent,
           rclmat_slopes, include.lowest = TRUE, othersNA = TRUE) 

# resample 
slope_agg_percent_reclassified_resamp <-   resample(slope_agg_percent_reclassified, mcd) 

 # remove few odd values (reclassified slope > 1) -- probably faster with classify
slope_agg_percent_reclassified_resamp[slope_agg_percent_reclassified_resamp > 1] <- 1



writeRaster(slope_agg_percent_reclassified_resamp,
           filename = here("Data", "Slope", "output",
                           "nomask_slope_reclassified_res000449.tif"),
           names = "slope", overwrite = T)

slope_agg_percent_reclassified_resamp <-
  here("Data", "Slope", "output",
                           "nomask_slope_reclassified_res000449.tif") %>% 
  rast()
  
# tic()  
# slope <- 
#   mask(slope_agg_percent_reclassified_resamp, mcd_gl_every_year_mask,
#        filename = here("Data", "Slope", "output",
#                        "slope_reclassified_where_gl_every_year2001_2020_res000449.tif"),
#               overwrite =T)
# toc()

# or 
slope <- 
  mask(slope_agg_percent_reclassified_resamp, mcd_igbp6to10or16_every_year_1_or_NA_1lyr_mask,
       filename = here("Data", "Slope", "output",
                       "slope_reclassified_where_igbp6to10or16_every_year2001_2020_res000449.tif"),
       overwrite =T)

################################### get the file (masked)
slope <-
  here("Data", "Slope", "output",
       "slope_reclassified_where_igbp6to10or16_every_year2001_2020_res000449.tif") %>% 
  rast()
plot(slope)
################################### get the file
plot(slope_agg_percent_reclassified_resamp)
```


# Next we need uncertainty data for NPP partitioning

Sun et al. (2021) Fig 6d CV
https://agupubs.onlinelibrary.wiley.com/doi/10.1029/2021JG006472

```{r}
sun_cv <- here("Data","Sun_et_al_CV","fengyuhao1030-Grassland_NPP-f7a9c9e",
               "Figs","Fig_6","Fig_6d",
               "fBNPP_CV_Proj_Clip_Scale_0001.tif") %>% 
  rast()
sun_cv


#sun_cv_proj <- project(sun_cv, mcd_gl_every_year_mask)

# Aggregate, change projection and resample (15sec)
ncell(sun_cv)/1e6
sun_cv_agg <- aggregate(sun_cv, fact = 9)
sun_cv_agg
#sun_cv_agg %>% plot()

## unscale here
sun_cv_agg_unscale <- sun_cv_agg * 0.0001


sun_et_al_cv_5arcmin <- 
  project(sun_cv_agg_unscale, "EPSG:4326") %>% 
  resample(., template_rast_5arcmin)
names(sun_et_al_cv_5arcmin) <- "sun_et_al_cv"


# writeRaster(sun_et_al_cv_5arcmin,
#             file = here("Data","Sun_et_al_CV", "Sun_et_al_intermediate_inputs",
#                         "CV_Sun_et_al_unscaled_5arcmin.tif"),
#             overwrite = T)


# --------------------------------------------------------------- Regional Means
# The CV values are not given for all pixels we have NPP data for. The approach below is to extract the regional (regions used in analysis) mean CV values, then add the mean value for the region to the empty pixels in that region. The approach used map algebra.
sun_et_al_regional_cv <- 
  extract(sun_et_al_cv_5arcmin, reg_wgs_vect, fun = "mean", na.rm = T)
sun_et_al_regional_cv <- sun_et_al_regional_cv %>% 
  mutate(subregion = reg_wgs_vect$Region_new,
         regnumber = 1:nrow(.))


reg_wgs_rast <-
  reg_wgs %>% 
  mutate(regnumber = 1:nrow(.)) %>% 
  as(., "Spatial") %>% 
  vect() %>% 
  rasterize(., template_rast_5arcmin,  field = "regnumber")

plot(reg_wgs_rast)

# classify. Change values of reg_wgs_rast so that these represent average CV values
  ## create reclassification matrix
rcl_sun_et_al <- sun_et_al_regional_cv %>% 
  dplyr::select(regnumber, sun_et_al_cv ) %>% 
  as.matrix()

  ## classify
sun_et_al_raster_fills_the_globe <-
  reg_wgs_rast %>% 
  classify(rcl_sun_et_al)
names(sun_et_al_raster_fills_the_globe) <- "sun_et_al_cv"


sun_et_al_raster_fills_the_globe  %>% qtm(., title = "Filled Sun et al CV")
plot(sun_et_al_cv_5arcmin, main = "original Sun et al CV before filling")
sun_et_al_cv_filled_nomask_5arcmin <- 
  terra::cover(sun_et_al_cv_5arcmin, sun_et_al_raster_fills_the_globe, values = NA)
plot(sun_et_al_cv_filled_nomask_5arcmin, main = "CV after filling") # for comparison








# ------------------------------------------------------- mask to relevant areas
# sun_et_al_cv_filled_5arcmin <-
#   mask(sun_et_al_cv_filled_nomask_5arcmin, r_fraction_gl_0toNA )
# names(sun_et_al_cv_where_gl_every_year_2001_2020_5arcmin) <- "sun_et_al_cv"

# or
sun_et_al_cv_filled_5arcmin <-
  mask(sun_et_al_cv_filled_nomask_5arcmin, r_fraction_gl_0toNA_igbp6to10or16 )
names(sun_et_al_cv_filled_5arcmin) <- "sun_et_al_cv"

plot(sun_et_al_cv_filled_5arcmin)

```




# Combine, aggregate and resample input data to 5arcmin 
In this chunck, masked to areas where there is some gl every year 2001-2020

```{r}
# ---------------------------------- mask to areas where gl every year 2001-2010 
# npp_tc_slope_where_gl_every_year_2001_2020_500m <-
#   c(npp, treecover, slope) # no MAT, MAP or sun_et_al_cc here
#   
# 
# # --------------------------------------- Aggregate and resample this input data
#   # masked input for AGB
# tic()
# npp_tc_slope_where_gl_every_year_2001_2020_agg  <-
#   npp_tc_slope_where_gl_every_year_2001_2020_500m %>% 
#   aggregate(., fact = 18, na.rm = T, filename =
#               here("Data", "AGB", "Inputs_for_AGBcalculations",
#                    "npp_tc_slope_where_gl_every_year_2001_2020_agg18.tif"),
#             overwrite =T) 
# toc() # 2-3h
# 
# 
# 
#   ## resample
# tic()
# npp_tc_slope_where_gl_every_year_2001_2020_5arcmin <-
#   npp_tc_slope_where_gl_every_year_2001_2020_agg %>% 
#   resample(., template_rast_5arcmin, filename =
#              here("Data", "AGB", "Inputs_for_AGBcalculations",
#                   "npp_tc_slope_where_gl_every_year_2001_2020_5arcmin.tif"),
#            overwrite = T)
# toc() # 1min
# 
# 
# 
# # ----------------------------- combine with other input data
# data_for_simulation_masked_where_gl_2001_2020 <-
#   c(npp_tc_slope_where_gl_every_year_2001_2020_5arcmin, 
#     MAT_5arcmin, MAP_5arcmin,
#     sun_et_al_cv_filled_5arcmin)
# 
# 
# writeRaster(data_for_simulation_masked_where_gl_2001_2020,
#             filename = here("Data", "AGB", "Inputs_for_AGBcalculations",
#                             "input_data_for_agb_sim_masked_where_gl_2001_2020.tif"),
#             overwrite =T)
# 
# 
# ################################################## get the gfile
# data_for_simulation_masked_where_gl_2001_2020 <-
#   here("Data", "AGB", "Inputs_for_AGBcalculations",
#        "input_data_for_agb_sim_masked_where_gl_2001_2020.tif") %>% 
#   rast()
# 
# plot(data_for_simulation_masked_where_gl_2001_2020$npp2010)
# plot(data_for_simulation_masked_where_gl_2001_2020$tc2010med)
# plot(data_for_simulation_masked_where_gl_2001_2020$sun_et_al_cv)
# plot(data_for_simulation_masked_where_gl_2001_2020$MAT)
# 
# 





# same but using different mask r_fraction_gl_0toNA_igbp6to10or16 
tic()
npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_agg  <-
  c(npp, treecover, slope) %>% # masked with igbp6to10or16gl
  aggregate(., fact = 18, na.rm = T, filename =
              here("Data", "AGB", "Inputs_for_AGBcalculations",
                   "npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_agg18.tif"),
            overwrite =T) 
toc() # 2-3h


  ## resample
tic()
npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_5arcmin <-
  npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_agg %>% 
  resample(., template_rast_5arcmin, filename =
             here("Data", "AGB", "Inputs_for_AGBcalculations",
                  "npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_5arcmin.tif"),
           overwrite = T)
toc() # 1min


# ----------------------------- combine with other input data
data_for_simulation_masked_where_igbp6to10or16gl_2001_2020 <-
  c(npp_tc_slope_where_igbp6to10or16gl_every_year_2001_2020_5arcmin, 
    MAT_5arcmin, 
    MAP_5arcmin,
    sun_et_al_cv_filled_5arcmin)


writeRaster(data_for_simulation_masked_where_igbp6to10or16gl_2001_2020,
            filename = here("Data", "AGB", "Inputs_for_AGBcalculations",
                            "input_data_for_agb_sim_masked_where_igbp6to10or16gl_2001_2020.tif"),
            overwrite =T)



####### get the file
data_for_simulation_masked_where_igbp6to10or16gl_2001_2020 <-
  here("Data", "AGB", "Inputs_for_AGBcalculations",
       "input_data_for_agb_sim_masked_where_igbp6to10or16gl_2001_2020.tif") %>% 
  rast()
```




# Aggregate and resample nomasked input data to 5arcmin 

```{r}
# Files without any mcd mask
 # npp
mymergedfiles_npp <- 
  list.files(path = here("Data", "MODIS", "NPP", "v061_processed"),
             pattern="npp_merged_",
             full.names=TRUE)
npp_nomask <- rast(mymergedfiles_npp) 
 # treecover
mymergedfiles_treecover <- 
  list.files(path = here("Data", "MODIS", "MOD44B", "v06_processed"),
             pattern="tc_merged_",
             full.names=TRUE)
treecover_nomask <- rast(mymergedfiles_treecover) 

 # slope
slope_nomask <- 
  here("Data", "Slope", "output",
       "nomask_slope_reclassified_res000449.tif") %>% 
  rast()


npp_tc_slope_nomask_500m <- 
  c(npp_nomask, treecover_nomask, slope_nomask) 


  # unmasked input for AGB
tic()
npp_tc_slope_nomask_agg   <-
  npp_tc_slope_nomask_500m %>% 
  aggregate(., fact = 18, na.rm = T, filename =
              here("Data", "AGB", "Inputs_for_AGBcalculations",
                   "npp_tc_slope_nomask_agg18.tif")) 
toc() # 2h


  ## resample
tic()
npp_tc_slope_nomask_5arcmin <-
  npp_tc_slope_nomask_agg  %>% 
  resample(., template_rast_5arcmin, filename =
             here("Data", "AGB", "Inputs_for_AGBcalculations",
                  "npp_tc_slope_nomask_5arcmin.tif"),
           overwrite = T)
toc() # 2-15min


# Combine with non masked MAT, MAP and Sun et al CV
data_for_simulation_nomask <- 
  c(npp_tc_slope_nomask_5arcmin,
    sun_et_al_cv_filled_nomask_5arcmin,
    MAT_5arcmin_nomask,
    MAP_5arcmin_nomask)


writeRaster(data_for_simulation_nomask,
            filename = here("Data", "AGB", "Inputs_for_AGBcalculations",
                            "input_data_for_agb_sim_nomask.tif"))

############################################ get the file
data_for_simulation_nomask <-
  here("Data", "AGB", "Inputs_for_AGBcalculations",
       "input_data_for_agb_sim_nomask.tif") %>% 
  rast()
```


# 500m (not sure if needed)

```{r}

 # MAT
MAT_res_000449_nomask <- 
  here("Data","TerraClimate", "output","nomask_MAT_1991_2020_res000449.tif") %>% 
  rast()
 # MAP
MAP_res_nomask_000449 <- 
  here("Data", "TerraClimate", "output","nomask_MAP_1991_2020_res000449.tif") %>% 
  rast()
```



# Combine data_for_simulation to include GLW4 da animals

```{r}
######## get the data
data_for_simulation_masked_where_igbp6to10or16gl_2001_2020 <-
  here("Data", "AGB", "Inputs_for_AGBcalculations",
       "input_data_for_agb_sim_masked_where_igbp6to10or16gl_2001_2020.tif") %>% 
  rast()

glw4_da_and_area <- here("Data", "GLW", "GLW4", "output", "glw4_da_and_area.tif") |> rast()
plot(glw4_da_and_area) # NA values e.g. buffalo north america

# convert NA to 0
glw4_da_and_area <- 
  classify(glw4_da_and_area, cbind(NA,0))


glw4_da_and_area_masked <-
  mask(glw4_da_and_area, data_for_simulation_masked_where_igbp6to10or16gl_2001_2020$npp2001)


data_for_simulation_also_GLW4_animals <- 
  c(data_for_simulation_masked_where_igbp6to10or16gl_2001_2020, 
    glw4_da_and_area_masked)

names(data_for_simulation_also_GLW4_animals) # good!

writeRaster(data_for_simulation_also_GLW4_animals,
            filename = here("Data", "AGB", "Inputs_for_AGBcalculations",
                            "input_data_for_agb_sim_also_GLW4_da_animals.tif"),
            overwrite =T)


########### get the data_for_simulation_also_GLW4_animals
data_for_simulation_also_GLW4_animals <-
  here("Data", "AGB", "Inputs_for_AGBcalculations",
       "input_data_for_agb_sim_also_GLW4_da_animals.tif") %>% 
  rast()

```


# Save some outputs separately with 5arcmin resolution

needed elsewhere

```{r}
tic()
# NPP
npp_nomask_5arcmin <- 
  aggregate(npp_nomask, fact = 18, na.rm = T) |> 
  resample(template_rast_5arcmin,
           filename = here("Data", "MODIS", "NPP", "v061_output",
                           "npp_nomask_5arcmin.tif"),
           overwrite = T)

# Treecover
tcm_nomask_5arcmin <- 
  aggregate(treecover_nomask, fact = 18, na.rm = T) |> 
  resample(template_rast_5arcmin,
           filename = here("Data", "MODIS", "MOD44B", "v06_output",
                           "tcm_nomask_5arcmin.tif"),
           overwrite = T)
toc()
```

