
Protein production capacity from aboveground biomass (AGB)

? nyt vuosi 2020 --> pitäisikö olla mediaani vuosista 2018-2022

# !!!!!!!!!!!!!!!!! vaihdettu reviisiossa
- valittu vuosi 2020 ja kaikki Y2010 vaihdettu Y2020


# Calculate protein to energy conversion using FAOSTAT statistics

```{r}
# ---------------------------------------------------------- FAOSTAT download 
foodBalanceSheets_semiraw <-
  here("Data", "Input", "faostat", "foodBalanceSheets_E_All_Data_NOFLAG.csv") %>% 
  read.csv() %>% 
  # dplyr::select(-Element.Code,-Unit ,-c(Y2011:Y2019))
  dplyr::select(-Element.Code,-Unit , -c(Y2010:Y2019), -c(Y2021)) #


my_elementlist_fao <- c(
  "Production",
  "Food supply quantity (kg/capita/yr)",
  "Food supply (kcal/capita/day)",
  "Protein supply quantity (g/capita/day)")

my_itemlist_fao_animals <- c("Meat", "Bovine Meat", "Mutton & Goat Meat")





# ---------------------------------------- FAOSTAT animal protein to kcal
fbs_animals <- foodBalanceSheets_semiraw %>% 
  filter(Item %in% my_itemlist_fao_animals,
         Element %in% my_elementlist_fao) 


 ## pivot wider
fbs_animals_wide <- fbs_animals %>% 
  pivot_wider(names_from = Element, values_from = Y2020) %>% 
  rename(prod_1000mt = "Production", # eg numbers in FIN equals Lukestat
         supply.KgCapYr = "Food supply quantity (kg/capita/yr)",
         supply.KcalCapD = "Food supply (kcal/capita/day)",
         prot.gCapD = "Protein supply quantity (g/capita/day)") %>% 
  
  
  ## how many kcal do we get from amount of meat containing one kg of protein - required for simple conversion from protein production to energy production
  ## create variable protein to kcal 
  mutate(one_kg_prot_to_kcal =  (supply.KcalCapD / prot.gCapD) * 1000,
   #      one_kg_prot_to_kg_meat_product = ((supply.KgCapYr/365*1000) / prot.gCapD),
         protein_fraction = (prot.gCapD*365/1000) / supply.KgCapYr) %>% # protein fraction for meat products -- not even needed?
  # one_kg_prot_to_kg_meat_product is the same as 1/protein_fraction
  dplyr::select(-Item.Code) %>% 
  
  
  mutate_all(function(x) ifelse(is.nan(x) | is.infinite(x), NA, x))  # convert NaN values to NA






# fill empty columns (values 0 or NA) with global averages similardly as done earlier with crops
  ## ex with Bovine meat. First, find global average for one_kg_prot_to_kcal
# one_kg_prot_to_kcal_bovine <- fbs_animals_wide %>% 
#   filter(Item == "Bovine Meat", Area == "World") %>%
#   pull(one_kg_prot_to_kcal)
# 
#   ## then replace 0 or NA values with this value
# test_bovine  <- fbs_animals_wide %>% 
#   filter(Item == "Bovine Meat") %>% 
#   mutate(one_kg_prot_to_kcal = 
#            ifelse(is.na(one_kg_prot_to_kcal) | one_kg_prot_to_kcal == 0 , # if one_kg_prot_to_kcal fraction = NA or 0,
#                                    one_kg_prot_to_kcal_bovine, # fill with the global average value
#                                    one_kg_prot_to_kcal))  # otherwise let it just be




# do this with function for all meat items and fill also protein_fraction col if that's needed later
f_fill_missing_nutritient_values_animals<- function(animalcategory) {

   ## pull one_kg_prot_to_kcal for spesific animal species
  one_kg_prot_to_kcal_animalcategory <- fbs_animals_wide %>% 
  filter(Item == animalcategory, Area == "World") %>%
  pull(one_kg_prot_to_kcal)
  
  ## pull protein fraction for spesific crop
  protein_fraction_animalcategory <- fbs_animals_wide %>% 
  filter(Item == animalcategory, Area == "World") %>%
  pull(protein_fraction)
  
  
  ## fill empty one_kg_prot_to_kcal amd protein_fraction  with global averages
  fbs_animals_wide_filled <- fbs_animals_wide %>% 
    filter(Item == animalcategory) %>%
    
    
    ## fill one_kg_prot_to_kcal column
    mutate(one_kg_prot_to_kcal = 
           ifelse(is.na(one_kg_prot_to_kcal) | one_kg_prot_to_kcal == 0 , # if protein fraction = NA or 0,
                                   one_kg_prot_to_kcal_animalcategory, # fill with the global average value
                                   one_kg_prot_to_kcal)) %>% # otherwise let it just be    
    
    ## fill protein_fraction column
    mutate(protein_fraction = 
           ifelse(is.na(protein_fraction) | protein_fraction == 0 , 
                                   protein_fraction_animalcategory,
                                   protein_fraction)) 
  

  return(fbs_animals_wide_filled)
}





# Lapply this for all the animal species  --> returns a list
fbs_animals_wide_filled_list <- 
  lapply(my_itemlist_fao_animals, 
         f_fill_missing_nutritient_values_animals) 

# next we use bind_rows to make this reasonable looking dataframe
fbs_animals_wide_filled <-
  bind_rows(fbs_animals_wide_filled_list)  


# Add Sudan rows to South Sudan (give new FAO id etc to South Sudan but use other values from Sudan)
# sudan <- fbs_animals_wide_filled %>% 
#   filter(Area == "Sudan") %>% 
#   dplyr::select(-Area.Code, -Area)
# 
# south_sudan <- sudan %>% 
#   mutate(Area.Code = 277,
#          Area = "South Sudan")
# 
# # Combine to the fbs_animals_wide_filled
# fbs_animals_wide_filled <- 
#   rbind(fbs_animals_wide_filled, south_sudan)
```


# Aboveground biomass (AGB) om HYDE grazing lands 



! Note that r_fraction_gl contains areas that are not visible in r_agb dataset

```{r}
# ---------------------------------------------- find HYDE grazing land fraction
r_fraction_gl <- 
  here("Data", "Input", "from_harddrive", ## add so that this file is in same place
       "r_fraction_hyde_grazing_lands_masked_to_MODIS_IGBP_areas_0toNA.tif") %>%   #### changed  was HYDE area only
  rast() # if value > 1 it should be just 1
plot(r_fraction_gl, main = "fraction of grazing lands in this study (HYDE-MODIS integration)")


r_fraction_gl[r_fraction_gl >1] <- 1
r_fraction_gl_0toNA <- classify(r_fraction_gl, cbind(0,NA))
r_fraction_gl_NAto0 <- classify(r_fraction_gl, cbind(NA,0))


# Simulated AGB (median over 2008-2012) masked to HYDE areas (some areas, where IGBP classes 6-10or16 are not covered are not present in this AGB data. But they are visible in r_fraction_gl)
# !! update -- year 2020
r_agb_2020_in_hyde_gl_kg_ha <-
  here("Data", "Input", "from_harddrive",
       "AGB2020_withIGPB6to10or16_in_hydeareas_kg_ha_5arcmin.tif") %>% 
  rast()
# old
# r_agb_5y_med_in_hyde_gl_kg_ha <-
#   here("Data", "Input", "from_harddrive",
#        "5y_medAGB_withIGPB6to10or16_in_hydeareas_kg_ha_5arcmin.tif") %>% 
#   rast()


 ## AGB per pixel without any cropland mask. present result kg km2 but this is for global sums
r_agb_kg_perpix <-
  r_agb_2020_in_hyde_gl_kg_ha *
  cellSize(r_agb_2020_in_hyde_gl_kg_ha, unit = "ha") *
  # r_agb_5y_med_in_hyde_gl_kg_ha * 
  # cellSize(r_agb_5y_med_in_hyde_gl_kg_ha, unit = "ha") * 
  r_fraction_gl






#------------------------------------------------ FCR created in livestock_fcr.R
# Next we convert AGB to protein using the following FCR values Herrero et al. (2013) ------------------------------------------------ katso vaihe vaiheelta että on niitä arvoja joita dataframessa on
# fcr_complete <-
#   read.csv(here("Vilma_fcr_repo", "livestock_fcr_repo", "Outputs",
# # "fcr_grazing_by_animalspecies_bycountry_HerreroMekonnen.csv")) %>% 
#             "fcr_grazing_by_animalspecies_bycountry_HerreroOnly.csv")) %>% ####################################### !!!!!!! Fix with Kajsa
#   as_tibble()
# 
# 
#   ## combine with fao geometries and select production system = Grazing, and drop SG milk
# adm_10m_fao_id_simple_combined_with_fcr <-
#   fcr_complete %>%
#  # filter(prod_sys_2 == "Grazing", Animal.category != "Sg milk") %>%
#  # left_join(adm_10m_fao_id_simple, ., by = c("FAO_ID" = "Area.Code"))
#   left_join(adm10_simple_faoadded, ., 
#              by = c("SOVEREIGNT" = "AreaName")) 
#             #  by = c("FAO_ID" = "Area.Code")) # no South Sudan in fcr_complete



# #------------------ Create FCR raster for spesific Animal.Category with function
# f_FCR_raster <- function(animalcategory) {
#   
#   ## create reclassification matrix
#   rcl_mat_countries_to_fcr <-
#   adm_10m_fao_id_simple_combined_with_fcr %>% 
#   filter(Animal.category == animalcategory) %>% 
#   dplyr::select(FAO_ID, FCR_herrero) %>% # was FCR_uniform_median
#   st_drop_geometry() %>% 
#   as.matrix(., ncol = 2)
# 
# 
#   ## classify. Creates a raster that shows FCR for different countries
#   r_fcr_in_f  <- cntry_raster %>% # in_f refers that in function -- not global var
#     classify(rcl_mat_countries_to_fcr, 
#              others =  # ----------------------------------------------------------------------- tämä tekee väärän jutun, luulen
#                fcr_complete %>% filter(Animal.category == animalcategory) %>% 
#                pull(FCR_herrero) %>% mean()) # fill others with mean value # was FCR_uniform_median
#   
#   names(r_fcr_in_f) <- paste0(animalcategory, "_FCR")
# 
# return(r_fcr_in_f)
# }


#f_FCR_raster("Dairy cattle") # test


# create average FCR raster based on mean of simulated FCRs of beef, cattle and sheepgoats
# fcr_average <-
# mean(
#   f_FCR_raster("Dairy cattle"),
#   f_FCR_raster("Beef cattle"),
#   f_FCR_raster("Sg meat")
# ) 
# names(fcr_average) <- "fcr_average"

#  ## create fcr raster
# r_fcr <- c(
#   f_FCR_raster("Dairy cattle"),
#   f_FCR_raster("Beef cattle"),
#   f_FCR_raster("Sg meat")
#  # fcr_average
# )
# names(r_fcr) <- c("fcr_dairy", "fcr_beef", "fcr_sheep_and_goat_meat")
# 
# 
# plot(r_fcr)



```

# Feed Conversion Ratios (FCR)  

```{r}
# get data
fcr_country <- 
  read_excel(here("Data", "Input", "fromKajsa", "FCRs_bycountry.xlsx")) %>% 
  filter(Region != is.na(Region))



fcr_country <- fcr_country %>%
  mutate(FAO_from_countrynamme = countrycode(.$Country, origin = "country.name",  destination = "fao"))

# Similar
# kok <- fcr_country %>% 
#   dplyr::select(Country, FAO_from_countrynamme)
# 
# kok2 <- adm10_simple_faoadded %>% 
#    dplyr::select(ADMIN, FAO_ID) %>% 
#   st_drop_geometry()

# Regional medians
fcr_regional_medians <- 
  fcr_country %>% 
  group_by(Region) %>% 
  summarise(FCR_reg_median = median(FCR_totalFeedDivTotalProtoutput, na.rm = T))

fcr_regional_medians # almost sensible -- what to do with russian?


# Clean data and convert NA and 0 to regional median values
fcr_country_clean <- fcr_country %>% 
  dplyr::select(Country, FAO_from_countrynamme, Region, FCR_totalFeedDivTotalProtoutput)


# Liitä alueelliset mediaaniarvot alkuperäiseen tietokehykseen
fcr_country_clean <- fcr_country_clean %>% 
  left_join(fcr_regional_medians, by = "Region")


# Replace NA and zero values with the regional median value
fcr_country_clean <- fcr_country_clean %>% 
  mutate(
    FCR_kgDMperkgProtein = case_when(
      is.na(FCR_totalFeedDivTotalProtoutput) ~ FCR_reg_median,  # Replace NA with regional median
      FCR_totalFeedDivTotalProtoutput == 0 ~ FCR_reg_median,   # Replace zeros with regional median
      .default = FCR_totalFeedDivTotalProtoutput)  # Otherwise, keep the original value
  )

fcr_country_clean$FCR_kgDMperkgProtein %>% summary()

# Laske yhteenveto FCR arvoista alueittain
# fcr_country_clean %>%
#   group_by(Region) %>%
#   summarise(
#     Min = min(FCR_kgDMperkgProtein, na.rm = TRUE),
#     Max = max(FCR_kgDMperkgProtein, na.rm = TRUE),
#     Mean = mean(FCR_kgDMperkgProtein, na.rm = TRUE),
#     Median = median(FCR_kgDMperkgProtein, na.rm = TRUE),
#     SD = sd(FCR_kgDMperkgProtein, na.rm = TRUE)
#   )




######################################################## combine with geometries
adm_10m_fao_id_simple_combined_with_fcr <-
  fcr_country_clean %>%
  left_join(adm10_simple_faoadded, .,
             by = c("FAO_ID" = "FAO_from_countrynamme"))




# Fill missing countries. If a FCR value of a country is missing, fill it with region median FCR value

# Update the 'Region' column based on the conditions for 'REGION_UN'
adm_10m_fao_id_simple_combined_with_fcr <- adm_10m_fao_id_simple_combined_with_fcr %>%
  mutate(
    Region = case_when(
      is.na(Region) & REGION_UN == "Asia" ~ "EASA",  # East and Southeast Asia
      is.na(Region) & REGION_UN == "Americas" ~ "LAM",  # Latin America and the Caribbean
      is.na(Region) & REGION_UN == "Africa" ~ "SSA",  # Sub-Saharan Africa
      is.na(Region) & REGION_UN == "Europe" ~ "EUR",  # Western Europe
      is.na(Region) & REGION_UN == "Oceania" ~ "OCE",  # Oceania
      .default =  Region  # Otherwise, keep the original value
    )
  )



# Sitten liitetään fcr_regional_medians uudelleen
adm_10m_fao_id_simple_combined_with_fcr <- 
  adm_10m_fao_id_simple_combined_with_fcr %>%
  left_join(fcr_regional_medians, by = "Region", suffix = c(".x", ""))


# Lopuksi täytetään puuttuvat FCR_kgDMperkgProtein arvot
adm_10m_fao_id_simple_combined_with_fcr <- 
  adm_10m_fao_id_simple_combined_with_fcr %>%
  mutate(
    FCR_kgDMperkgProtein = case_when(
      is.na(FCR_kgDMperkgProtein) ~ FCR_reg_median,
      .default =  FCR_kgDMperkgProtein
    )
  )


################################################################# rasterize
r_fcr <- 
  rasterize(adm_10m_fao_id_simple_combined_with_fcr,
            template_rast_5arcmin,
            field = "FCR_kgDMperkgProtein")

global(r_fcr, "mean", na.rm = T) # 222
median(values(r_fcr), na.rm = T) # 187


plot(r_fcr, main = "FCR for each country. Unit: kgDM/kgProtein")

# writeRaster(r_fcr,
#             filename = here("Delete", "FCR_totalFeedDivTotalProtoutput.tif"))
```



# Current herd compositions --- no needed anymore

Unfortunately, this data cannot be shared. It can be requested from Mario Herrero and then from Kajsa Sahlin who processed the data further. Because the data is not openly available I builded the script so that elements that contains info about herd structures are kind of like addition parts and not included in main functions for example

```{r}
# df_herd <- 
#   read_excel(here("Data", "Input", "herrero",
#                   "modified_animal_composition2000.xlsx"))
# df_herd_long <- df_herd %>%
#   pivot_longer(cols = c(Dairy_cattle, Beef_cattle, Sheep_and_goats),
#                names_to = "Animal_type",
#                values_to = "share")
# 
# 
# df_herd_long <- df_herd_long %>%
#   mutate(fao_from_dfcountryname = countrycode(.$Country, origin = "country.name",  destination = "fao"))
# ## warns that some are missing. However, either fao_from_iso3eh or fao_from_SOVEREIGNT includes most
# 
# 
# adm_10m_fao_id_simple_combined_with_herd_structures <-
#   left_join(adm10_simple_faoadded, df_herd_long, 
#              by = c("FAO_ID" = "fao_from_dfcountryname")) 
# 
# 
# 
# 
# #------------------ Create HERD raster for specific Animal Category with function
# f_HERD_raster <- function(animalcategory) {
#   
#   ## create reclassification matrix
#   rcl_mat_countries_to_herd <-
#   adm_10m_fao_id_simple_combined_with_herd_structures %>% 
#   filter(Animal_type  == animalcategory) %>% 
#   dplyr::select(FAO_ID, share) %>%
#   st_drop_geometry() %>% 
#   as.matrix(., ncol = 2)
# 
#   ## classify. Creates a raster that shows HERD values for different countries
#   r_herd_in_f  <- cntry_raster %>%
#     classify(rcl_mat_countries_to_herd, 
#              others = NA)
#                # df_herd_long %>% filter(Animal_type == animalcategory) %>% http://127.0.0.1:14347/graphics/plot_zoom_png?width=2327&height=1252
#                # pull(share ) %>% mean()) # fill others with mean value
#   
#    # Apply the land mask
#   r_herd_in_f <- r_herd_in_f * land_mask
#   
#   names(r_herd_in_f) <- paste0(animalcategory, "_share")
# 
#   return(r_herd_in_f)
# }
# 
# f_HERD_raster("Dairy_cattle")
# 
# 
# # get rasters for each animal categories
# r_herd <- c(
#   f_HERD_raster("Dairy_cattle"),
#   f_HERD_raster("Beef_cattle"),
#   f_HERD_raster("Sheep_and_goats")
# )
# names(r_herd) <- c("dairy_share", "beef_share", "sheep_and_goats_share")
# sum(r_herd) # test that sums to 100%
# plot(r_herd)
```




# Convert AGB in HYDE grazing lands to protein



```{r}
# ------------------------------------------------ protein from AGB per ha
r_protein_from_AGB_kg_ha <-
  r_agb_2020_in_hyde_gl_kg_ha / r_fcr
  #r_agb_5y_med_in_hyde_gl_kg_ha / r_fcr 

names(r_protein_from_AGB_kg_ha) <- c("current_herd_str") #, "average"

plot(r_protein_from_AGB_kg_ha, 
     main = "protein from AGB when 43% of AGB is used ") 

# !! bit different extent
# kok <- r_protein_from_AGB_kg_ha$current_herd_str
# kok <- ifel(kok > 0, 1, NA)
# 
# kok2 <- r_fraction_gl_0toNA
# kok2 <- ifel(kok2 > 0, 1, NA)
# 
# plot(kok)
# plot(kok2)



# ------------------- Protein that current herd structure could provide (kg/ha)
# Calculate and classify the protein amount for each animal category

# # Dairy share: Calculate protein from dairy and classify the result
# protein_from_dairy <- (0.65 * r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$dairy_share) / r_fcr$fcr_dairy
# protein_from_dairy_classified <- classify(protein_from_dairy, cbind(NA, 0))
# 
# # Beef share: Calculate protein from beef and classify the result
# protein_from_beef <- (0.65 * r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$beef_share) / r_fcr$fcr_beef
# protein_from_beef_classified <- classify(protein_from_beef, cbind(NA, 0))
# 
# # Sheep and goats share: Calculate protein from sheep and goats and classify the result
# protein_from_sheep_and_goats <- (0.65 * r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$sheep_and_goats_share) / r_fcr$fcr_sheep_and_goat_meat
# protein_from_sheep_and_goats_classified <- classify(protein_from_sheep_and_goats, cbind(NA, 0))
# 
# # Combine all categories to get the final result
# r_protein_from_065xAGB_current_herd_str_kg_ha <- protein_from_dairy_classified +
#                                                 protein_from_beef_classified +
#                                                 protein_from_sheep_and_goats_classified

# # Classify the final result
# r_protein_from_065xAGB_current_herd_str_kg_ha <- classify(r_protein_from_065xAGB_current_herd_str_kg_ha, cbind(0, NA))
# names(r_protein_from_065xAGB_current_herd_str_kg_ha) <- "current_herd_str"

# ------------------- Combine results
# Combine the protein calculations for different categories and the current herd structure
#r_protein_from_AGB_kg_ha <- c(r_protein_from_AGB_kg_ha, r_protein_from_065xAGB_current_herd_str_kg_ha)

# Plot the final result
#plot(r_protein_from_AGB_kg_ha)

# ------------------------------------------------ protein from AGB per pixel
r_protein_from_AGB_kg_perpix <- 
  r_protein_from_AGB_kg_ha *
  cellSize(r_protein_from_AGB_kg_ha, unit = "ha") *
  r_fraction_gl



# ---------------------------- add also fcr of current herd structures here
  # this is actually not used anywhere and it is not correct to claculate amount of protein originating from AGB by using this
# r_fcr_current_herd_str <-
#   r_fcr$fcr_dairy * r_herd$dairy_share +
#   r_fcr$fcr_beef * r_herd$beef_share +
#   r_fcr$fcr_sheep_and_goat_meat * r_herd$sheep_and_goats_share
# names(r_fcr_current_herd_str) <- "fcr_current_herd"
# plot(r_fcr_current_herd_str)  #
# 
# #kokeile tuleeko samanlainen proteiinimäärä jos laskee tällä
# r_protein_from_AGB_kg_ha2 <-
#   (0.65*r_agb_5y_med_in_hyde_gl_kg_ha) / r_fcr_current_herd_str 
# 
# r_protein_from_AGB_kg_ha2 %>% summary()
# r_protein_from_AGB_kg_ha$current_herd_str %>% summary()
```






# Calculate global sums
Convert per ha values to perpix



```{r}
# Total sustainable AGB amount in petagrams
global(1000*r_agb_kg_perpix, "sum", na.rm = T)/1e15 # 1.7 pg --> 1.7 million tons compared to 6 of Mottet et al
global(1000*(r_agb_kg_perpix/0.42), "sum", na.rm = T)/1e15 # 4 without PUF - still very cautious / small

# Total protein produce
  # with current: 9.7
global(r_protein_from_AGB_kg_perpix, fun = "sum", na.rm = T)/1e9 # 9.7 








# livestock protein masked to cropland areas --- uncomment when need to calculte these
# !!!!!!!!!!!! move to somewhere else -- maybe a place where all global sums are calculated?


# r_protein_from_AGB_kg_perpix_masked_to_croplands <-
#   mask(r_protein_from_AGB_kg_perpix, r_physical_areas_crops_sum_ha_perpix)
# 
# global(r_protein_from_AGB_kg_perpix_masked_to_croplands, "sum", na.rm = T)/1e9
# 
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$dairy) /1e9 #  dairy
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$beef) /1e9 #  beef
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$sheep_and_goats) /1e9 #  sheep and goats
# 
# # 
# 
# # livestock protein masked to unsuitable areas in hyde_suitabilities.Rmd
# r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10 <-
#   mask(r_protein_from_AGB_kg_perpix, SI_5arcmin_below10_croplandmask)
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$dairy) /1e9 # 1.4 dairy
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$beef) /1e9 # 0.3 beef
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$sheep_and_goats) /1e9 # 0.4 sheep and goats


###-------------------------------------------- Calculate g protein per day per capita
  # based on FAOSTAT population in 2020 was 7791709 * 1000 persons
(1000 * global(r_protein_from_AGB_kg_perpix, fun = "sum", na.rm = T)) / # gram
  (1000* foodBalanceSheets_semiraw %>% filter(Area == "World", Item == "Population") %>% pull(Y2020)) / # persons
  365  # days in a year
# 3.4 g/capita/d with current herd structures

    # compare this to daily requirements (http://dx.doi.org/10.1016/S0140-6736(18)31788-4)
    # average person weight = 70kg and protein need = 0.8 per kg --> needs 56g of protein per day
((1000 * global(r_protein_from_AGB_kg_perpix, fun = "sum", na.rm = T)) / # gram of protein
  (1000* foodBalanceSheets_semiraw %>% filter(Area == "World", Item == "Population") %>% pull(Y2020)) / # persons
  365 ) /# days in a year
  (0.8*70) # 6% (if all population adults and distributed evenly...)

    # for who many persons this would be enough?
    # in 2020 a people would need 56g * 365d = 20440 g = 20.4kg protein in a year. Our esimate would be enough for 
(1000 * global(r_protein_from_AGB_kg_perpix, fun = "sum", na.rm = T)) / # total protein in grams
    (0.8*70*365) / # 20440 grams needed in a year
  1e6 


# ------------------------------------------------ comparisons to other studies
# In Van Zanten et al (2016) paper (Defining a land boundary for sustainable livestock consumption) they synthetised following numbers for beef (their Table 1)
van_zanten_protein_beef_g_capita_day <-
  c(5, 2, 1, 10, 2, 10, 0)
mean(van_zanten_protein_beef_g_capita_day) # mean 4.3 g per capita d!!
sd(van_zanten_protein_beef_g_capita_day)


van_zanten_protein_milk_g_capita_day <-
  c(2, 4, 8, 8, 16, 6)
mean(van_zanten_protein_milk_g_capita_day) # mean 7.3 t capita day


```





# Rasterize protein fraction - not needed here!

```{r}
#fbs_animals_wide_filled$protein_fraction %>% hist()

# combine fbs_animals_wide_filled with geometries

faoid_and_protein_fraction  <-
  fbs_animals_wide_filled %>%
  filter(Item == "Meat") %>% # Bovine meat
  dplyr::select(Area.Code, protein_fraction)

## create reclassification matrix
rcl_mat_countries_to_protein_fraction <-
  matrix(c(
    faoid_and_protein_fraction$Area.Code,
    faoid_and_protein_fraction$protein_fraction),
    ncol = 2)


## create raster where fao id has been replaced with protein_fraction
r_protein_fraction_Meat_from_FAOtables  <-
  cntry_raster %>%
  classify(rcl_mat_countries_to_protein_fraction,
           others =
             (fbs_animals_wide_filled %>%
                filter(Area == "World", Item == "Meat") %>%
                pull(protein_fraction)))# use global averages for missing countries

names(r_protein_fraction_Meat_from_FAOtables) <- "protein_fraction_Meat"
#plot(r_protein_fraction)
```




# Save rasters for intermediate use

```{r}
writeRaster(r_protein_from_AGB_kg_ha,
             here("Data", "Intermediate_input","r_protein_from_AGB_kg_ha.tif"),
            overwrite = T)


# writeRaster(r_kcal_from_AGB_MM_ha,
#               here("Data", "Intermediate_input","r_kcal_from_AGB_MM_ha.tif"),
#              overwrite = T)
# 
#  writeRaster(r_kcal_from_AGB_MM_perpix,
#               here("Data", "Intermediate_input","r_kcal_from_AGB_MM_perpix.tif"),
#              overwrite = T)



 writeRaster(r_protein_from_AGB_kg_perpix,
              here("Data", "Intermediate_input","r_protein_from_AGB_kg_perpix.tif"),
             overwrite = T)



```






