---
title: "Untitled"
author: "JP"
date: "`r Sys.Date()`"
output: html_document
---

# GAEZ attainable grass yields

"For most crops the yields are given in kg dry weight per hectare. For alfalfa, miscanthus, napier grass, reed canary grass, pasture legumes and grasses the yields are in 10 kg dry weight (above ground biomass) per hectare. For sugar beet and sugarcane the yields are in kg sugar per hectare and for olive and oil palm in kg oil per hectare. Cotton yields are given as kg lint per hectare."

! Voi olla että näitä kyseisiä biomassatasoja pitää skaalata AGB/GAEZ_AGB suhteella


Estimating AGB based protein is useful only in areas where amount of AGB protein is not known but where there could be some AGB based grazing production and thus AGB based protein.
In other words, grazing lands can expand only when there is cropland in a cell. We know the AGB based protein yields for most of these cells but not all of them.
Therefore, we need GAEZ-based results in regions where there is cropland but where the amount of AGB protein is not known


```{r}

r_attainable_yield <-
  list.files(here("Data", "Input", "gaez"), 
             pattern = ".tif", full.names = T) |> 
  map(rast) |> 
  rast()


# select grass related GAEZ attainable yields
r_attainable_yield_grasses <-
  c(r_attainable_yield$ycLr_alf,# Alfalfa
    r_attainable_yield$ycHr_nap) # Napier grass --> says that high input but actually input field is empty!!
    #, 
#    r_attainable_yield$ycLr_jtr, # Jatropha --> not good for animals (poisonous)
 #   r_attainable_yield$ycLr_mis, # Miscanthus
#    r_attainable_yield$ycLr_rcg) # Reed canary grass
  #  r_attainable_yield$ycLr_swg) # Switchgrass

r_attainable_yield_grasses <-
  terra::ifel(r_attainable_yield_grasses <= 0, NA,
              r_attainable_yield_grasses)

# Convert the unit to kg/ha
r_attainable_yield_grasses_kgDM_ha <-
  r_attainable_yield_grasses * 10

summary(0.42*r_attainable_yield_grasses_kgDM_ha) # napier grass > alfalfa yields > AGB yields




############################################## ADD LPJmL grass (C3) yields
##### LPJ
grass_lpj <- here("Data", "Input", "other", "yield_grass_rainfed_fao_yield_gap.asc") |> rast()
plot(grass_lpj)

# disaggregate from 0.5 degree to 5arcmin
grass_lpj_5arcmin <- disagg(grass_lpj, fact = 6)


# Combine with GAEZ
r_attainable_yield_grasses_kgDM_ha <-
  c(r_attainable_yield_grasses_kgDM_ha, grass_lpj_5arcmin)

plot(r_attainable_yield_grasses_kgDM_ha)
############################################## 



# calculate mean yield with na.rm = T
r_attainable_yield_grasses_kgDM_ha_mean <-
  mean(r_attainable_yield_grasses_kgDM_ha, na.rm = T)
plot(r_attainable_yield_grasses_kgDM_ha_mean)

# Save
writeRaster(r_attainable_yield_grasses_kgDM_ha_mean, 
            here("Data", "Intermediate_input", "r_GAEZ_LPJmL_attainable_yield_grasses_kgDM_ha_mean_nomask.tif"),
            overwrite = T)





###############################################  Relationship between AGB and GAEZ + LPJ yields
# mask gaez yield to pixels with AGB data
r_agb_2020_in_hyde_gl_kg_ha <-
  here("Data", "Input", "from_harddrive",
       "AGB2020_withIGPB6to10or16_in_hydeareas_kg_ha_5arcmin.tif") %>% 
  rast()

r_attainable_yield_grasses_kgDM_ha_mean_glmask <-
  mask(r_attainable_yield_grasses_kgDM_ha_mean,
       r_agb_2020_in_hyde_gl_kg_ha)


summary(r_agb_2020_in_hyde_gl_kg_ha) # median 360 mean 531 DM kg/ha
summary(r_attainable_yield_grasses_kgDM_ha_mean_glmask) # median 1500 mean 2500 DM kg/ha


r_agb_div_by_gaez_grasses <-
  classify(r_agb_2020_in_hyde_gl_kg_ha, cbind(NA,0)) / 
  classify(r_attainable_yield_grasses_kgDM_ha_mean_glmask, cbind(NA,0))

# inf to NA 
r_agb_div_by_gaez_grasses <- ifel(r_agb_div_by_gaez_grasses == Inf, NA, r_agb_div_by_gaez_grasses)

summary(r_agb_div_by_gaez_grasses) # 
mean(values(r_agb_div_by_gaez_grasses), na.rm = T ) #  3.2
median(values(r_agb_div_by_gaez_grasses), na.rm = T ) # 0.229 -- use this to scale the attn yields to AGB yields


# -------------------------------------------- results needed outside current gl pixels
# need to be masked to cropland areas with no grazing land!
# cropland outside grazing land
r_cl_outside_gl <- 
  mask(r_fraction_cl_0toNA, r_fraction_gl_0toNA, inverse = T)
plot(r_cl_outside_gl, main = "cl outside gl - AGB protein needed for these pixels")


r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl <-
  mask(r_attainable_yield_grasses_kgDM_ha_mean,
       r_cl_outside_gl)


r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl_agb_scaled <-
  r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl * 
  (median(values(r_agb_div_by_gaez_grasses), na.rm = T ) )



summary(0.42*r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl_agb_scaled) # med 262 mean 315 kg/ha

plot(r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl_agb_scaled,
     main = "GAEZ and LpJ avg grass yield (kg DM/ha) - scaled with AGB yields")
# We assume after scaling PUF (harvest coefficient) etc are already being take into account



############################### Convert to protein using FCR (assumed PUF = 0.42 already when scaling)
r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl <-
  r_attainable_yield_grasses_kgDM_ha_mean_on_cl_outside_gl_agb_scaled / r_fcr

names(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl) <- "prot_kg_ha"

summary(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl) # median 3, mean 3.8,,
# with agb protein median vas 2 and mean 3


plot(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl)


## protein per pixel
r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl <-
  r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl * 
  cellSize(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl, unit = "ha") *
  r_cl_outside_gl

names(r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl) <-
  "prot_kg_perpix"

global(r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl, "sum", na.rm = T)/1e9 # 1.1 mmt in X million ha


# Save
  # perpix
writeRaster(r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl,
            here("Data", "Intermediate_input", "avg_protein_from_GAEZ_and_LPJ_grasses_kg_perpix_on_cl_outside_gl.tif"),
            overwrite = T)


  # per ha
writeRaster(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl,
            here("Data", "Intermediate_input", "avg_protein_from_GAEZ_and_LPJ_grasses_kg_ha_on_cl_outside_gl.tif"),
            overwrite = T)

############ get the file
r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl <-
  here("Data", "Intermediate_input", "avg_protein_from_GAEZ_and_LPJ_grasses_kg_perpix_on_cl_outside_gl.tif") %>% 
  rast()

r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl <-
  here("Data", "Intermediate_input", "avg_protein_from_GAEZ_and_LPJ_grasses_kg_ha_on_cl_outside_gl.tif") %>% 
  rast()
############ get the file

```



# GAEZ/LPJ based AGB protein yields outside current gl production areas but existing yields on existing production areas


```{r}
  # perpix
r_agb_prot_existing_plus_new_areas_kg_perpix <-
  classify(r_protein_from_AGB_kg_perpix$current_herd_str, cbind(NA,0)) +
  classify(r_avg_protein_from_GAEZ_grasses_kg_perpix_on_cl_outside_gl, cbind(NA,0))

r_agb_prot_existing_plus_new_areas_kg_perpix <-
  classify(r_agb_prot_existing_plus_new_areas_kg_perpix, cbind(0,NA))
names(r_agb_prot_existing_plus_new_areas_kg_perpix) <- "protKgperpix_current_herd"

  # per ha
r_agb_prot_existing_plus_new_areas_kg_ha <-
  classify(r_protein_from_AGB_kg_ha$current_herd_str, cbind(NA,0)) + 
  classify(r_avg_protein_from_GAEZ_grasses_kg_ha_on_cl_outside_gl, cbind(NA,0))

r_agb_prot_existing_plus_new_areas_kg_ha <- classify(r_agb_prot_existing_plus_new_areas_kg_ha, cbind(0,NA))


(plt_AGB_protein_gaez_derived_and_existing <-
  create_index_map(r_index =
   r_agb_prot_existing_plus_new_areas_kg_ha,
                   tocrs = "ESRI:54030",
                   index_main_title = "AGB prot yield GAEZ-derived (outside gl) plus existing", # outside current croplands
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 1, 5, 10, 25, 50, Inf), # 1/10 of crops
                   breaknames = c("0-1","1-5", "5-10",
                                  "10-25","25-50",  ">50")))


names(r_agb_prot_existing_plus_new_areas_kg_ha) <- "protKgha_current_herd"


# SAVE
writeRaster(r_agb_prot_existing_plus_new_areas_kg_perpix,
            filename= here("Data", "Intermediate_input", 
                           "r_agb_prot_existing_plus_new_areas_kg_perpix.tif"),
            overwrite = T)

writeRaster(r_agb_prot_existing_plus_new_areas_kg_ha,
            filename= here("Data", "Intermediate_input", 
                           "r_agb_prot_existing_plus_new_areas_kg_ha.tif"),
            overwrite = T)


############ get the file
r_agb_prot_existing_plus_new_areas_kg_perpix <-
  here("Data", "Intermediate_input", "r_agb_prot_existing_plus_new_areas_kg_perpix.tif") %>% 
  rast()

r_agb_prot_existing_plus_new_areas_kg_ha <-
  here("Data", "Intermediate_input", "r_agb_prot_existing_plus_new_areas_kg_ha.tif") %>% 
  rast()
############ get the file




```





# Compare this with regression derived yields


