
Protein production capacity from aboveground biomass (AGB)


# This is just to get rid of error

```{r}
plot(cntry_raster) # Error in x$.self$finalize() : attempt to apply non-function
cntry_raster
```

# Calculate protein to energy conversion using FAOSTAT statistics

```{r}
# ---------------------------------------------------------- FAOSTAT download 
foodBalanceSheets_semiraw <-
  here("Data", "Input", "faostat", "foodBalanceSheets_E_All_Data_NOFLAG.csv") %>% 
  read.csv() %>% 
  dplyr::select(-Element.Code,-Unit ,-c(Y2011:Y2019))


my_elementlist_fao <- c(
  "Production",
  "Food supply quantity (kg/capita/yr)",
  "Food supply (kcal/capita/day)",
  "Protein supply quantity (g/capita/day)")

my_itemlist_fao_animals <- c("Meat", "Bovine Meat", "Mutton & Goat Meat")





# ---------------------------------------- FAOSTAT animal protein to kcal
fbs_animals <- foodBalanceSheets_semiraw %>% 
  filter(Item %in% my_itemlist_fao_animals,
         Element %in% my_elementlist_fao) 


 ## pivot wider
fbs_animals_wide <- fbs_animals %>% 
  pivot_wider(names_from = Element, values_from = Y2010) %>% 
  rename(prod_1000mt = "Production", # eg numbers in FIN equals Lukestat
         supply.KgCapYr = "Food supply quantity (kg/capita/yr)",
         supply.KcalCapD = "Food supply (kcal/capita/day)",
         prot.gCapD = "Protein supply quantity (g/capita/day)") %>% 
  
  
  ## how many kcal do we get from amount of meat containing one kg of protein - required for simple conversion from protein production to energy production
  ## create variable protein to kcal 
  mutate(one_kg_prot_to_kcal =  (supply.KcalCapD / prot.gCapD) * 1000,
         protein_fraction = (prot.gCapD*365/1000) / supply.KgCapYr) %>% # protein fraction for meat products -- not even needed?
  dplyr::select(-Item.Code) %>% 
  
  
  mutate_all(function(x) ifelse(is.nan(x) | is.infinite(x), NA, x))  # convert NaN values to NA






# fill empty columns (values 0 or NA) with global averages similardly as done earlier with crops
  ## ex with Bovine meat. First, find global average for one_kg_prot_to_kcal
# one_kg_prot_to_kcal_bovine <- fbs_animals_wide %>% 
#   filter(Item == "Bovine Meat", Area == "World") %>%
#   pull(one_kg_prot_to_kcal)
# 
#   ## then replace 0 or NA values with this value
# test_bovine  <- fbs_animals_wide %>% 
#   filter(Item == "Bovine Meat") %>% 
#   mutate(one_kg_prot_to_kcal = 
#            ifelse(is.na(one_kg_prot_to_kcal) | one_kg_prot_to_kcal == 0 , # if one_kg_prot_to_kcal fraction = NA or 0,
#                                    one_kg_prot_to_kcal_bovine, # fill with the global average value
#                                    one_kg_prot_to_kcal))  # otherwise let it just be




# do this with function for all meat items and fill also protein_fraction col if that's needed later
f_fill_missing_nutritient_values_animals<- function(animalcategory) {

   ## pull one_kg_prot_to_kcal for spesific animal species
  one_kg_prot_to_kcal_animalcategory <- fbs_animals_wide %>% 
  filter(Item == animalcategory, Area == "World") %>%
  pull(one_kg_prot_to_kcal)
  
  ## pull protein fraction for spesific crop
  protein_fraction_animalcategory <- fbs_animals_wide %>% 
  filter(Item == animalcategory, Area == "World") %>%
  pull(protein_fraction)
  
  
  ## fill empty one_kg_prot_to_kcal amd protein_fraction  with global averages
  fbs_animals_wide_filled <- fbs_animals_wide %>% 
    filter(Item == animalcategory) %>%
    
    
    ## fill one_kg_prot_to_kcal column
    mutate(one_kg_prot_to_kcal = 
           ifelse(is.na(one_kg_prot_to_kcal) | one_kg_prot_to_kcal == 0 , # if protein fraction = NA or 0,
                                   one_kg_prot_to_kcal_animalcategory, # fill with the global average value
                                   one_kg_prot_to_kcal)) %>% # otherwise let it just be    
    
    ## fill protein_fraction column
    mutate(protein_fraction = 
           ifelse(is.na(protein_fraction) | protein_fraction == 0 , 
                                   protein_fraction_animalcategory,
                                   protein_fraction)) 
  

  return(fbs_animals_wide_filled)
}





# Lapply this for all the animal species  --> returns a list
fbs_animals_wide_filled_list <- 
  lapply(my_itemlist_fao_animals, 
         f_fill_missing_nutritient_values_animals) 

# next we use bind_rows to make this reasonable looking dataframe
fbs_animals_wide_filled <- bind_rows(fbs_animals_wide_filled_list)  

```


# Aboveground biomass (AGB) om HYDE grazing lands and Feed Conversion Ratios (FCR)  

! add explanation to text that missing values have been filled with country's average AGB values?
! or use regression instead?


! Note that r_fraction_gl contains areas that are not visible in r_agb dataset

```{r}
# ---------------------------------------------- find HYDE grazing land fraction
r_fraction_gl <- 
  here("Data", "Input", "from_harddrive", ## add so that this file is in same place
       "r_fraction_hyde_grazing_lands_masked_to_MODIS_IGBP_areas_0toNA.tif") %>%   #### changed  was HYDE area only
  rast() # if value > 1 it should be just 1
plot(r_fraction_gl, main = "fraction of grazing lands in this study (HYDE-MODIS integration)")



# Simulated AGB (median over 2008-2012) masked to HYDE areas (some areas, where IGBP classes 6-10or16 are not covered are not present in this AGB data. But they are visible in r_fraction_gl)
r_agb_5y_med_in_hyde_gl_kg_ha <-
  here("Data", "Input", "from_harddrive",
       "5y_medAGB_withIGPB6to10or16_in_hydeareas_kg_ha_5arcmin.tif") %>% 
  rast()



 ## AGB per pixel without any cropland mask. present result kg km2 but this is for global sums
r_agb_kg_perpix <-
  r_agb_5y_med_in_hyde_gl_kg_ha * 
  cellSize(r_agb_5y_med_in_hyde_gl_kg_ha, unit = "ha") * 
  r_fraction_gl






#------------------------------------------------ FCR created in livestock_fcr.R
# Next we convert AGB to protein using FCR values of Mekonnen et al and Herrero et al
fcr_complete <-
  read.csv(here("Vilma_fcr_repo", "livestock_fcr_repo", "Outputs",
                "fcr_grazing_by_animalspecies_bycountry_HerreroMekonnen.csv")) %>% 
  as_tibble()


  ## combine with fao geometries and select production system = Grazing, and drop SG milk
adm_10m_fao_id_simple_combined_with_fcr <-
  fcr_complete %>%
  filter(prod_sys_2 == "Grazing", Animal.category != "Sg milk") %>%
 # left_join(adm_10m_fao_id_simple, ., by = c("FAO_ID" = "Area.Code"))
  left_join(adm10_simple_faoadded, ., 
             by = c("SOVEREIGNT" = "AreaName")) 
            #  by = c("FAO_ID" = "Area.Code")) # no South Sudan in fcr_complete



#------------------ Create FCR raster for spesific Animal.Category with function
f_FCR_raster <- function(animalcategory) {
  
  ## create reclassification matrix
  rcl_mat_countries_to_fcr <-
  adm_10m_fao_id_simple_combined_with_fcr %>% 
  filter(Animal.category == animalcategory) %>% 
  dplyr::select(FAO_ID, FCR_uniform_median) %>% # was FCR_truncated_median
  st_drop_geometry() %>% 
  as.matrix(., ncol = 2)


  ## classify. Creates a raster that shows FCR for different countries
  r_fcr_in_f  <- cntry_raster %>% # in_f refers that in function -- not global var
    classify(rcl_mat_countries_to_fcr, 
             others = 
               fcr_complete %>% filter(Animal.category == animalcategory) %>% 
               pull(FCR_uniform_median) %>% mean()) # fill others with mean value
  
  names(r_fcr_in_f) <- paste0(animalcategory, "_FCR")

return(r_fcr_in_f)
}


#f_FCR_raster("Dairy cattle") # test


# create average FCR raster based on mean of simulated FCRs of beef, cattle and sheepgoats
# fcr_average <-
# mean(
#   f_FCR_raster("Dairy cattle"),
#   f_FCR_raster("Beef cattle"),
#   f_FCR_raster("Sg meat")
# ) 
# names(fcr_average) <- "fcr_average"

 ## create fcr raster
r_fcr <- c(
  f_FCR_raster("Dairy cattle"),
  f_FCR_raster("Beef cattle"),
  f_FCR_raster("Sg meat")
 # fcr_average
)
names(r_fcr) <- c("fcr_dairy", "fcr_beef", "fcr_sheep_and_goat_meat")


plot(r_fcr)




```

# Current herd compositions

Unfortunately, this data cannot be shared. It can be requested from Mario Herrero and then from Kajsa Sahlin who processed the data further. Because the data is not openly available I builded the script so that elements that contains info about herd structures are kind of like addition parts and not included in main functions for example

```{r}
df_herd <- 
  read_excel(here("Data", "Input", "herrero",
                  "modified_animal_composition2000.xlsx"))
df_herd_long <- df_herd %>%
  pivot_longer(cols = c(Dairy_cattle, Beef_cattle, Sheep_and_goats),
               names_to = "Animal_type",
               values_to = "share")


df_herd_long <- df_herd_long %>%
  mutate(fao_from_dfcountryname = countrycode(.$Country, origin = "country.name",  destination = "fao"))
## warns that some are missing. However, either fao_from_iso3eh or fao_from_SOVEREIGNT includes most


adm_10m_fao_id_simple_combined_with_herd_structures <-
  left_join(adm10_simple_faoadded, df_herd_long, 
             by = c("FAO_ID" = "fao_from_dfcountryname")) 




#------------------ Create HERD raster for specific Animal Category with function
f_HERD_raster <- function(animalcategory) {
  
  ## create reclassification matrix
  rcl_mat_countries_to_herd <-
  adm_10m_fao_id_simple_combined_with_herd_structures %>% 
  filter(Animal_type  == animalcategory) %>% 
  dplyr::select(FAO_ID, share) %>%
  st_drop_geometry() %>% 
  as.matrix(., ncol = 2)

  ## classify. Creates a raster that shows HERD values for different countries
  r_herd_in_f  <- cntry_raster %>%
    classify(rcl_mat_countries_to_herd, 
             others = NA)
               # df_herd_long %>% filter(Animal_type == animalcategory) %>% 
               # pull(share ) %>% mean()) # fill others with mean value
  
   # Apply the land mask
  r_herd_in_f <- r_herd_in_f * land_mask
  
  names(r_herd_in_f) <- paste0(animalcategory, "_share")

  return(r_herd_in_f)
}

f_HERD_raster("Dairy_cattle")


# get rasters for each animal categories
r_herd <- c(
  f_HERD_raster("Dairy_cattle"),
  f_HERD_raster("Beef_cattle"),
  f_HERD_raster("Sheep_and_goats")
)
names(r_herd) <- c("dairy_share", "beef_share", "sheep_and_goats_share")
sum(r_herd) # test that sums to 100%
plot(r_herd)
```




# Convert AGB in HYDE grazing lands to protein

Calculate how much protein and kcal HYDE grazing lands can produce from AGB. Amount of produced protein is different for different animal species


Calculate this also for current herd structure. 


E.g. consider grid cell with AGB 1000 kg/ha and herd shares of 20% dairy, 50% beef and 30% sheep and goats. Let's assume FCRs in that cell are 200, 400 and 400 respectively. Now dairy cattle consumes 1000 * 0.20 kg of AGB per ha, beef cattle 1000 * 0.50 and sheepgoats 1000 * 0.30. Then these amount of feed they consume will be converted to protein: (1000 * 0.20)/200 + (1000 * 0.50)/400 + (1000 * 0.30)/400 



```{r}
# ------------------------------------------------ protein from AGB per ha
r_protein_from_065xAGB_kg_ha <-
  (0.65*r_agb_5y_med_in_hyde_gl_kg_ha) / r_fcr 

names(r_protein_from_065xAGB_kg_ha) <- c("dairy", "beef", "sheep_and_goats") #, "average"

plot(r_protein_from_065xAGB_kg_ha, 
     main = "protein from AGB when 65% of AGB is used ") 



# ---------------------------- protein that current herd structure could provide
  # kg/ha
r_protein_from_065xAGB_current_herd_str_kg_ha <-
  (0.65*r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$dairy_share) /
  r_fcr$fcr_dairy +
  (0.65*r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$beef_share) /
  r_fcr$fcr_beef +
  (0.65*r_agb_5y_med_in_hyde_gl_kg_ha * r_herd$sheep_and_goats_share) /
  r_fcr$fcr_sheep_and_goat_meat 
  
names(r_protein_from_065xAGB_current_herd_str_kg_ha) <- "current_herd_str"

# ---------------------- combine 
r_protein_from_065xAGB_kg_ha <- 
  c(r_protein_from_065xAGB_kg_ha, 
    r_protein_from_065xAGB_current_herd_str_kg_ha)

plot(r_protein_from_065xAGB_kg_ha) # this kind of fig also needed for kcals



# ------------------------------------------------ protein from AGB per pixel
r_protein_from_065xAGB_kg_perpix <- 
  r_protein_from_065xAGB_kg_ha *
  cellSize(r_protein_from_065xAGB_kg_ha, unit = "ha") *
  r_fraction_gl



# ---------------------------- add also fcr of current herd structures here
  # this is actually not used anywhere and it is not correct to claculate amount of protein originating from AGB by using this
# r_fcr_current_herd_str <-
#   r_fcr$fcr_dairy * r_herd$dairy_share +
#   r_fcr$fcr_beef * r_herd$beef_share +
#   r_fcr$fcr_sheep_and_goat_meat * r_herd$sheep_and_goats_share
# names(r_fcr_current_herd_str) <- "fcr_current_herd"
# plot(r_fcr_current_herd_str)  #
# 
# #kokeile tuleeko samanlainen proteiinimäärä jos laskee tällä
# r_protein_from_065xAGB_kg_ha2 <-
#   (0.65*r_agb_5y_med_in_hyde_gl_kg_ha) / r_fcr_current_herd_str 
# 
# r_protein_from_065xAGB_kg_ha2 %>% summary()
# r_protein_from_065xAGB_kg_ha$current_herd_str %>% summary()
```


# Convert protein production to kcal
AGB converted to protein above. Now we convert protein to kcal for all different animal species


```{r}
# kcal production raster, example with dairy cows
faoid_and_kcals_dairy  <- 
  fbs_animals_wide_filled %>%
  filter(Item == "Bovine Meat") %>%
  dplyr::select(Area.Code, one_kg_prot_to_kcal)

## create reclassification matrix
rcl_mat_countries_to_one_kg_prot_to_kcal <-
  matrix(c(
    faoid_and_kcals_dairy$Area.Code,
    faoid_and_kcals_dairy$one_kg_prot_to_kcal),
    ncol = 2)


## create raster where fao id has been replaced with one_kg_prot_to_kcal
r_one_kg_prot_to_kcal_dairy  <- 
  cntry_raster %>% 
  classify(rcl_mat_countries_to_one_kg_prot_to_kcal, 
           others = 
             (fbs_animals_wide_filled %>% 
                filter(Area == "World", Item == "Bovine Meat") %>%
                pull(one_kg_prot_to_kcal)))
            # use global averages for missing countries

plot(r_one_kg_prot_to_kcal_dairy, main = "one_kg_prot_to_kcal_dairy by country")


## multiply with protein production raster
plot((r_one_kg_prot_to_kcal_dairy * r_protein_from_065xAGB_kg_ha$dairy /1e6),
     main = "Amount of kcal dairy cattle could produce from AGB (MM kcal per ha)")








# ------------------------ the same for different animal species with a function
f_energy_from_protein_MMkcal_ha <- function(animalcategory_fao, 
                                            animalcategory_FCR) {

  ## get protein fraction for spesific crop species
  faoid_and_one_kg_prot_to_kcal <-
    fbs_animals_wide_filled %>%
    filter(Item == animalcategory_fao) %>%
    dplyr::select(Area.Code, one_kg_prot_to_kcal)

  ## reclassification matrix
  rcl_mat_countries_to_one_kg_prot_to_kcal <-
    matrix(c(
      faoid_and_one_kg_prot_to_kcal$Area.Code,
      faoid_and_one_kg_prot_to_kcal$one_kg_prot_to_kcal),
      ncol = 2)

  ## classify. Creates a raster that shows one_kg_to_prot for each country
  r_one_kg_prot_to_kcal  <- 
    cntry_raster %>% 
    classify(rcl_mat_countries_to_one_kg_prot_to_kcal, 
             others = 
               (fbs_animals_wide_filled %>%
                       filter(Area == "World", Item == animalcategory_fao) %>% 
                  pull(one_kg_prot_to_kcal))) # fill empty values with global averages
  
  
  # protein production raster (first chunck)
  r_protein_production_animals <-
    subset(r_protein_from_065xAGB_kg_ha, animalcategory_FCR) 
  

  ## kcal production raster
  r_kcal_MM <-
    r_protein_production_animals* r_one_kg_prot_to_kcal / 1e6
  names(r_kcal_MM) <- paste0(animalcategory_FCR, "_MMkcal_ha")

  return(r_kcal_MM) # MM per pix or per ha?
}


f_energy_from_protein_MMkcal_ha(animalcategory_fao = "Bovine Meat",
                          animalcategory_FCR = "dairy") %>% 
  plot(., main = "Amount of kcal dairy cows could produce from AGB (MM kcal per ha)")




# Derive energy production raster for all animal species
r_kcal_from_065xAGB_MM_ha <- c(
  f_energy_from_protein_MMkcal_ha(animalcategory_fao = "Bovine Meat",
                                  animalcategory_FCR = "dairy"), #
  
  f_energy_from_protein_MMkcal_ha(animalcategory_fao = "Bovine Meat",
                                  animalcategory_FCR = "beef"),
  
  f_energy_from_protein_MMkcal_ha(animalcategory_fao = "Mutton & Goat Meat",
                                  animalcategory_FCR = "sheep_and_goats") ,
  
  # add kcal produced by current herd structure
  f_energy_from_protein_MMkcal_ha(animalcategory_fao = "Meat",
                                  animalcategory_FCR = "current_herd_str")  
)

r_kcal_from_065xAGB_MM_ha
plot(r_kcal_from_065xAGB_MM_ha)


```



# Calculate global sums
Convert per ha values to perpix



```{r}
# Total protein produce
  # dairy 31 mmt, beef 7 mmt, sheep 6 mmt, current herd 11 mmt
global(r_protein_from_065xAGB_kg_perpix, fun = "sum", na.rm = T)/1e9


hist(r_protein_from_065xAGB_kg_perpix$dairy) # Some valoues sus high (over 2e5 and max > 8e5)
quantile(values(r_protein_from_065xAGB_kg_perpix$dairy), 0.99, na.rm = T)

  ## outliers removed 
# f_global_sum_without_outliers(r_protein_from_065xAGB_kg_perpix$dairy) /1e9 # 27 #was 41 dairy
# f_global_sum_without_outliers(r_protein_from_065xAGB_kg_perpix$beef) /1e9 # 6  was9 beef
# f_global_sum_without_outliers(r_protein_from_065xAGB_kg_perpix$sheep_and_goats) /1e9 # 6 #was 9 sheep_and_goats
# 
# 
# 
# global(f_raster_without_outliers(r_protein_from_065xAGB_kg_perpix$dairy), "sum", na.rm = T)/1e9 # 30
# trimmed_value_proteinAGB <-
#   quantile(values(r_protein_from_AGB_kg_perpix), 0.99, na.rm = T)
# r_protein_from_AGB_kg_perpix_trim <- r_protein_from_AGB_kg_perpix
# r_protein_from_AGB_kg_perpix_trim[r_protein_from_AGB_kg_perpix_trim > trimmed_value_proteinAGB] <- NA
# global(r_protein_from_AGB_kg_perpix_trim, fun = "sum", na.rm = T)/1e9 # dairy 35 mmt, beef and sheep 11 mmt --> !!!! should be different quantiles for diff layers! --- update with function



# total kcal production
r_kcal_from_065xAGB_MM_perpix <- 
  r_kcal_from_065xAGB_MM_ha * cellSize(r_kcal_from_065xAGB_MM_ha, unit ="ha") *
  r_fraction_gl
names(r_kcal_from_065xAGB_MM_perpix) <- c("dairy_MMkcal_perpix", "beef_MMkcal_perpix",
                                     "sheep_and_goats_MMkcal_perpix", "current_herd_str_MMkcal_perpix" )
# dairy 315 beef 69 sheep 105, current herd 153
global(r_kcal_from_065xAGB_MM_perpix*1e6, fun = "sum", na.rm = T)/1e12 

hist(r_kcal_from_065xAGB_MM_perpix*1e6) # sus high vals (over 2e9 for dairy)

  ## outliers removed --- these numbers in paper
# f_global_sum_without_outliers(r_kcal_from_065xAGB_MM_perpix$dairy_MMkcal_ha) /1e6 # 286  # was 440 dairy
# f_global_sum_without_outliers(r_kcal_from_065xAGB_MM_perpix$beef_MMkcal_ha) /1e6 # 65 # was 100 beef
# f_global_sum_without_outliers(r_kcal_from_065xAGB_MM_perpix$sheep_and_goats_MMkcal_ha) /1e6 # 98 # was 151 sheep_and_goats





# livestock protein masked to cropland areas --- uncomment when need to calculte these
# !!!!!!!!!!!! move to somewhere else -- maybe a place where all global sums are calculated?


# r_protein_from_AGB_kg_perpix_masked_to_croplands <-
#   mask(r_protein_from_065xAGB_kg_perpix, r_physical_areas_crops_sum_ha_perpix)
# 
# global(r_protein_from_AGB_kg_perpix_masked_to_croplands, "sum", na.rm = T)/1e9
# 
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$dairy) /1e9 #  dairy
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$beef) /1e9 #  beef
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands$sheep_and_goats) /1e9 #  sheep and goats
# 
# # 
# 
# # livestock protein masked to unsuitable areas in hyde_suitabilities.Rmd
# r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10 <-
#   mask(r_protein_from_065xAGB_kg_perpix, SI_5arcmin_below10_croplandmask)
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$dairy) /1e9 # 1.4 dairy
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$beef) /1e9 # 0.3 beef
# f_global_sum_without_outliers(r_protein_from_AGB_kg_perpix_masked_to_croplands_where_SIbelow10$sheep_and_goats) /1e9 # 0.4 sheep and goats
```





# Rasterize protein fraction - needed when calculating LUR -- maybe not even needed!

```{r}
#fbs_animals_wide_filled$protein_fraction %>% hist()

# combine fbs_animals_wide_filled with geometries

# kcal production raster, example with dairy cows
faoid_and_protein_fraction  <- 
  fbs_animals_wide_filled %>%
  filter(Item == "Meat") %>%
  dplyr::select(Area.Code, protein_fraction)

## create reclassification matrix
rcl_mat_countries_to_protein_fraction <-
  matrix(c(
    faoid_and_protein_fraction$Area.Code,
    faoid_and_protein_fraction$protein_fraction),
    ncol = 2)


## create raster where fao id has been replaced with protein_fraction
r_protein_fraction_Meat_from_FAOtables  <- 
  cntry_raster %>% 
  classify(rcl_mat_countries_to_protein_fraction, 
           others = 
             (fbs_animals_wide_filled %>% 
                filter(Area == "World", Item == "Meat") %>%
                pull(protein_fraction)))# use global averages for missing countries

names(r_protein_fraction_Meat_from_FAOtables) <- "protein_fraction_Meat"
#plot(r_protein_fraction)
```




# Save rasters for intermediate use

```{r}
writeRaster(r_protein_from_065xAGB_kg_ha,
             here("Data", "Intermediate_input","r_protein_from_065xAGB_kg_ha.tif"),
            overwrite = T)


writeRaster(r_kcal_from_065xAGB_MM_ha,
             here("Data", "Intermediate_input","r_kcal_from_065xAGB_MM_ha.tif"),
            overwrite = T)

writeRaster(r_kcal_from_065xAGB_MM_perpix,
             here("Data", "Intermediate_input","r_kcal_from_065xAGB_MM_perpix.tif"),
            overwrite = T)



writeRaster(r_protein_from_065xAGB_kg_perpix,
             here("Data", "Intermediate_input","r_protein_from_065xAGB_kg_perpix.tif"),
            overwrite = T)


writeRaster(r_protein_fraction_Meat_from_FAOtables,
             here("Data", "Intermediate_input","r_protein_fraction_Meat_from_FAOtables.tif"),
            overwrite = T)

```






