---
title: "Predicting crop yields"
author: "Johannes Piipponen"
date: "`r Sys.Date()`"
output: html_document
---

Estimate how much the crop protein yield would be in grazing lands that could be converted from grazing lands to croplands



# Data and packages

```{r}
# SI
SI_plant_spesific_5arcmin <- 
  here("Data", "Input", "Zabel22_SI",
       "plantspecific_suitabilititiesl_for_crops_1to17_1980_2009hist_5arcmin.tif") %>% 
  rast() # maybe kok[kok<1]<- NA

# this is not a good idea (plan was to run regression only for SI > 0 values but it does remove too much data)
#SI_plant_spesific_5arcmin <- classify(SI_plant_spesific_5arcmin, cbind(0, NA))

# Crop protein production yields
r_prot_allcrops_sum_kg_ha <-
  here("Data", "Intermediate_input", "protein_production_27crops_sum_kg_ha.tif") %>% 
  rast()


quantile(values(r_prot_allcrops_sum_kg_ha), probs = 0.95, na.rm = T) # 826.4333  

r_protein_and_SI_crops <- c(r_prot_allcrops_sum_kg_ha, SI_plant_spesific_5arcmin)





#plot(r_protein_and_SI_crops)
```


# Check correlations

```{r}
# check the data
# test1 <- r_prot_allcrops_sum_kg_ha
# test1[test1 >826.4333] <- NA
# r_protein_and_SI_test <- c(test1, SI_plant_spesific_5arcmin)
# 
# #pairs(r_protein_and_SI_test) # too slow
# 
# df_protein_and_SI_test <- as.data.frame(r_protein_and_SI_test, xy = T)
# df_protein_and_SI_test <- df_protein_and_SI_test %>%
#   na.omit()
# 
# correlation_matrix_crops <-
#   cor(df_protein_and_SI_test %>%  dplyr::select(-x,-y)) #
# 
# library(corrplot)
# corrplot(correlation_matrix_crops, method = "circle")
# 
# # plot --- no very visible linear trend
# ggplot(df_protein_and_SI_test[sample(nrow(df_protein_and_SI_test), 1e4), ], 
#        aes(x = crop_protein_kg_ha , y = plnt10)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   labs(title = "Scatter Plot",
#        x = "Crop Protein (kg/ha)",
#        y = "plnt1 Suitability") +
#   theme_minimal()

```

remove outliers function

```{r remove outliers function}

remove_outliers_quantiles <- function(df, lower, upper) {
  lower_quantile <- quantile(df$crop_protein_kg_ha, lower, na.rm = TRUE)
  upper_quantile <- quantile(df$crop_protein_kg_ha, upper, na.rm = TRUE)
  return(c(lower_quantile, upper_quantile))
}

```




# Linear regression model for different countries

In this section, we first create a list of countries with their geometric information. We then use lists to associate relevant rasters with each country. Next, we convert raster data into data frames and apply a linear regression using the lm() function to explain crop protein yield based on suitability. Outliers are now filtered out based on the 95th percentile of crop protein values --> can be some other outlier detection level as well!

We implement a safe wrapper function for linear regression modeling, which fits a model for each country using the filtered data. The model results are combined with the original country polygons, and the data is cleaned up by selecting relevant columns and unnesting the model results into a single data frame.

# prepare data for regression

```{r}
all_countries <- adm10_simple_faoadded

# create list of countries
country_sf_list <- all_countries %>%
  mutate(country_sf_list = map(ADMIN, ~filter(adm10_simple_faoadded, ADMIN == .x))) %>%
  pull(country_sf_list)
#country_sf_list[[53]] %>% plot() # FI

# find a relevant raster for each country by cropping and masking
country_raster_list <- map(country_sf_list, ~crop_and_mask(r_protein_and_SI_crops, .x))
#country_raster_list[[53]] %>% plot() # FI

# Convert to data frame
country_df_list <- map(country_raster_list, ~as.data.frame(.x, xy = T) %>% as_tibble())
#country_df_list[[53]] # FI


#  find outliers for every country (list again)
#outlier_level_list <- map_dbl(country_df_list, ~quantile(.x$crop_protein_kg_ha, probs = 0.95, na.rm = T)) 
#outlier_level_list[[53]] # 1243.164 FI


# outlier_level_list_1_99 <-  map(country_df_list, ~ remove_outliers_quantiles(.x, 0.01, 0.99))
# outlier_level_list_5_95 <-  map(country_df_list, ~ remove_outliers_quantiles(.x, 0.05, 0.95))
outlier_level_list_0_90 <-  map(country_df_list, ~ remove_outliers_quantiles(.x, 0.00, 0.90)) # best model



country_df_list <- map2(country_df_list, outlier_level_list_0_90, ~.x %>%
                                   filter(crop_protein_kg_ha >= .y[1] & crop_protein_kg_ha <= .y[2]) %>%
                                   dplyr::select(-c(x, y)))


# f_outlier_level_list <- function(lowXpercent, topXpercent) {
#   outlier_level_list <-  map(country_df_list, ~ remove_outliers_quantiles(.x, lowXpercent, topXpercent)) 
#   return(outlier_level_list)
# }


# Filtter so that outliers are removed. Also remove x and y
# country_df_list <- map2(country_df_list, outlier_level_list_1_99, ~.x %>%
#                                    filter(crop_protein_kg_ha >= .y[1] & crop_protein_kg_ha <= .y[2]) %>%
#                                    dplyr::select(-c(x, y)))

# country_df_list <- map2(country_df_list, outlier_level_list_5_95, ~.x %>%
#                                    filter(crop_protein_kg_ha >= .y[1] & crop_protein_kg_ha <= .y[2]) %>%
#                                    dplyr::select(-c(x, y)))

# country_df_list <-
#   map2(country_df_list, f_outlier_level_list(0.00, 0.9), ~.x %>%
#                                    filter(crop_protein_kg_ha >= .y[1] & crop_protein_kg_ha <= .y[2]) %>%
#                                    dplyr::select(-c(x, y)))
# # # # 
# # 
# country_df_list[[53]] # # data for regression



```

# Regression
```{r}


#  test if the distribution is skewed for the right (then gamma distribution could be an option:
# country_df_filtered_list[[57]][,1] %>% pull() %>% hist()



# do regression basd on this filtered data -- use  safe function of purrr to avoind troubles
# Wrapper function for linear regression
fit_lm_possibly <- function(data) {
  
  model_output <- lm(crop_protein_kg_ha ~ 
                       plnt1 + plnt2 + plnt3 + plnt4 + plnt5 + 
                       plnt6 + plnt7 + plnt8 + plnt9 + plnt10 +
                       plnt11 + plnt12 + plnt13 + plnt14 + plnt15 +
                       plnt16 + plnt17,
                     data = data) 
  
  # model_coef <- model_output %>% 
  #   broom::tidy() %>%
  #   dplyr::select(term, estimate, p.value)
  
  model_coef <- model_output %>% 
    broom::tidy() %>%
    dplyr::select(term, estimate,  p.value) %>% 
    pivot_wider(., names_from = term, values_from = c(estimate, p.value))
  
  
  model_stats <- model_output %>% 
    broom::glance()
  
  # Combine model_coef and model_stats into a single row data frame
  result <- model_coef %>%
    mutate(r.squared = model_stats$r.squared, # these from model stats
           adj.r.squared = model_stats$adj.r.squared,
           AIC = model_stats$AIC,
           BIC = model_stats$BIC,
           nobs = model_stats$nobs,
           sigma = model_stats$sigma,
           statistic = model_stats$statistic,
           p.value_Ftest = model_stats$p.value) %>% 
     rename(estimate_intercept = "estimate_(Intercept)")
  
  

  return(result)
}




# Safe version of the linear regression function
fit_lm_safe <- purrr::possibly(fit_lm_possibly, otherwise = NULL)


# model_list 

model_list <- map(country_df_list, fit_lm_safe)
model_list[[53]] # results for Finland




# combine nested country_df_list with polygons
all_countries <- all_countries %>% 
  mutate(country_df_list = country_df_list,
         model_list = model_list)



# clean
all_countries_cleaned <- all_countries %>% 
  dplyr::select(ADMIN, ISO_A3_EH, REGION_UN, geom, model_list)



all_countries_cleaned <- all_countries_cleaned %>%
  mutate(model_list = map(model_list, ~ as_tibble(.x))) %>%
  unnest(cols = model_list, keep_empty = T)



#View(all_countries_cleaned)




```


# Use nearest neighbour to give values for countries with insignificant regression values



```{r}
# 1) Find countries where the model is not significant. For these countries, only the distance (distance from a country with a F_testp-value > 0.05 to a significant country) is needed.

model_significant <- all_countries_cleaned %>% filter(p.value_Ftest < 0.05) #  countries
model_nonsignificant <- all_countries_cleaned %>% filter(p.value_Ftest >= 0.05 | is.na(p.value_Ftest)) #  countries

model_significant %>% nrow()
median(all_countries_cleaned$adj.r.squared, na.rm=T)
median(all_countries_cleaned$AIC, na.rm=T)



model_significant %>%   dplyr::select(adj.r.squared) %>% summary() 
model_nonsignificant %>%   dplyr::select(adj.r.squared) %>% summary() 

# Find distances for those 78 countries. First, find centroids (needed for distances)
p_adm0_centroids_st <- all_countries_cleaned %>% 
  dplyr::select(ADMIN, geom) %>% 
  st_centroid()


# Filter centroids for countries with significant and non-significant models
significant_centroids <- p_adm0_centroids_st %>%
  filter(ADMIN %in% model_significant$ADMIN) # 124

nonsignificant_centroids <- p_adm0_centroids_st %>%
  filter(ADMIN %in% model_nonsignificant$ADMIN) # 78  --------> etäisyydet tarvitaan vain näille (esim Uruguay listalla)




# Calculate distances between non-significant and significant model centroids
distTemp <- st_distance(nonsignificant_centroids, significant_centroids) %>% 
  as_tibble() # 77 rows because we have 77 nonsignificant_centroids. 125 cols as we have 125 significant countries
# Row = centroid of non-significant countries AND
# Column = centroid of significant countries BETWEEN
# So (1,1) from the centroid of country Dhekelia Sovereign Base Area to the centroid of country Indonesia
# So (1,2) from the centroid of country Dhekelia Sovereign Base Area to the centroid of country Malaysia





#Find the index of the closest significant model for each non-significant model
closest_significant_indices <- apply(distTemp, 1, which.min) # for each row, find the minimum distance, and select the significant column (col) with the shortest distance to the non-significant country (row)


#Attach the closest significant models to the non-significant models.
model_nonsignificant_with_closest <- model_nonsignificant %>%
  mutate(closest_significant_model = model_significant$ADMIN[closest_significant_indices], 
         closest_significant_index = closest_significant_indices)

# Combine the table "significant" with non-significant ones, for which the closest significant country is known.
all_countries_with_replacement <- model_significant %>%
  bind_rows(model_nonsignificant_with_closest) 









# -------------------------add intercept and filled suitabiliities for variables
# Fill estimate_intercept first
updated_regression_results <- all_countries_with_replacement %>%
  mutate(
    estimate_intercept_filled = 
      ifelse(p.value_Ftest >= 0.05 | is.na(p.value_Ftest),
             model_significant$estimate_intercept[closest_significant_index],
             estimate_intercept)
  )




# Define function to fill estimate values
fill_estimate <- function(df, var) {
  # Create the variable names dynamically based on the input variable
  var_estimate <- paste0("estimate_", var) # The name of the estimate variable for this plant
  var_estimate_filled <- paste0(var_estimate, "_filled") # The name of the filled estimate variable for this plant
  
  # Create the new filled estimate variable
  df <- df %>%
    mutate(
      # Use !! to unquote the variable names so that they are evaluated as variables in the dataframe
      !!var_estimate_filled := 
        ifelse(p.value_Ftest >= 0.05 | is.na(p.value_Ftest),
               model_significant[[var_estimate]][closest_significant_index],
               !!sym(var_estimate))
    )
  return(df)
}
# := on erikoisoperaattori, joka on saatavilla joissakin R-paketeissa, kuten data.table ja rlang. Se mahdollistaa nimien luomisen ja arvojen määrittämisen dynaamisesti ohjelmointikoodissa. Tässä tapauksessa se mahdollistaa dynaamisesti luotujen muuttujien määrittämisen mutate-funktion sisällä.
# 
# sym on funktio, joka kuuluu rlang-pakettiin. Se ottaa merkkijonon ja palauttaa symbolin. Symbolit ovat ohjelmointikielissä yleisiä ja ne edustavat muuttujien nimiä. rlang::sym muuntaa merkkijonon symboliksi, jonka tidyverse-funktiot tunnistavat muuttujaksi. Tämä on välttämätöntä, kun muuttujan nimet luodaan dynaamisesti ohjelmointikoodin suorituksen aikana.
# 

vars_to_fill <- names(SI_plant_spesific_5arcmin)


# Apply the fill_estimate function to each variable using purrr::reduce
updated_regression_results <- 
  reduce(vars_to_fill, fill_estimate, .init = updated_regression_results)
# Tässä purrr::reduce ottaa kolme argumenttia:
# vars_to_fill on lista muuttujista, joihin fill_estimate-funktiota sovelletaan.
# fill_estimate on funktio, jota sovelletaan jokaiseen elementtiin listassa.
# .init = updated_regression_results kertoo, että toimintoa sovelletaan aluksi updated_regression_results-datakehikkoon.


# Now we need to convert NA values to 0 (filled columns) as otherwise e.g. Mongolia will be completely NA. This is because some plant suitabilities are NA for Mongolooa
# Specify the columns to mutate
vars_to_mutate <-  updated_regression_results %>%
  dplyr::select(ends_with("_filled")) %>% 
  st_drop_geometry()%>%
  colnames()



# Replace NA values with 0
updated_regression_results <- updated_regression_results %>%
  mutate(across(all_of(vars_to_mutate), ~replace_na(., 0)))
```




# check how predicted protein yield map looks like 


```{r}

# Intercept
r_intercept_plants_SI<- 
  rasterize(updated_regression_results, template_rast_5arcmin, 
            field="estimate_intercept_filled")



# Beta coefficients
# Function to rasterize a given variable
rasterize_var <- function(var) {
  var_beta_coef_filled <- paste0("estimate_", var, "_filled") # The name of the filled beta coefficient variable for this plant
  assign(paste0("r_beta_coef_", var), 
         rasterize(updated_regression_results, template_rast_5arcmin, field = var_beta_coef_filled),
         envir = .GlobalEnv)
}

# Use map to apply the function to each variable
r_beta_coefficients <- map(vars_to_fill, rasterize_var) %>% 
  rast()






# Copy paste way

r_predicted_protein_yield_kg_ha <-
  r_intercept_plants_SI + 
  r_beta_coefficients$estimate_plnt1_filled * SI_plant_spesific_5arcmin$plnt1 + 
  r_beta_coefficients$estimate_plnt2_filled * SI_plant_spesific_5arcmin$plnt2 + 
  r_beta_coefficients$estimate_plnt3_filled * SI_plant_spesific_5arcmin$plnt3 +
  r_beta_coefficients$estimate_plnt4_filled * SI_plant_spesific_5arcmin$plnt4 +
  r_beta_coefficients$estimate_plnt5_filled * SI_plant_spesific_5arcmin$plnt5 +
  r_beta_coefficients$estimate_plnt6_filled * SI_plant_spesific_5arcmin$plnt6 +
  r_beta_coefficients$estimate_plnt7_filled * SI_plant_spesific_5arcmin$plnt7 +
  r_beta_coefficients$estimate_plnt8_filled * SI_plant_spesific_5arcmin$plnt8 +
  r_beta_coefficients$estimate_plnt9_filled * SI_plant_spesific_5arcmin$plnt9 +
  r_beta_coefficients$estimate_plnt10_filled * SI_plant_spesific_5arcmin$plnt10 +
  r_beta_coefficients$estimate_plnt11_filled * SI_plant_spesific_5arcmin$plnt11 +
  r_beta_coefficients$estimate_plnt12_filled * SI_plant_spesific_5arcmin$plnt12 +
  r_beta_coefficients$estimate_plnt13_filled * SI_plant_spesific_5arcmin$plnt13 +
  r_beta_coefficients$estimate_plnt14_filled * SI_plant_spesific_5arcmin$plnt14 +
  r_beta_coefficients$estimate_plnt15_filled * SI_plant_spesific_5arcmin$plnt15 +
  r_beta_coefficients$estimate_plnt16_filled * SI_plant_spesific_5arcmin$plnt16 +
  r_beta_coefficients$estimate_plnt17_filled * SI_plant_spesific_5arcmin$plnt17


# We need to mask out areas where OVERALL SUITABILITY < 1. Otherwise high crop yield in Tibet for ecample although there cannot be crops based on SI
# this means that we assume there cannot be crop production on areas where SI = 0
SI_5arcmin_gte1 <- terra::ifel(SI_5arcmin < 1, NA, SI_5arcmin)
r_predicted_protein_yield_kg_ha <-
  mask(r_predicted_protein_yield_kg_ha,
       SI_5arcmin_gte1)

  # also negative values can be cpnverted to 0 
r_predicted_protein_yield_kg_ha <- 
  classify(r_predicted_protein_yield_kg_ha, 
           matrix(c(-Inf, 0, 0), ncol = 3, byrow = TRUE))


 # there are also a couple of values that are > 1e6 so we remove them
quantile(values(r_predicted_protein_yield_kg_ha), probs = 0.99, na.rm = T) #563.5783 
# 
r_predicted_protein_yield_kg_ha <-
  classify(r_predicted_protein_yield_kg_ha,
           matrix(c(563.5783, Inf, NA), ncol = 3, byrow = TRUE))
# #
# 

# 
# summary(r_predicted_protein_yield_outside_cl_but_in_gl_areas_and_in_suitable_areas_kg_ha)
# summary(r_prot_allcrops_sum_kg_ha)  
  
pal_protein_lajolla <- scico(n = 6, palette = "lajolla", end = 0.85) 

(plt_crop_protein_predicted <- 
  create_index_map(r_index = r_predicted_protein_yield_kg_ha,
                   tocrs = "ESRI:54030",
                   index_main_title = "PREDICTED crop protein yields on areas with overall suitability > 1 ",
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 10, 50, 100, 250, 500, Inf),
                   breaknames = c("0-10","10-50", "50-100",
                                  "100-250","250-500",  ">500")))   







# create mask. First outside cl mask as this is where we need data
r_predicted_protein_yield_outside_cl_kg_ha <-
  mask(r_predicted_protein_yield_kg_ha,
       r_physical_areas_crops_sum_ha_perpix, inverse = T)


  # moreover gl mask needed (otherwise shows all the possible areas)
r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha <-
  mask(r_predicted_protein_yield_outside_cl_kg_ha,
       r_fraction_gl_0toNA)

# croplandmask -- then we can see how close predited values are to actual values
r_predicted_protein_yield_croplandmask_kg_ha <-
  mask(r_predicted_protein_yield_kg_ha,
       r_physical_areas_crops_sum_ha_perpix)






(plt_crop_protein <- 
  create_index_map(r_index = r_prot_allcrops_sum_kg_ha, 
                   tocrs = "ESRI:54030",
                   index_main_title = "Protein production of SPAM food crops in 2010",
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 10, 50, 100, 250, 500, Inf),
                   breaknames = c("0-10","10-50", "50-100",
                                  "100-250","250-500",  ">500"))) 



(plt_crop_protein_predicted_outside_cl_but_in_gl <- 
  create_index_map(r_index = r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha,
                   tocrs = "ESRI:54030",
                   index_main_title = "PREDICTED crop protein yields outside current croplands", #  
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 10, 50, 100, 250, 500, Inf),
                   breaknames = c("0-10","10-50", "50-100",
                                  "100-250","250-500",  ">500"))) 



writeRaster(r_predicted_protein_yield_kg_ha,
            filename = here("Data", "Intermediate_input",
                            "r_predicted_crop_protein_yield_kg_ha.tif"),
            overwrite = T)

writeRaster(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha,
            filename = here("Data", "Intermediate_input",
                            "r_predicted_crop_protein_yield_outside_cl_but_in_gl_areas_kg_ha.tif"),
            overwrite = T)
```

# predicted outside current production areas but existing yields on existing production areas

```{r}

 # crops
r_crop_prot_existing_plus_new_areas_kg_ha <-
  classify(r_prot_allcrops_sum_kg_ha, cbind(NA,0)) + 
  classify(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha, cbind(NA,0))

r_crop_prot_existing_plus_new_areas_kg_ha <-
  classify(r_crop_prot_existing_plus_new_areas_kg_ha, cbind(0,NA))


(plt_crop_protein_predicted_and_existing <- 
  create_index_map(r_index = r_crop_prot_existing_plus_new_areas_kg_ha,
                   tocrs = "ESRI:54030",
                   index_main_title = "Crop prot yield predicted plus existing", #  
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 10, 50, 100, 250, 500, Inf),
                   breaknames = c("0-10","10-50", "50-100",
                                  "100-250","250-500",  ">500"))) 





writeRaster(r_crop_prot_existing_plus_new_areas_kg_ha,
            filename= here("Data", "Intermediate_input", 
                           "r_crop_prot_existing_plus_new_areas_kg_ha.tif"),
            overwrite = T)
```



# Global sums

```{r}
# laske globaalit summat kakista --- liian suuri viivakuvion proteiini tulee!


#---------------------------------------------------------------- Perpix test
# current cropland areas
r_predicted_protein_yield_croplandmask_kg_perpix <-
  r_predicted_protein_yield_croplandmask_kg_ha * 
  cellSize(r_predicted_protein_yield_croplandmask_kg_ha, unit = "ha") *
  r_fraction_cl_0toNA

global(r_predicted_protein_yield_croplandmask_kg_perpix, "sum", na.rm=T)/1e9 # 213




  # outside current cl areas but in gl areas
r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_perpix <-
  r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha *
  cellSize(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha, unit ="ha") *
  r_fraction_gl_0toNA

global(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_perpix, "sum", na.rm=T)/1e9 # 137



#---- can run only after suitability.Rmd
# r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test10 <-
#   (r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha/ 1000) *
#   cellSize(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha, unit = "ha")*
#   r_potential_cropland_expansion_outside_croplands_fraction[[10]]
# # 
# 
# r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test75 <- 
#   (r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha/ 1000) * 
#   cellSize(r_predicted_protein_yield_outside_cl_but_in_gl_areas_kg_ha, unit = "ha")*
#   r_potential_cropland_expansion_outside_croplands_fraction[[75]]
# 
# 
# plot(r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test10)
# plot(r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test75)
# 
# 
# global(r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test10, "sum", na.rm = T)/1e6 # 144 mmt
# global(r_predicted_protein_yield_outside_cl_but_in_gl_areas_mt_perpix_test75, "sum", na.rm = T)/1e6 # 9mmt

```


