# Gaez 

Miten saadaan selville onko GAEZ dataa saatavilla kaikille tarvittaville alueille?
--> siis meidän gl+cl alueille, joilla SI > 0


Työn vaiheet
1) tunnista alueet joilla muutos gl --> cl mahdollista
2) etsi GAEZ:n ehdottamat mahdolliset satotasot näille alueille
3) tutki kuinka suuren osuuden pellosta kasvilajit SPAM datan mukaan käyttävät (esim vehnä 30% ja maissi 70% kyseisten kasvilajien kokonaismaanäkäytöstä). --> Tämä tehdään admin1 tasolla (maataso liian karkea)
4) laske GAEZ sadot em kohdat huomioiden eri kasvilajeille (pitää muuntaa kuiva-aineesta tuoresadoiksi)
5) muuta eri kasvilajien sato proteiiniksi ja yhdistä
( 6) vertaa omiin regressioestimaatteihin)


# Relevant area for GAEZ crops

```{r}

# ne gl alueet joilla ei ole cl alueita
# LISÄKSI: SI < 1 pois
gl_areas_with_no_current_cl <- mask(r_fraction_gl_0toNA, r_fraction_cl_0toNA, inverse = T)
plot(gl_areas_with_no_current_cl, main = "gl areas with no current cl")

# SI_5arcmin already correct SI data
gl_areas_with_no_current_cl_with_SIgte1 <-
  mask(gl_areas_with_no_current_cl,
       ifel(SI_5arcmin < 1, NA, SI_5arcmin) ) # SI already updated tata but doulbe check
plot(gl_areas_with_no_current_cl_with_SIgte1, main = "suitable gl areas with no current cl")

gl_areas_with_no_current_cl_with_SIgte1

```

# GAEZ attainable yields

Convert to fresh yields


```{r}
# List attainable yield raster (all where pattern is tif)
r_attainable_yield <-
  list.files(here("Data", "Input", "gaez"), 
             pattern = ".tif", full.names = T) |> 
  map(rast) |> 
  rast()

r_attainable_yield # 48 lyr
names(r_attainable_yield) # 48 lyr
r_attainable_yield$ycLr_con |> plot()

# Create df of the GAEZ names under Theme ID 4 'Suitability and Attainable Yield'
gaez_names <-  tibble(
  Crop_acronym = c("alf", "ban", "bck", "brl", "bsg", "cab", "car", "chk", "cit", "coc", 
                   "cof", "con", "cot", "cow", "csv", "flx", "fml", "grd", "grm", "jtr", 
                   "mis", "mze", "nap", "oat", "olp", "olv", "oni", "pea", "phb", "pig", 
                   "pml", "rcd", "rcg", "rcw", "rsd", "rub", "rye", "sfl", "soy", "spo", 
                   "srg", "sub", "suc", "swg", "tea", "tob", "tom", "whe", "wpo", "yam"),
  Crop_name = c("Alfalfa", "Banana", "Buckwheat", "Barley", "Biomass sorghum", "Cabbage", 
                "Carrot", "Chickpea", "Citrus", "Cocoa", "Coffee", "Coconut", "Cotton", 
                "Cowpea", "Cassava", "Flax", "Foxtail millet", "Groundnut", "Gram", "Jatropha", 
                "Miscanthus", "Maize", "Napier grass", "Oat", "Oil palm", "Olive", "Onion", 
                "Dry pea", "Phaseolus bean", "Pigeonpea", "Pearl millet", "Dryland rice", 
                "Reed canary grass", "Wetland rice", "Rapeseed", "Rubber", "Rye", "Sunflower", 
                "Soybean", "Sweet potato", "Sorghum", "Sugar beet", "Sugar cane", "Switchgrass", 
                "Tea", "Tobacco", "Tomato", "Wheat", "White potato", "Yam"))


# Tibble of GAEZ crop names combined with SPAM crops names
# Includes also conversio factor that GAEZ has used to conver fresh yields to DM yields
# conversion factors can be found in table 9-1 in gaez documentation pdf
combined_data_gaez_spam <- tribble(
  ~gaezname, ~gaez_long_name, ~spamname, ~spam_long_name, ~conversion_factor,
  "ban", "Banana", "BANA", "Bananas", 0.25,
  "brl", "Barley", "BARL", "Barley and products", 0.87,
  "chk", "Chickpea", "CHIC", "Pulses, Other and products", 1.00,
  "con", "Coconut", "CNUT", "Coconuts - Incl Copra", 1.00,
  "cow", "Cowpea", "COWP", "Pulses, Other and products", 1.00,
  "csv", "Cassava", "CASS", "Cassava and products", 0.35,
  "grd", "Groundnut", "GROU", "Groundnuts", 0.65,
  "mze", "Maize", "MAIZ", "Maize and products", 0.86,
  "soy", "Soybean", "SOYB", "Soybeans", 0.90,
  "srg", "Sorghum", "SORG", "Sorghum and products", 0.87,
  "whe", "Wheat", "WHEA", "Wheat and products", 0.87,
  "pml", "Pearl millet", "PMIL", "Millet and products", 0.90,
  "phb", "Phaseolus bean", "BEAN", "Beans", 1.00,
  "pig", "Pigeonpea", "PIGE", "Pulses, Other and products", 1.00,
  "spo", "Sweet potato", "SWPO", "Sweet potatoes", 0.25,
  "wpo", "White potato", "POTA", "Potatoes and products", 0.20,
  "yam", "Yam", "YAMS", "Yams", 0.35,
  "pea", "Dry pea", "OPUL", "Pulses, Other and products", 1.00,
  "rcd", "Dryland rice", "DRYLANDRICE", "Rice and products", 0.87,
  "rcw", "Wetland rice", "WETLANDRICE", "Rice and products", 0.87,
  "tom", "Tomato", "VEGETOMATOES", "Vegetables", 1.00,
  "cab", "Cabbage", "VEGECAPPAGES", "Vegetables", 1.00,
  "oni", "Onion", "VEGEONION", "Vegetables", 1.00,
  "cit", "Citrus", "TROF", "Fruits, other", 1.00
) |> 
  mutate(gaezname_full = paste0("ycLr_", gaezname))
# OCER, ORTS, LENT, PLNT missing at least --> should we add oat or rye?

############# Select SPAM crops from the GAEZ raster layers
# Create a vector of GAEZ crop acronyms that match the SPAM crops
gaez_acronyms <- 
  gaez_names %>% 
  filter(Crop_acronym %in% combined_data_gaez_spam$gaezname) %>% 
  pull(Crop_acronym) # 24

# Create a pattern to match the crop acronyms in the raster layer names
pattern_gaez <- paste0("ycLr_", gaez_acronyms, collapse = "|")

# Select the raster layers that match the pattern
r_GAEZattnyld_for_spamcrops <- 
  subset(r_attainable_yield, 
         grep(pattern_gaez, names(r_attainable_yield)))

# Convert values (0 or smaller) to NA
r_GAEZattnyld_for_spamcrops <- 
  terra::ifel(r_GAEZattnyld_for_spamcrops <= 0, NA,
              r_GAEZattnyld_for_spamcrops) # still gaez names

plot(r_GAEZattnyld_for_spamcrops)

# Change names to SPAM names --> order must remain the same
ordered_indices <- 
  match(combined_data_gaez_spam$gaezname_full,
        names(r_GAEZattnyld_for_spamcrops))

r_GAEZattnyld_for_spamcrops <-
  r_GAEZattnyld_for_spamcrops[[ordered_indices]]

names(r_GAEZattnyld_for_spamcrops) <- combined_data_gaez_spam$spamname





# ---------------------------------------------------- convert DM to fresh yields
# r_GAEZattnyld_for_spamcrops_fresh_whea <-
#   subset(r_GAEZattnyld_for_spamcrops, "WHEA") /
#   (combined_data_gaez_spam |>
#      filter(spamname == "WHEA") |>
#      pull(conversion_factor) )

# or with GAEZ names
# r_GAEZattnyld_for_spamcrops_fresh_whea <-
#   subset(r_GAEZattnyld_for_spamcrops, "ycLr_whe") /
#   (combined_data_gaez_spam |>
#      filter(gaezname_full == "ycLr_whe") |>
#      pull(conversion_factor) )


# create function for this
f_DM_to_fresh <-
  function(select_spamname) {
    
    r_GAEZattnyld_for_spamcrops_fresh <-
      
    (subset(r_GAEZattnyld_for_spamcrops, 
            select_spamname)) /
      
      (combined_data_gaez_spam |> 
     filter(spamname == select_spamname) |> 
     pull(conversion_factor) )
    
    return(r_GAEZattnyld_for_spamcrops_fresh)
  }

# run function for each unique spamnames
r_GAEZattnyld_for_spamcrops_fresh <-
  purrr::map(unique(combined_data_gaez_spam$spamname),
             f_DM_to_fresh) |>  # creates a list
  rast() # creates a rast out of list --> 24 lyr
  





# ------------------------------------------------------------------------------- EDIT: all of these could be plotted in facet in supplementary --> only when converted to fresh matter
plot(r_GAEZattnyld_for_spamcrops_fresh)


# -------------------------------------------------------- mask to relevant area
r_GAEZattnyld_for_spamcrops_fresh_relevant_areas <-
  mask(r_GAEZattnyld_for_spamcrops_fresh,
       gl_areas_with_no_current_cl_with_SIgte1)

plot(r_GAEZattnyld_for_spamcrops_fresh_relevant_areas) # fresh


```





# Create new fbs_crop_wide where additional GAEZ crops

```{r}
foodBalanceSheets_semiraw <-
  here("Data", "Input", "faostat", "foodBalanceSheets_E_All_Data_NOFLAG.csv") %>%
  read.csv() %>%
  dplyr::select(-Element.Code,-Unit , -c(Y2010:Y2019), -c(Y2021)) #
names(foodBalanceSheets_semiraw)

# ---------------------------------------------------------- crops
my_itemlist_fao <- c(  # in mapspam webpage order
  "Wheat and products", "Rice and products","Maize and products",
  "Barley and products", "Millet and products", "Sorghum and products",
  "Cereals, Other", "Potatoes and products", "Sweet potatoes",
  "Yams", "Cassava and products", "Roots, Other",  
  "Beans", "Peas", "Pulses, Other and products",
  "Soyabeans","Groundnuts", "Coconuts - Incl Copra",
  "Bananas", "Plantains", "Fruits, other", "Vegetables" # categories for SPAM names
)

my_elementlist_fao <- c(
  "Production","Food supply quantity (kg/capita/yr)",
  "Food supply (kcal/capita/day)", "Protein supply quantity (g/capita/day)"
)


## filter
fbs_crops <- foodBalanceSheets_semiraw  %>% 
  filter(Item %in% my_itemlist_fao,
         Element %in% my_elementlist_fao)



## to wide format. Select only relevant variables
fbs_crops_wide_gaez <- fbs_crops %>% 
  pivot_wider(names_from = Element, values_from = Y2020) %>% # 2020 
  rename(prod_1000mt = "Production",
         supply.KgCapYr = "Food supply quantity (kg/capita/yr)",
         supply.KcalCapD = "Food supply (kcal/capita/day)",
         prot.gCapD = "Protein supply quantity (g/capita/day)") 



## add column where item names corresponds to mapsmap names
## see Definitions and standards - Food Balances (2010-) for the classes
## cannot add information for all crops like this -- add MILL, COWP, PIGE, OPUL, LENT, TROF afterwards
fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  mutate(spamname = case_when(
    Item == "Wheat and products" ~ "WHEA",
    Item == "Rice and products"  ~ "RICE",
    Item == "Barley and products" ~ "BARL",
    Item == "Maize and products" ~ "MAIZ",
    Item == "Millet and products"  ~ "PMIL", 
    #    Item == "Millet and products"  ~ "MILL", # ei toimi koska 2x Millet and products
    Item == "Sorghum and products"  ~ "SORG",
    Item == "Cereals, Other"   ~ "OCER",
    Item == "Potatoes and products" ~ "POTA",
    Item == "Sweet potatoes"  ~ "SWPO",
    Item == "Yams"  ~ "YAMS",
    Item == "Cassava and products"   ~ "CASS",    
    Item == "Roots, Other"  ~ "ORTS",
    Item == "Beans"  ~ "BEAN",
    Item == "Pulses, Other and products"  ~ "CHIC", 
    #    Item == "Pulses, Other and products" ~ "COWP", 
    #    Item == "Pulses, Other and products" ~ "PIGE",     
    #    Item == "Pulses, Other and products"  ~ "OPUL", 
    #    Item == "Pulses, Other and products"  ~ "LENT", 
    Item == "Soyabeans"  ~ "SOYB",
    Item == "Groundnuts"   ~ "GROU",
    Item == "Coconuts - Incl Copra"  ~ "CNUT",
    Item == "Bananas"  ~ "BANA",  
    Item == "Plantains"  ~ "PLNT",
    Item == "Fruits, other"  ~ "TROF",    
    #    Item == "Fruits, other"  ~ "TEMF",    
    Item == "Vegetables"  ~ "VEGE"
  ))


# could be done with function but was faster like this..
additional_rows <- list(
  fbs_crops_wide_gaez %>% filter(Item == "Millet and products") %>% mutate(spamname = "MILL"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "COWP"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "PIGE"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "OPUL"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "LENT"),
  fbs_crops_wide_gaez %>% filter(Item == "Fruits, other") %>% mutate(spamname = "TEMF"),
  
  # !!
  # These rows added only when combining GAEZ attainable yields 
  fbs_crops_wide_gaez %>% filter(Item == "Rice and products") %>% mutate(spamname = "DRYLANDRICE"),
  fbs_crops_wide_gaez %>% filter(Item == "Rice and products") %>% mutate(spamname = "WETLANDRICE"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGETOMATOES"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGECAPPAGES"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGEONION"))

fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  bind_rows(additional_rows)









# ------------------------- Calculate fraction of protein and kcal in a product (conversion factor) 
## protein and kcal
fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  mutate(protein_fraction = (prot.gCapD * 365 / 1000) / supply.KgCapYr,
         kcalPerKg = (supply.KcalCapD * 365)/ supply.KgCapYr)  %>% 
  
  ## remove cases where supply = 0 (cases that yield NA values)  
  ## # also converts NaN to NA
  mutate_all(function(x) ifelse(is.nan(x) | is.infinite(x), NA, x)) 




# ---------------------------------------------------------- Fill empty rows
# Create a function that fills empty values of cols protein_fraction and kcalPerKg
f_fill_missing_nutritient_values_crops <- function(cropname_mapspam) {
  
  ## pull protein fraction for spesific crop
  protein_fraction_spamcrop <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam, Area == "World") %>%
    pull(protein_fraction)
  
  
  ## pull KcalPerKg for spesific crop
  KcalPerKg_spamcrop <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam, Area == "World") %>%
    pull(kcalPerKg)
  
  
  ## fill empty protein fraction and kcalPerKg with global averages
  fbs_crops_wide_gaez_filled <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam) %>%
    ## fill protein_fraction column
    mutate(protein_fraction = 
             ifelse(is.na(protein_fraction) | protein_fraction == 0 , # if protein fraction = NA or 0,
                    protein_fraction_spamcrop, # fill with the global average value
                    protein_fraction)) %>% # otherwise let it just be
    
    ## fill KcalPerKg column
    mutate(kcalPerKg = 
             ifelse(is.na(kcalPerKg) | kcalPerKg == 0 , 
                    KcalPerKg_spamcrop,
                    kcalPerKg)) 
  
  return(fbs_crops_wide_gaez_filled)
}



test1 <- f_fill_missing_nutritient_values_crops("SOYB") # returns a df

# Lapply this for all the crops --> returns a list
fbs_crops_wide_gaez_filled_list <- 
  lapply(fbs_crops_wide_gaez$spamname |> unique(),
         f_fill_missing_nutritient_values_crops) 




# next we use bind_rows to make this reasonable looking dataframe
fbs_crops_wide_gaez_filled <- bind_rows(fbs_crops_wide_gaez_filled_list)



```



# Convert yields to protein

```{r}
#------------------------------------------------------- crop protein production
  ## create a function to convert crop production to protein
f_crop_protein_production_kg_ha_GAEZ <- function(cropname_mapspam) {
  
  ## get protein fraction for spesific crop species
  faoid_and_prot_frac <-
    fbs_crops_wide_gaez_filled %>%
    filter(spamname == cropname_mapspam) %>%
    dplyr::select(Area.Code, protein_fraction)
  
  ## reclassification matrix
  rcl_mat_countries_to_prot_frac <-
    matrix(c(
      faoid_and_prot_frac$Area.Code,
      faoid_and_prot_frac$protein_fraction),
      ncol = 2)
  
  ## classify. Creates a raster that shows protein fraction for each country
  
    # first define values for others
  fbs_avg_proteinfract_of_a_crop <-
  fbs_crops_wide_gaez_filled %>%
  filter(spamname == cropname_mapspam, Area == "World") %>%
  pull(protein_fraction)
  
  r_fraction_prot  <- 
    cntry_raster %>% #  south sudan?
    classify(rcl_mat_countries_to_prot_frac, 
             others = fbs_avg_proteinfract_of_a_crop)
  
  
  # production raster  ----------- gaez data here!!
  r_production_crop <- 
    subset(r_GAEZattnyld_for_spamcrops_fresh_relevant_areas,
           cropname_mapspam) 
           
  
  
  ## protein production raster
  r_prot <-
    r_production_crop * r_fraction_prot
  names(r_prot) <- paste0(cropname_mapspam, "_prot")
  
  return(r_prot)
}

( gaez_MAIZ_prot <- f_crop_protein_production_kg_ha_GAEZ("MAIZ") )
( gaez_DRYLANDRICEprot <- f_crop_protein_production_kg_ha_GAEZ("DRYLANDRICE") )



# Use map() to apply this function to all the crops
r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha <-
  map(combined_data_gaez_spam$spamname, # väärä nimi
      f_crop_protein_production_kg_ha_GAEZ) |> 
  rast()


# Combine DRYLANDRICE_prot"  "WETLANDRICE_prot" "VEGETOMATOES_prot" "VEGECAPPAGES_prot" "VEGEONION_prot
r_VEGE_prot <-
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGETOMATOES_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGECAPPAGES_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGEONION_prot) |> 
  sum(na.rm=T)
names(r_VEGE_prot) <- "VEGE_prot"


r_RICE_prot <-
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$DRYLANDRICE_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$WETLANDRICE_prot) |> 
  sum(na.rm=T)
names(r_RICE_prot) <- "RICE_prot"




r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha <- 
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha, r_VEGE_prot, r_RICE_prot) |> 
  
  # remove some layers
  subset(c("DRYLANDRICE_prot", "WETLANDRICE_prot", "VEGETOMATOES_prot", "VEGECAPPAGES_prot", "VEGEONION_prot"), negate = T)




r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha |> plot() # check the areas, e.g. South Sudan


## just a test
# test_gaez_prot_yields <-
#   sum(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha, na.rm=T)
# plot(test_gaez_prot_yields)
# 
# test_gaez_prot_yields_south_sudan <-
#   crop_and_mask(test_gaez_prot_yields,
#        (adm_10m %>% filter(ADMIN == "South Sudan") %>% dplyr::select(ADMIN) |> 
#   as("Spatial") |>   vect()))
# 
# plot(test_gaez_prot_yields_south_sudan, main ="Potential GAEZ yields for Sudan - no cropmix") 
```

# SPAM crop mix --> takes few hours to run

We use this to define how much SPAM crops are grown in different regions.
This is used to calculate the protein production of the crops outside current cl areas.
So we assume that crops could be grown in the same proportions as they are grown now.

```{r}
r_physical_areas_ha_perpix <-
  here("Data", "Intermediate_input","r_physical_areas_ha_perpix.tif") |> 
  rast() # 27 different crops

# fraction of dif crop species in 5arcmin cell
r_fraction_cl_by_species <-
  r_physical_areas_ha_perpix / # e.g. 50 ha WHEA per cell
  r_physical_areas_crops_sum_ha_perpix # yhteensä esim 2000 ha viljelykasveja tässä solussa
# Tämän pitäisi olla nyt sellainen, että selviää paljonko esim vehnäala solussa on suhteessa kaikkien viljelykasvien alaan
# eli vehnä ala / kaikki viljelykasvit ala   molemmat siis hehtaareissa

r_fraction_cl_by_species
summary(r_fraction_cl_by_species$WHEA)
plot(r_fraction_cl_by_species$WHEA)
#sum(r_fraction_cl_by_species, na.rm=T) |> summary() # min and max = 1



# -----------------------------------------  Load GADM ADMIN 0, ADMIN 1, ADMIN 2
gadm_path <- 
  here("Data", "Input", "other", "polygons","gadm_410-levels.gpkg")
( layers <- st_layers(gadm_path) )

gadm_admin0 <- st_read(gadm_path, layer = "ADM_0") 
gadm_admin1 <- st_read(gadm_path, layer = "ADM_1")
gadm_admin2 <- st_read(gadm_path, layer = "ADM_2")

dim(gadm_admin1) # 3662 12
dim(gadm_admin2) # 47217 14

gadm_admin0 <- gadm_admin0 |>
  mutate(rowNbr = row_number()) 

gadm_admin1 <- gadm_admin1 |> 
  mutate(rowNbr = row_number()) |>
  dplyr::select(rowNbr, COUNTRY, GID_1, NAME_1, geom)

gadm_admin2 <- gadm_admin2 |>
  mutate(rowNbr = row_number()) |>
  dplyr::select(rowNbr, COUNTRY, GID_1, NAME_1, GID_2, NAME_2, geom)


# ---------------------------------------------------------------- extract crop mix
# ADMIN 0
df_crop_mix_adm0 <- 
  extract(r_fraction_cl_by_species, gadm_admin0, fun = mean, na.rm= T) 
# Convert NaN values in cols WHEA:VEGE to NA
df_crop_mix_adm0 <- df_crop_mix_adm0 |> 
  mutate(across(WHEA:VEGE, ~replace(., is.nan(.), NA)))

# convert NA to 0 (this we only do for admin0
df_crop_mix_adm0 <- df_crop_mix_adm0 |>
  mutate(across(WHEA:VEGE, ~replace(., is.na(.), 0)))

# Combine with admin1 geometries
df_crop_mix_adm0 <- 
  left_join(gadm_admin0, df_crop_mix_adm0, by = c("rowNbr" = "ID"))



# ADMIN 1
df_crop_mix_adm1 <- 
  extract(r_fraction_cl_by_species, gadm_admin1, fun = mean, na.rm= T) # 10 sec

df_crop_mix_adm1 <- df_crop_mix_adm1 %>%
  mutate(across(WHEA:VEGE, ~replace(., is.nan(.), NA)))

df_crop_mix_adm1 <- 
  left_join(gadm_admin1, df_crop_mix_adm1, by = c("rowNbr" = "ID"))



# ADMIN 2
df_crop_mix_adm2 <-
  extract(r_fraction_cl_by_species, gadm_admin2, fun = mean, na.rm= T)

df_crop_mix_adm2 <- df_crop_mix_adm2 %>%
  mutate(across(WHEA:VEGE, ~replace(., is.nan(.), NA)))

df_crop_mix_adm2 <- 
  left_join(gadm_admin2, df_crop_mix_adm2, by = c("rowNbr" = "ID"))


# ----------------------------------------------------------- rasterize crop mix
# Create the directory for intermediate inputs if it doesn't exist
output_dir <- here("Data", "Intermediate_input", "delete_these_intermed_inputs")
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}


# Function for rasterizing crop mix and saving to disk
f_rasterize_crop_mix_1arcmin <- function(my_admin_df, cropspecies) {
  
  r_crop_share_of_agland <- rasterize(my_admin_df, template_rast_1arcmin, 
                                      field = cropspecies)
  names(r_crop_share_of_agland) <- paste0("share_", cropspecies)
  
  # Save raster to disk
  output_path <- here(output_dir, paste0("share_", cropspecies, ".tif"))
  writeRaster(r_crop_share_of_agland, filename = output_path, overwrite = TRUE)
  
  return(r_crop_share_of_agland)
}

#test <- f_rasterize_crop_mix_1arcmin(df_crop_mix_adm1, "WHEA") # 1min
# get a file we just did
#rast(here(output_dir, "share_WHEA.tif")) |> plot()




### ADMIN 0
r_crop_share_of_agland_adm0_1arcmin <- 
  lapply(names(r_physical_areas_ha_perpix), function(cropspecies) {
    f_rasterize_crop_mix_1arcmin(my_admin_df = df_crop_mix_adm0, cropspecies)
  }) |> 
  rast() # 

# Save the raster to Intermediate_inputs
writeRaster(r_crop_share_of_agland_adm0_1arcmin, 
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm0_1arcmin.tif"),
            overwrite = TRUE)


### ADMIN 1
# Run function for all crops and for dataframe df_crop_mix_adm1
r_crop_share_of_agland_adm1_1arcmin <- 
  lapply(names(r_physical_areas_ha_perpix), function(cropspecies) {
    f_rasterize_crop_mix_1arcmin(my_admin_df = df_crop_mix_adm1, cropspecies)
  }) |> 
  rast() # 30min


writeRaster(r_crop_share_of_agland_adm1_1arcmin, 
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm1_1arcmin.tif"),
            overwrite = TRUE)



### ADMIN 2
r_crop_share_of_agland_adm2_1arcmin <-
  lapply(names(r_physical_areas_ha_perpix), function(cropspecies) {
    f_rasterize_crop_mix_1arcmin(my_admin_df = df_crop_mix_adm2, cropspecies)
  }) |> 
  rast() # less than 1hour

writeRaster(r_crop_share_of_agland_adm2_1arcmin, 
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm2_1arcmin.tif"),
            overwrite = TRUE)

plot(r_crop_share_of_agland_adm2_1arcmin)
# Optionally, clean up intermediate files or the entire directory
#unlink(output_dir, recursive = TRUE)


######### get the rasters we created
r_crop_share_of_agland_adm0_1arcmin  <-
  here("Data", "Intermediate_input",
       "r_crop_share_of_agland_adm0_1arcmin.tif") |> 
  rast()

r_crop_share_of_agland_adm1_1arcmin <-
  here("Data", "Intermediate_input",
       "r_crop_share_of_agland_adm1_1arcmin.tif") |> 
  rast()

r_crop_share_of_agland_adm2_1arcmin <-
  here("Data", "Intermediate_input",
       "r_crop_share_of_agland_adm2_1arcmin.tif") |> 
  rast()


# total share should be 1 (sum of 27 layers = 1)
# Funktio tarkistamaan ja rajoittamaan osuuksien summan arvoon 1
limit_to_one <- function(r) {
  total <- app(r, sum, na.rm = TRUE)
  r_corrected <- ifel(total > 1, r / total, r)
  return(r_corrected)
}


r_crop_share_of_agland_adm0_1arcmin <-
  limit_to_one(r_crop_share_of_agland_adm0_1arcmin)

r_crop_share_of_agland_adm1_1arcmin <-
  limit_to_one(r_crop_share_of_agland_adm1_1arcmin)

r_crop_share_of_agland_adm2_1arcmin <-
  limit_to_one(r_crop_share_of_agland_adm2_1arcmin)
 # 1 h
#plot(r_crop_share_of_agland_adm2_1arcmin)
# sum(r_crop_share_of_agland_adm2_1arcmin, na.rm=T) # saa vain arvoja 1
# summa on pienempi vai silloin jos polygoni saa NA arvoja (ne kts myöh)


# ----------------------------------------------------------------- to 5arcmin
# use modal as aggregation function
r_crop_share_of_agland_adm0_5arcmin <- 
  aggregate(r_crop_share_of_agland_adm0_1arcmin, fact=5, fun = "modal", na.rm = T,
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm0_5arcmin.tif"),
            overwrite = T)


r_crop_share_of_agland_adm1_5arcmin <- 
  aggregate(r_crop_share_of_agland_adm1_1arcmin, fact=5, fun = "modal", na.rm = T,
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm1_5arcmin.tif"),
            overwrite = T)

r_crop_share_of_agland_adm2_5arcmin <-
  aggregate(r_crop_share_of_agland_adm2_1arcmin, fact=5, fun = "modal", na.rm = T,
            filename = here("Data", "Intermediate_input",
                            "r_crop_share_of_agland_adm2_5arcmin.tif"),
            overwrite = T)
#sum(r_crop_share_of_agland_adm2_5arcmin, na.rm=T) |> summary()- # fev vals > 1


plot(r_crop_share_of_agland_adm0_5arcmin$share_WHEA)
plot(r_crop_share_of_agland_adm1_5arcmin$share_WHEA)
plot(r_crop_share_of_agland_adm2_5arcmin$share_WHEA)



# ------------------------------------- fill NA values (when NA in all of the layers) of adm2 with adm1 values, then, if still NA values, fill those from ADMIN 0 data

# Tarkistetaan, missä soluissa kaikissa layereissä on NA arvoja
r_crop_share_of_agland_adm2_allNA <-
  allNA(r_crop_share_of_agland_adm2_5arcmin)

plot(r_crop_share_of_agland_adm2_allNA, main = "no cl in these cells") # 1lyr -- for these NA cells we try to get values from adm1



# Apply the cover function to fill NA values in admin2 raster with values from admin1 raster for each crop layer --> only when all layers have NA values

fill_na_with_admin1 <- function(crop) {
  admin1_layer <- r_crop_share_of_agland_adm1_5arcmin[[crop]]
  admin2_layer <- r_crop_share_of_agland_adm2_5arcmin[[crop]]
  mask_layer <- mask(admin2_layer, 
                     r_crop_share_of_agland_adm2_allNA)
  filled_layer <- cover(mask_layer, admin1_layer)
  names(filled_layer) <- crop
  return(filled_layer)
}

# Käytetään funktiota kaikille layereille
r_crop_share_of_agland_adm2_filled_from_adm1 <-
  lapply(names(r_crop_share_of_agland_adm1_5arcmin),
         fill_na_with_admin1) |> 
  rast() # 4min




# .. if still NA values, fill from ADMIN 0
r_crop_share_of_agland_adm2_filled_from_adm1_allNA <-
  allNA(r_crop_share_of_agland_adm2_filled_from_adm1)


fill_na_with_admin0 <- function(crop) {
  admin0_layer <- r_crop_share_of_agland_adm0_5arcmin[[crop]]
  admin2_layer_filled <- r_crop_share_of_agland_adm2_filled_from_adm1[[crop]] 
  mask_layer <- mask(admin2_layer_filled, 
                     r_crop_share_of_agland_adm2_filled_from_adm1_allNA)
  filled_layer <- cover(mask_layer, admin0_layer)
  names(filled_layer) <- crop
  return(filled_layer)
}


# Käytetään funktiota kaikille layereille
r_crop_share_of_agland <-
  lapply(names(r_crop_share_of_agland_adm0_5arcmin), 
         fill_na_with_admin0) |> 
  rast()

# use limit to one function again to ensure the total cl fraction is max 1
r_crop_share_of_agland <- limit_to_one(r_crop_share_of_agland)

writeRaster(r_crop_share_of_agland, 
            here("Data", "Intermediate_input",
                 "r_SPAMcrops_share_of_agland_adm2_filled_5arcmin.tif"), overwrite = T)

# ---------------------------------------------- tests
plot(r_crop_share_of_agland_adm2_5arcmin$share_WHEA)
plot(r_crop_share_of_agland_adm2_filled_from_adm1$share_WHEA)
plot(r_crop_share_of_agland$share_WHEA)

sum(r_crop_share_of_agland, na.rm=T) |> summary()
sum(r_crop_share_of_agland, na.rm=T) |> plot()

r_crop_share_of_agland_south_sudan <-
  crop_and_mask(r_crop_share_of_agland,
       (adm_10m %>% filter(ADMIN == "South Sudan") %>% dplyr::select(ADMIN) |>
  as("Spatial") |>   vect()))
plot(r_crop_share_of_agland_south_sudan)



############### get the file
r_crop_share_of_agland <- 
  here("Data", "Intermediate_input",
       "r_SPAMcrops_share_of_agland_adm2_filled_5arcmin.tif") |> 
  rast()
names(r_crop_share_of_agland)
```


# Arvioi nyt crop mix soluille jotka ovat cl alueiden ulkopuolella

Eli maskataan relevantille alueelle (alueelle joka nykyisten cl ulkopuolella mutta joka kuuluu nykyisiin gl alueisiin)

```{r}
# First GAEZ attn yields perpix (for certain SPAM crops)
r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix <-
  r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha * # esim vehnää 3000 kg/ha (gl alueilla joilla SI > 1)
  cellSize(template_rast_5arcmin, unit = "ha") * # solun koko
  gl_areas_with_no_current_cl_with_SIgte1 # maatalousmaan osuus solussa, esim 20 % (gl alueet joilla SI > 1)

r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix
#plot(r_GAEZattnyld_for_spamcrops_relevant_areas_perpix)



  
# Arvioidaan sato kullekin pix niin että kerrotaan proteiinisato sillä osuudella, jota kyseistä kasvia nyt viljellään

r <- r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix

r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix <- sum(
  c(
  r$WHEA_prot * r_crop_share_of_agland$share_WHEA,
  r$RICE_prot * r_crop_share_of_agland$share_RICE,
  r$BARL_prot * r_crop_share_of_agland$share_BARL,
  r$MAIZ_prot * r_crop_share_of_agland$share_MAIZ,
  r$PMIL_prot * r_crop_share_of_agland$share_PMIL,
  #r$MILL_prot * r_crop_share_of_agland$MILL,
  r$SORG_prot * r_crop_share_of_agland$share_SORG,
  # r$OCER_prot * r_crop_share_of_agland$OCER,
  r$CASS_prot * r_crop_share_of_agland$share_CASS,
  r$POTA_prot * r_crop_share_of_agland$share_POTA,
  r$SWPO_prot * r_crop_share_of_agland$share_SWPO,
  r$YAMS_prot * r_crop_share_of_agland$share_YAMS,
  #r$ORTS_prot * r_crop_share_of_agland$ORTS,
  r$BEAN_prot * r_crop_share_of_agland$share_BEAN,
  r$CHIC_prot * r_crop_share_of_agland$share_CHIC,
  r$COWP_prot * r_crop_share_of_agland$share_COWP,
  r$PIGE_prot * r_crop_share_of_agland$share_PIGE,
  r$OPUL_prot * r_crop_share_of_agland$share_OPUL,
  # r$LENT_prot * r_crop_share_of_agland$LENT,
  r$SOYB_prot * r_crop_share_of_agland$share_SOYB,
  r$GROU_prot * r_crop_share_of_agland$share_GROU,
  r$CNUT_prot * r_crop_share_of_agland$share_CNUT,
  r$BANA_prot * r_crop_share_of_agland$share_BANA,
  #r$PLNT_prot * r_crop_share_of_agland$PLNT,
  r$TROF_prot * r_crop_share_of_agland$share_TROF,
  # r$TEMF_prot * r_crop_share_of_agland$TEMF,
  r$VEGE_prot * r_crop_share_of_agland$share_VEGE),
  na.rm = TRUE)

r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix
names(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix) <-
  "gaez_prot_outside_cl"

### global sums
global(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix, "sum", na.rm = T)/1e9 # -- 70 mmt much lower than estimation with regression! (about 19% of existing crop prot production)
# Using regression, this amount outside cl was almost 4x higher than this (273 mmt





# Convert back to kg per ha
r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha <-
  r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix /
  (cellSize(template_rast_5arcmin, unit = "ha") * gl_areas_with_no_current_cl_with_SIgte1)

names(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha) <-
  "gaez_prot_outside_cl"

plot(ifel(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha > 0 ,1,NA))

# save perpix raster
writeRaster(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix,
            filename= here("Data", "Intermediate_input",
                           "r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix.tif"),
            overwrite = T)

writeRaster(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha,
            filename= here("Data", "Intermediate_input",
                           "r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha.tif"),
            overwrite = T)



####### get the file
r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix <-
  here("Data", "Intermediate_input",
       "r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix.tif") |> 
  rast()

r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha <-
  here("Data", "Intermediate_input",
       "r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha.tif") |> 
  rast()



## Again crop and mask to south sudan
test_gaez_prot_in_south_sudan <-
  crop_and_mask(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha,
                (adm_10m %>% filter(ADMIN == "South Sudan") %>% dplyr::select(ADMIN) |>
  as("Spatial") |>   vect()))
plot(test_gaez_prot_in_south_sudan) # these outside current croplands! and there is potential
```

# GAEZ-based crop protein yields outside current cl production areas but existing yields on existing production areas

```{r}
# crops
  # perpix
r_crop_prot_existing_plus_new_areas_kg_perpix <-
  classify(1000*r_prot_allcrops_sum_mt_perpix, cbind(NA, 0)) +
  classify(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix, cbind(NA, 0))

r_crop_prot_existing_plus_new_areas_kg_perpix <-
  classify(r_crop_prot_existing_plus_new_areas_kg_perpix, cbind(0, NA))

names(r_crop_prot_existing_plus_new_areas_kg_perpix) <-
  "crop_protein_kg_perpix"
  
  # per ha
r_crop_prot_existing_plus_new_areas_kg_ha <-
  classify(r_prot_allcrops_sum_kg_ha, cbind(NA,0)) + 
  classify(r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha, cbind(NA,0))

r_crop_prot_existing_plus_new_areas_kg_ha <-
  classify(r_crop_prot_existing_plus_new_areas_kg_ha, cbind(0,NA))





# SAVE
writeRaster(r_crop_prot_existing_plus_new_areas_kg_perpix,
            filename= here("Data", "Intermediate_input",
                           "r_crop_prot_existing_plus_new_areas_kg_perpix.tif"),
            overwrite = T)


writeRaster(r_crop_prot_existing_plus_new_areas_kg_ha,
            filename= here("Data", "Intermediate_input",
                           "r_crop_prot_existing_plus_new_areas_kg_ha.tif"),
            overwrite = T)



############## get the file
r_crop_prot_existing_plus_new_areas_kg_perpix <-
  here("Data", "Intermediate_input",
       "r_crop_prot_existing_plus_new_areas_kg_perpix.tif") |> 
  rast()


r_crop_prot_existing_plus_new_areas_kg_ha <-
  here("Data", "Intermediate_input",
       "r_crop_prot_existing_plus_new_areas_kg_ha.tif") |> 
  rast()



```



# Plot

```{r}
pal_protein_lajolla <- scico(n = 6, palette = "lajolla", begin = 0.15, end = 1, direction = -1) # previously ended 0.85



r_separate_existing_and_GAEZ_derived_protein_yields <-
  c(r_prot_allcrops_sum_kg_ha, 
    r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_ha)

names(r_separate_existing_and_GAEZ_derived_protein_yields) <- c("Existing", "GAEZ-derived")
plot(r_separate_existing_and_GAEZ_derived_protein_yields)



(plt_crop_protein_GAEZ_derived_and_existing <- 
  create_index_map(r_index = r_separate_existing_and_GAEZ_derived_protein_yields,
                   tocrs = "ESRI:54030",
                   index_main_title = "Crop prot yield existing and GAEZ derived areas", #  
                   index_label = "[kg/ha]",
                   colorpal = pal_protein_lajolla,
                   breakvals = c(0, 10, 50, 100, 250, 500, Inf),
                   breaknames = c("0-10","10-50", "50-100",
                                  "100-250","250-500",  ">500"))) 



```





# Additional - estimate totalcrop yields (not protein) to SPAM yields
Are attainable yields just considerably lower in general?



```{r}
# # WHEA
# r_gaez_estimate_attn_yield_kg_perpix_WHEA <-
#   r_GAEZattnyld_for_spamcrops_fresh$WHEA * # 
#   r_physical_areas_ha_perpix$WHEA 
# # MAIZ
# r_gaez_estimate_attn_yield_kg_perpix_MAIZ <-
#   r_GAEZattnyld_for_spamcrops_fresh$MAIZ * # 
#   r_physical_areas_ha_perpix$MAIZ
# # SOYB
# r_gaez_estimate_attn_yield_kg_perpix_SOYB <-
#   r_GAEZattnyld_for_spamcrops_fresh$SOYB * # 
#   r_physical_areas_ha_perpix$SOYB
# # BARL
# r_gaez_estimate_attn_yield_kg_perpix_BARL <-
#   r_GAEZattnyld_for_spamcrops_fresh$BARL * # 
#   r_physical_areas_ha_perpix$BARL
# 
# r_gaez_estimate_attn_yield_kg_perpix_comb <-
#   c(
#   r_gaez_estimate_attn_yield_kg_perpix_WHEA,
#   r_gaez_estimate_attn_yield_kg_perpix_MAIZ,
#   r_gaez_estimate_attn_yield_kg_perpix_SOYB,
#   r_gaez_estimate_attn_yield_kg_perpix_BARL)
#   
# global(r_gaez_estimate_attn_yield_kg_perpix_comb, "sum", na.rm = T)/1e9 # 
# # WHEA 282
# # MAIZ  414
# # SOYB  89
# # BARL  74
# 
# 
# #------------------------------------------ Now SPAM yields
# r_spam_yields_in2020_comb <-
#   c(
#     r_global_prod_crops_mt$WHEA,
#     r_global_prod_crops_mt$MAIZ,
#     r_global_prod_crops_mt$SOYB,
#     r_global_prod_crops_mt$BARL)
# global(r_spam_yields_in2020_comb, "sum", na.rm = T)/1e6 #
# # WHEA 764
# # MAIZ  1170
# # SOYB  354
# # BARL  154
# 
# 
# #### Maakohtaiset erot - Buruni ym lähellä toisiaan?
# 
# df_r_gaez_estimate_attn_yield_kg_perpix_comb_bycountry <-
#   extract(r_gaez_estimate_attn_yield_kg_perpix_comb/1000,
#           adm_10m,
#           fun = sum, na.rm = T) |> 
#   mutate(ADMIN = adm_10m$ADMIN) # now converted to metric tons
# 
# # same for spam
# df_r_spam_yields_in2020_comb_bycountry <-
#   extract(r_spam_yields_in2020_comb,
#           adm_10m,
#           fun = sum, na.rm = T) |> 
#   mutate(ADMIN = adm_10m$ADMIN)
# 
# 
# # join them using left_join
# df_test <-
#   left_join(df_r_gaez_estimate_attn_yield_kg_perpix_comb_bycountry,
#             df_r_spam_yields_in2020_comb_bycountry,
#             by = "ADMIN") 

```




