# Gaez KCAL

# Relevant area for GAEZ crops

```{r}

# ne gl alueet joilla ei ole cl alueita
# LISÄKSI: SI < 1 pois
gl_areas_with_no_current_cl <- mask(r_fraction_gl_0toNA, r_fraction_cl_0toNA, inverse = T)
plot(gl_areas_with_no_current_cl, main = "gl areas with no current cl")

# SI_5arcmin already correct SI data
gl_areas_with_no_current_cl_with_SIgte1 <-
  mask(gl_areas_with_no_current_cl,
       ifel(SI_5arcmin < 1, NA, SI_5arcmin) ) # SI already updated tata but doulbe check
plot(gl_areas_with_no_current_cl_with_SIgte1, main = "suitable gl areas with no current cl")

gl_areas_with_no_current_cl_with_SIgte1

```

# GAEZ attainable yields

Convert to fresh yields (GAEZ originally as DM yields)


```{r}
# List attainable yield raster (all where pattern is tif)
r_attainable_yield <-
  list.files(here("Data", "Input", "gaez"), 
             pattern = ".tif", full.names = T) |> 
  map(rast) |> 
  rast()

r_attainable_yield # 48 lyr
names(r_attainable_yield) # 48 lyr
r_attainable_yield$ycLr_con |> plot()

# Create df of the GAEZ names under Theme ID 4 'Suitability and Attainable Yield'
gaez_names <-  tibble(
  Crop_acronym = c("alf", "ban", "bck", "brl", "bsg", "cab", "car", "chk", "cit", "coc", 
                   "cof", "con", "cot", "cow", "csv", "flx", "fml", "grd", "grm", "jtr", 
                   "mis", "mze", "nap", "oat", "olp", "olv", "oni", "pea", "phb", "pig", 
                   "pml", "rcd", "rcg", "rcw", "rsd", "rub", "rye", "sfl", "soy", "spo", 
                   "srg", "sub", "suc", "swg", "tea", "tob", "tom", "whe", "wpo", "yam"),
  Crop_name = c("Alfalfa", "Banana", "Buckwheat", "Barley", "Biomass sorghum", "Cabbage", 
                "Carrot", "Chickpea", "Citrus", "Cocoa", "Coffee", "Coconut", "Cotton", 
                "Cowpea", "Cassava", "Flax", "Foxtail millet", "Groundnut", "Gram", "Jatropha", 
                "Miscanthus", "Maize", "Napier grass", "Oat", "Oil palm", "Olive", "Onion", 
                "Dry pea", "Phaseolus bean", "Pigeonpea", "Pearl millet", "Dryland rice", 
                "Reed canary grass", "Wetland rice", "Rapeseed", "Rubber", "Rye", "Sunflower", 
                "Soybean", "Sweet potato", "Sorghum", "Sugar beet", "Sugar cane", "Switchgrass", 
                "Tea", "Tobacco", "Tomato", "Wheat", "White potato", "Yam"))


# Tibble of GAEZ crop names combined with SPAM crops names
# Includes also conversio factor that GAEZ has used to conver fresh yields to DM yields
# conversion factors can be found in table 9-1 in gaez documentation pdf
combined_data_gaez_spam <- tribble(
  ~gaezname, ~gaez_long_name, ~spamname, ~spam_long_name, ~conversion_factor,
  "ban", "Banana", "BANA", "Bananas", 0.25,
  "brl", "Barley", "BARL", "Barley and products", 0.87,
  "chk", "Chickpea", "CHIC", "Pulses, Other and products", 1.00,
  "con", "Coconut", "CNUT", "Coconuts - Incl Copra", 1.00,
  "cow", "Cowpea", "COWP", "Pulses, Other and products", 1.00,
  "csv", "Cassava", "CASS", "Cassava and products", 0.35,
  "grd", "Groundnut", "GROU", "Groundnuts", 0.65,
  "mze", "Maize", "MAIZ", "Maize and products", 0.86,
  "soy", "Soybean", "SOYB", "Soybeans", 0.90,
  "srg", "Sorghum", "SORG", "Sorghum and products", 0.87,
  "whe", "Wheat", "WHEA", "Wheat and products", 0.87,
  "pml", "Pearl millet", "PMIL", "Millet and products", 0.90,
  "phb", "Phaseolus bean", "BEAN", "Beans", 1.00,
  "pig", "Pigeonpea", "PIGE", "Pulses, Other and products", 1.00,
  "spo", "Sweet potato", "SWPO", "Sweet potatoes", 0.25,
  "wpo", "White potato", "POTA", "Potatoes and products", 0.20,
  "yam", "Yam", "YAMS", "Yams", 0.35,
  "pea", "Dry pea", "OPUL", "Pulses, Other and products", 1.00,
  "rcd", "Dryland rice", "DRYLANDRICE", "Rice and products", 0.87,
  "rcw", "Wetland rice", "WETLANDRICE", "Rice and products", 0.87,
  "tom", "Tomato", "VEGETOMATOES", "Vegetables", 1.00,
  "cab", "Cabbage", "VEGECAPPAGES", "Vegetables", 1.00,
  "oni", "Onion", "VEGEONION", "Vegetables", 1.00,
  "cit", "Citrus", "TROF", "Fruits, other", 1.00
) |> 
  mutate(gaezname_full = paste0("ycLr_", gaezname))
# OCER, ORTS, LENT, PLNT missing at least --> should we add oat or rye?

############# Select SPAM crops from the GAEZ raster layers
# Create a vector of GAEZ crop acronyms that match the SPAM crops
gaez_acronyms <- 
  gaez_names %>% 
  filter(Crop_acronym %in% combined_data_gaez_spam$gaezname) %>% 
  pull(Crop_acronym) # 24

# Create a pattern to match the crop acronyms in the raster layer names
pattern_gaez <- paste0("ycLr_", gaez_acronyms, collapse = "|")

# Select the raster layers that match the pattern
r_GAEZattnyld_for_spamcrops <- 
  subset(r_attainable_yield, 
         grep(pattern_gaez, names(r_attainable_yield)))

# Convert values (0 or smaller) to NA
r_GAEZattnyld_for_spamcrops <- 
  terra::ifel(r_GAEZattnyld_for_spamcrops <= 0, NA,
              r_GAEZattnyld_for_spamcrops) # still gaez names

plot(r_GAEZattnyld_for_spamcrops)

# Change names to SPAM names --> order must remain the same
ordered_indices <- 
  match(combined_data_gaez_spam$gaezname_full,
        names(r_GAEZattnyld_for_spamcrops))

r_GAEZattnyld_for_spamcrops <-
  r_GAEZattnyld_for_spamcrops[[ordered_indices]]

names(r_GAEZattnyld_for_spamcrops) <- combined_data_gaez_spam$spamname





# ---------------------------------------------------- convert DM to fresh yields
# r_GAEZattnyld_for_spamcrops_fresh_whea <-
#   subset(r_GAEZattnyld_for_spamcrops, "WHEA") /
#   (combined_data_gaez_spam |>
#      filter(spamname == "WHEA") |>
#      pull(conversion_factor) )

# or with GAEZ names
# r_GAEZattnyld_for_spamcrops_fresh_whea <-
#   subset(r_GAEZattnyld_for_spamcrops, "ycLr_whe") /
#   (combined_data_gaez_spam |>
#      filter(gaezname_full == "ycLr_whe") |>
#      pull(conversion_factor) )


# create function for this
f_DM_to_fresh <-
  function(select_spamname) {
    
    r_GAEZattnyld_for_spamcrops_fresh <-
      
    (subset(r_GAEZattnyld_for_spamcrops, 
            select_spamname)) /
      
      (combined_data_gaez_spam |> 
     filter(spamname == select_spamname) |> 
     pull(conversion_factor) )
    
    return(r_GAEZattnyld_for_spamcrops_fresh)
  }

# run function for each unique spamnames
r_GAEZattnyld_for_spamcrops_fresh <-
  purrr::map(unique(combined_data_gaez_spam$spamname),
             f_DM_to_fresh) |>  # creates a list
  rast() # creates a rast out of list --> 24 lyr
  





# ------------------------------------------------------------------------------- EDIT: all of these could be plotted in facet in supplementary --> only when converted to fresh matter
plot(r_GAEZattnyld_for_spamcrops_fresh)


# -------------------------------------------------------- mask to relevant area
r_GAEZattnyld_for_spamcrops_fresh_relevant_areas <-
  mask(r_GAEZattnyld_for_spamcrops_fresh,
       gl_areas_with_no_current_cl_with_SIgte1)

plot(r_GAEZattnyld_for_spamcrops_fresh_relevant_areas) # fresh


```





# Create new fbs_crop_wide where additional GAEZ crops

```{r}
foodBalanceSheets_semiraw <-
  here("Data", "Input", "faostat", "foodBalanceSheets_E_All_Data_NOFLAG.csv") %>%
  read.csv() %>%
  dplyr::select(-Element.Code,-Unit , -c(Y2010:Y2019), -c(Y2021)) #
names(foodBalanceSheets_semiraw)

# ---------------------------------------------------------- crops
my_itemlist_fao <- c(  # in mapspam webpage order
  "Wheat and products", "Rice and products","Maize and products",
  "Barley and products", "Millet and products", "Sorghum and products",
  "Cereals, Other", "Potatoes and products", "Sweet potatoes",
  "Yams", "Cassava and products", "Roots, Other",  
  "Beans", "Peas", "Pulses, Other and products",
  "Soyabeans","Groundnuts", "Coconuts - Incl Copra",
  "Bananas", "Plantains", "Fruits, other", "Vegetables" # categories for SPAM names
)

my_elementlist_fao <- c(
  "Production","Food supply quantity (kg/capita/yr)",
  "Food supply (kcal/capita/day)", "Protein supply quantity (g/capita/day)"
)


## filter
fbs_crops <- foodBalanceSheets_semiraw  %>% 
  filter(Item %in% my_itemlist_fao,
         Element %in% my_elementlist_fao)



## to wide format. Select only relevant variables
fbs_crops_wide_gaez <- fbs_crops %>% 
  pivot_wider(names_from = Element, values_from = Y2020) %>% # 2020 
  rename(prod_1000mt = "Production",
         supply.KgCapYr = "Food supply quantity (kg/capita/yr)",
         supply.KcalCapD = "Food supply (kcal/capita/day)",
         prot.gCapD = "Protein supply quantity (g/capita/day)") 



## add column where item names corresponds to mapsmap names
## see Definitions and standards - Food Balances (2010-) for the classes
## cannot add information for all crops like this -- add MILL, COWP, PIGE, OPUL, LENT, TROF afterwards
fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  mutate(spamname = case_when(
    Item == "Wheat and products" ~ "WHEA",
    Item == "Rice and products"  ~ "RICE",
    Item == "Barley and products" ~ "BARL",
    Item == "Maize and products" ~ "MAIZ",
    Item == "Millet and products"  ~ "PMIL", 
    #    Item == "Millet and products"  ~ "MILL", # ei toimi koska 2x Millet and products
    Item == "Sorghum and products"  ~ "SORG",
    Item == "Cereals, Other"   ~ "OCER",
    Item == "Potatoes and products" ~ "POTA",
    Item == "Sweet potatoes"  ~ "SWPO",
    Item == "Yams"  ~ "YAMS",
    Item == "Cassava and products"   ~ "CASS",    
    Item == "Roots, Other"  ~ "ORTS",
    Item == "Beans"  ~ "BEAN",
    Item == "Pulses, Other and products"  ~ "CHIC", 
    #    Item == "Pulses, Other and products" ~ "COWP", 
    #    Item == "Pulses, Other and products" ~ "PIGE",     
    #    Item == "Pulses, Other and products"  ~ "OPUL", 
    #    Item == "Pulses, Other and products"  ~ "LENT", 
    Item == "Soyabeans"  ~ "SOYB",
    Item == "Groundnuts"   ~ "GROU",
    Item == "Coconuts - Incl Copra"  ~ "CNUT",
    Item == "Bananas"  ~ "BANA",  
    Item == "Plantains"  ~ "PLNT",
    Item == "Fruits, other"  ~ "TROF",    
    #    Item == "Fruits, other"  ~ "TEMF",    
    Item == "Vegetables"  ~ "VEGE"
  ))


# could be done with function but was faster like this..
additional_rows <- list(
  fbs_crops_wide_gaez %>% filter(Item == "Millet and products") %>% mutate(spamname = "MILL"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "COWP"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "PIGE"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "OPUL"),
  fbs_crops_wide_gaez %>% filter(Item == "Pulses, Other and products") %>% mutate(spamname = "LENT"),
  fbs_crops_wide_gaez %>% filter(Item == "Fruits, other") %>% mutate(spamname = "TEMF"),
  
  # !!
  # These rows added only when combining GAEZ attainable yields 
  fbs_crops_wide_gaez %>% filter(Item == "Rice and products") %>% mutate(spamname = "DRYLANDRICE"),
  fbs_crops_wide_gaez %>% filter(Item == "Rice and products") %>% mutate(spamname = "WETLANDRICE"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGETOMATOES"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGECAPPAGES"),
  fbs_crops_wide_gaez %>% filter(Item == "Vegetables") %>% mutate(spamname = "VEGEONION"))

fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  bind_rows(additional_rows)









# ------------------------- Calculate fraction of protein and kcal in a product (conversion factor) 
## protein and kcal
fbs_crops_wide_gaez <- fbs_crops_wide_gaez %>% 
  mutate(protein_fraction = (prot.gCapD * 365 / 1000) / supply.KgCapYr,
         kcalPerKg = (supply.KcalCapD * 365)/ supply.KgCapYr)  %>% 
  mutate(one_kg_prot_to_kcal =  (supply.KcalCapD / prot.gCapD) * 1000) |> 
  
  ## remove cases where supply = 0 (cases that yield NA values)  
  ## # also converts NaN to NA
  mutate_all(function(x) ifelse(is.nan(x) | is.infinite(x), NA, x)) 




# ---------------------------------------------------------- Fill empty rows
# Create a function that fills empty values of cols protein_fraction, kcalPerKg and one_kg_prot_to_kcal
f_fill_missing_nutritient_values_crops <- function(cropname_mapspam) {
  
  ## pull protein fraction for spesific crop
  protein_fraction_spamcrop <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam, Area == "World") %>%
    pull(protein_fraction)
  
  
  ## pull KcalPerKg for spesific crop
  KcalPerKg_spamcrop <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam, Area == "World") %>%
    pull(kcalPerKg)
  
  ## pull one_kg_prot_to_kcal for spesific crop
  one_kg_prot_to_kcal_spamcrop <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam, Area == "World") %>%
    pull(one_kg_prot_to_kcal)
  
  
  ## fill empty protein fraction and kcalPerKg with global averages
  fbs_crops_wide_gaez_filled <- fbs_crops_wide_gaez %>% 
    filter(spamname == cropname_mapspam) %>%
    ## fill protein_fraction column
    mutate(protein_fraction = 
             ifelse(is.na(protein_fraction) | protein_fraction == 0 , # if protein fraction = NA or 0,
                    protein_fraction_spamcrop, # fill with the global average value
                    protein_fraction)) %>% # otherwise let it just be
    
    ## fill KcalPerKg column
    mutate(kcalPerKg = 
             ifelse(is.na(kcalPerKg) | kcalPerKg == 0 , 
                    KcalPerKg_spamcrop,
                    kcalPerKg)) |> 
    
    ## fill one_kg_prot_to_kcal column
    mutate(one_kg_prot_to_kcal = 
             ifelse(is.na(one_kg_prot_to_kcal) | one_kg_prot_to_kcal == 0 , 
                    one_kg_prot_to_kcal_spamcrop,
                    one_kg_prot_to_kcal))
  
  return(fbs_crops_wide_gaez_filled)
}



test1 <- f_fill_missing_nutritient_values_crops("SOYB") # returns a df

# Lapply this for all the crops --> returns a list
fbs_crops_wide_gaez_filled_list <- 
  lapply(fbs_crops_wide_gaez$spamname |> unique(),
         f_fill_missing_nutritient_values_crops) 




# next we use bind_rows to make this reasonable looking dataframe
fbs_crops_wide_gaez_filled <- bind_rows(fbs_crops_wide_gaez_filled_list)



```



# Convert yields to PROTEIN

same as previously

```{r}
#------------------------------------------------------- crop protein production
  ## create a function to convert crop production to protein

f_crop_protein_production_kg_ha_GAEZ <- function(cropname_mapspam) {
  
  ## get protein fraction for spesific crop species
  faoid_and_prot_frac <-
    fbs_crops_wide_gaez_filled %>%
    filter(spamname == cropname_mapspam) %>%
    dplyr::select(Area.Code, protein_fraction)
  
  ## reclassification matrix
  rcl_mat_countries_to_prot_frac <-
    matrix(c(
      faoid_and_prot_frac$Area.Code,
      faoid_and_prot_frac$protein_fraction),
      ncol = 2)
  
  ## classify. Creates a raster that shows protein fraction for each country
  
    # first define values for others
  fbs_avg_proteinfract_of_a_crop <-
  fbs_crops_wide_gaez_filled %>%
  filter(spamname == cropname_mapspam, Area == "World") %>%
  pull(protein_fraction)
  
  r_fraction_prot  <- 
    cntry_raster %>% #  south sudan?
    classify(rcl_mat_countries_to_prot_frac, 
             others = fbs_avg_proteinfract_of_a_crop)
  
  
  # production raster  ----------- gaez data here!!
  r_production_crop <- 
    subset(r_GAEZattnyld_for_spamcrops_fresh_relevant_areas,
           cropname_mapspam) 
           
  
  
  ## protein production raster
  r_prot <-
    r_production_crop * r_fraction_prot
  names(r_prot) <- paste0(cropname_mapspam, "_prot")
  
  return(r_prot)
}

( gaez_MAIZ_prot <- f_crop_protein_production_kg_ha_GAEZ("MAIZ") )
( gaez_DRYLANDRICEprot <- f_crop_protein_production_kg_ha_GAEZ("DRYLANDRICE") )



# Use map() to apply this function to all the crops
r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha <-
  map(combined_data_gaez_spam$spamname, # väärä nimi
      f_crop_protein_production_kg_ha_GAEZ) |> 
  rast()


# Combine DRYLANDRICE_prot"  "WETLANDRICE_prot" "VEGETOMATOES_prot" "VEGECAPPAGES_prot" "VEGEONION_prot
r_VEGE_prot <-
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGETOMATOES_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGECAPPAGES_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$VEGEONION_prot) |> 
  sum(na.rm=T)
names(r_VEGE_prot) <- "VEGE_prot"


r_RICE_prot <-
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$DRYLANDRICE_prot,
    r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha$WETLANDRICE_prot) |> 
  sum(na.rm=T)
names(r_RICE_prot) <- "RICE_prot"




r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha <- 
  c(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha, r_VEGE_prot, r_RICE_prot) |> 
  
  # remove some layers
  subset(c("DRYLANDRICE_prot", "WETLANDRICE_prot", "VEGETOMATOES_prot", "VEGECAPPAGES_prot", "VEGEONION_prot"), negate = T)




r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha |> plot() # check the areas, e.g. South Sudan


## just a test
# test_gaez_prot_yields <-
#   sum(r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha, na.rm=T)
# plot(test_gaez_prot_yields)
# 
# test_gaez_prot_yields_south_sudan <-
#   crop_and_mask(test_gaez_prot_yields,
#        (adm_10m %>% filter(ADMIN == "South Sudan") %>% dplyr::select(ADMIN) |> 
#   as("Spatial") |>   vect()))
# 
# plot(test_gaez_prot_yields_south_sudan, main ="Potential GAEZ yields for Sudan - no cropmix") 
```

# Perpix

```{r}
# First GAEZ attn yields perpix (for certain SPAM crops)
r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix <-
  r_attainable_Prot_yield_spamcrops_relevant_areas_kg_ha * # esim vehnää 3000 kg/ha (gl alueilla joilla SI > 1)
  cellSize(template_rast_5arcmin, unit = "ha") * # solun koko
  gl_areas_with_no_current_cl_with_SIgte1 # maatalousmaan osuus solussa, esim 20 % (gl alueet joilla SI > 1)

r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix
```



# Protein to kcal using GAEZ perpix yields

```{r}
# First, change names in crop protein production raster
r_prot_allcrops_mt_perpix_namechange_gaez <-
  r_gaez_estimate_attn_total_protein_yield_relevant_areas_kg_perpix/1000 # to MMT
new_names <- sub("_prot$", "", names(r_prot_allcrops_mt_perpix_namechange_gaez)) # # !! 21 lyr only
names(r_prot_allcrops_mt_perpix_namechange_gaez) <- new_names



f_crop_kcal_prod_MM_perpix_gaez <- function(cropname_mapspam) {

  ## get protein fraction for spesific crop species
  faoid_and_kcalPerKg <-
    fbs_crops_wide_gaez_filled %>%
    filter(spamname == cropname_mapspam) %>%
    dplyr::select(Area.Code, 
                  one_kg_prot_to_kcal) # this in amount of kcal within 1kg or protein

  ## reclassification matrix
  rcl_mat_countries_to_one_kg_prot_to_kcal <-
    matrix(c(faoid_and_kcalPerKg$Area.Code, 
             faoid_and_kcalPerKg$one_kg_prot_to_kcal),
           ncol = 2)

  ## classify. Creates a raster with one_kg_prot_to_kcal for spesific crop
  
      # first define values for others
  fbs_avg_one_kg_prot_to_kcal_of_a_crop <-
  fbs_crops_wide_gaez_filled %>%
  filter(spamname == cropname_mapspam, Area == "World") %>%
  pull(one_kg_prot_to_kcal)
  
  
  r_one_kg_prot_to_kcal  <- 
    cntry_raster %>%
    classify(rcl_mat_countries_to_one_kg_prot_to_kcal, 
             others = fbs_avg_one_kg_prot_to_kcal_of_a_crop)

  
    # protein production raster
  r_production_crop_protein_mt_perpix <-
    subset(r_prot_allcrops_mt_perpix_namechange_gaez, # Protein production raster names originally WHEA_prot etc
           cropname_mapspam) 
  
  
  ## kcal production raster
  
  r_kcal_MM_perpix <-
    # divide as production unit = mt and our desired unit = million
    (r_production_crop_protein_mt_perpix * 1000)/1e6 * r_one_kg_prot_to_kcal
  
  names(r_kcal_MM_perpix) <- paste0(cropname_mapspam, "_MMkcal")

  return(r_kcal_MM_perpix)
}

# test
f_crop_kcal_prod_MM_perpix_gaez("WHEA") |> plot()


# lapply for all crops
r_kcal_allcrops_MM_perpix_gaez <- 
  lapply(new_names, ## 21
         f_crop_kcal_prod_MM_perpix_gaez) %>%
  rast()

# We don't take sum here as we need individual layers
# r_kcal_allcrops_sum_MM_perpix_gaez <- sum(r_kcal_allcrops_MM_perpix_gaez, na.rm = T)

```



# SPAM crop mix --> takes few hours to run
--> let's use the crop mix created in script where we evaluated PROTEIN outside cl areas


```{r}
############### get the file
r_crop_share_of_agland <- 
  here("Data", "Intermediate_input",
       "r_SPAMcrops_share_of_agland_adm2_filled_5arcmin.tif") |> 
  rast()
names(r_crop_share_of_agland) # share_WHEA etc
```


# Arvioi nyt crop mix soluille jotka ovat cl alueiden ulkopuolella

Eli maskataan relevantille alueelle (alueelle joka nykyisten cl ulkopuolella mutta joka kuuluu nykyisiin gl alueisiin)

```{r}
# First GAEZ attn yields perpix (for certain SPAM crops)
  # done above



  
# Arvioidaan sato kullekin pix niin että kerrotaan kcal sato sillä osuudella, jota kyseistä kasvia nyt viljellään

r <- r_kcal_allcrops_MM_perpix_gaez

r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix <- sum(
  c(
  r$WHEA_MMkcal * r_crop_share_of_agland$share_WHEA,
  r$RICE_MMkcal * r_crop_share_of_agland$share_RICE,
  r$BARL_MMkcal * r_crop_share_of_agland$share_BARL,
  r$MAIZ_MMkcal * r_crop_share_of_agland$share_MAIZ,
  r$PMIL_MMkcal * r_crop_share_of_agland$share_PMIL,
  #r$MILL_MMkcal * r_crop_share_of_agland$MILL,
  r$SORG_MMkcal * r_crop_share_of_agland$share_SORG,
  # r$OCER_MMkcal * r_crop_share_of_agland$OCER,
  r$CASS_MMkcal * r_crop_share_of_agland$share_CASS,
  r$POTA_MMkcal * r_crop_share_of_agland$share_POTA,
  r$SWPO_MMkcal * r_crop_share_of_agland$share_SWPO,
  r$YAMS_MMkcal * r_crop_share_of_agland$share_YAMS,
  #r$ORTS_MMkcal * r_crop_share_of_agland$ORTS,
  r$BEAN_MMkcal * r_crop_share_of_agland$share_BEAN,
  r$CHIC_MMkcal * r_crop_share_of_agland$share_CHIC,
  r$COWP_MMkcal * r_crop_share_of_agland$share_COWP,
  r$PIGE_MMkcal * r_crop_share_of_agland$share_PIGE,
  r$OPUL_MMkcal * r_crop_share_of_agland$share_OPUL,
  # r$LENT_MMkcal * r_crop_share_of_agland$LENT,
  r$SOYB_MMkcal * r_crop_share_of_agland$share_SOYB,
  r$GROU_MMkcal * r_crop_share_of_agland$share_GROU,
  r$CNUT_MMkcal * r_crop_share_of_agland$share_CNUT,
  r$BANA_MMkcal * r_crop_share_of_agland$share_BANA,
  #r$PLNT_MMkcal * r_crop_share_of_agland$PLNT,
  r$TROF_MMkcal * r_crop_share_of_agland$share_TROF,
  # r$TEMF_MMkcal * r_crop_share_of_agland$TEMF,
  r$VEGE_MMkcal * r_crop_share_of_agland$share_VEGE),
  na.rm = TRUE)

r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix
names(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix) <-
  "gaez_kcal_outside_cl"

### global sums
global(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix*1e6, "sum", na.rm = T)/1e12 # 2425 trillion kcal (about 23% of existing crop kcal prod)





# Convert back to kg per ha (only 1 layer now)
r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha <-
  r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix /
  (cellSize(template_rast_5arcmin, unit = "ha") * gl_areas_with_no_current_cl_with_SIgte1)

names(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha) <-
  "gaez_kcal_outside_cl"

plot(ifel(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha > 0 ,1,NA))


# save perpix raster
writeRaster(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix,
            filename= here("Data", "Intermediate_input",
                           "r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix.tif"),
            overwrite = T)
# per ha
writeRaster(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha,
            filename= here("Data", "Intermediate_input",
                           "r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha.tif"),
            overwrite = T)



####### get the file 
# use rast() function
r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix <-
  here("Data", "Intermediate_input",
       "r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix.tif") |> 
  rast()

r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha <-
  here("Data", "Intermediate_input",
       "r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha.tif") |> 
  rast()


## Again crop and mask to south sudan
test_gaez_kcal_in_south_sudan <-
  crop_and_mask(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha,
                (adm_10m %>% filter(ADMIN == "South Sudan") %>% dplyr::select(ADMIN) |>
  as("Spatial") |>   vect()))
plot(test_gaez_kcal_in_south_sudan) # these outside current croplands! and there is potential
```

# GAEZ-based crop kcal yields outside current cl production areas but existing yields on existing production areas

```{r}
# crops
  # perpix
r_crop_kcal_existing_plus_new_areas_MM_perpix <-
  classify(r_kcal_allcrops_sum_MM_perpix, cbind(NA, 0)) + # from another script
  classify(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_perpix, cbind(NA, 0))

r_crop_kcal_existing_plus_new_areas_MM_perpix <-
  classify(r_crop_kcal_existing_plus_new_areas_MM_perpix, cbind(0, NA))

names(r_crop_kcal_existing_plus_new_areas_MM_perpix) <-
  "crop_kcal_MM_perpix"

  
  # per ha
r_crop_kcal_existing_plus_new_areas_MM_ha <-
  classify(r_kcal_allcrops_sum_MM_ha, cbind(NA,0)) + # another script
  classify(r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha, cbind(NA,0))

r_crop_kcal_existing_plus_new_areas_MM_ha <-
  classify(r_crop_kcal_existing_plus_new_areas_MM_ha, cbind(0,NA))





# SAVE
writeRaster(r_crop_kcal_existing_plus_new_areas_MM_perpix,
            filename= here("Data", "Intermediate_input",
                           "r_crop_kcal_existing_plus_new_areas_MM_perpix.tif"),
            overwrite = T)


writeRaster(r_crop_kcal_existing_plus_new_areas_MM_ha,
            filename= here("Data", "Intermediate_input",
                           "r_crop_kcal_existing_plus_new_areas_MM_ha.tif"),
            overwrite = T)



############## get the file
r_crop_kcal_existing_plus_new_areas_MM_perpix <-
  here("Data", "Intermediate_input",
       "r_crop_kcal_existing_plus_new_areas_MM_perpix.tif") |> 
  rast()


r_crop_kcal_existing_plus_new_areas_MM_ha <-
  here("Data", "Intermediate_input",
       "r_crop_kcal_existing_plus_new_areas_MM_ha.tif") |> 
  rast()



```







# Plot

```{r}
pal_protein_lajolla <- scico(n = 6, palette = "lajolla", begin = 0.15, end = 1, direction = -1) # previously ended 0.85



r_separate_existing_and_GAEZ_derived_kcal_yields <-
  c(r_kcal_allcrops_sum_MM_ha, 
    r_gaez_estimate_attn_total_kcal_yield_relevant_areas_MM_ha)

names(r_separate_existing_and_GAEZ_derived_kcal_yields) <- c("Existing", "GAEZ-derived")
plot(r_separate_existing_and_GAEZ_derived_kcal_yields)


(plt_crop_kcal_gaez <- 
    create_index_map(r_index = r_crop_kcal_existing_plus_new_areas_MM_ha, 
                     tocrs = "ESRI:54030",
                     index_main_title = "kcal: Existing + new areas",
                     index_label = "[million kcal/ha]",
                     colorpal = pal_protein_lajolla,
                     breakvals = c(0, 0.2, 0.5, 1, 10, 50, Inf),
                     breaknames = c("<0.2","0.2-0.5",
                                  "0.5-1","1-10", "10-50", ">50"))) 



(plt_crop_kcal_GAEZ_derived_and_existing <- 
  create_index_map(r_index = r_separate_existing_and_GAEZ_derived_kcal_yields,
                   tocrs = "ESRI:54030",
                   index_main_title = "Crop kcal yield existing and GAEZ derived areas", #  
                   index_label = "[MMkcal/ha]",
                   colorpal = pal_protein_lajolla,
                     breakvals = c(0, 0.2, 0.5, 1, 10, 50, Inf),
                     breaknames = c("<0.2","0.2-0.5",
                                  "0.5-1","1-10", "10-50", ">50"))) 



```


