
# Reality check

Tutkitaan missä maankäytön muutokset ovat ongelmallisia

Ensin 1km tarkkuudella, sitten 10kn

!!!!!!!!!!!!!!!! tehdään tämä combined mask avulla

# Load data

```{r}
# Schneider areas with NO potentially available cropland
# ie pac = 0
pac_no_potential_30as <-
  here("Data", "Input", "from_harddrive",
       "pac_2010-2039_rcp8p5_rainfed_30arcsec_sub_NO_POTENTIAL.tif") |> 
  rast()

# Jung et al top BD areas only -- this should be with 5arcmin resolution!!
Jung_top_BD_areas_5arcmin <-
  here("Data", "Input", "from_harddrive",
       "Jung_top_BD_areas_5arcmin.tif") |> 
  rast()


# Primary vegetation from Florian
prim_veg_0toNA_30arcsec <-
  here("Data", "Input", "other", "primary_vege_from_Florian",
       "primary_vegetation_0toNA_30as.tif") |> 
  rast()


## Primary vegetation with min use
prim_veg_minuse_0toNA_30arcsec <- 
  here("Data", "Input", "other", "primary_vege_from_Florian",
       "primary_vegetation_minimaluse_0toNA_30as.tif") |> 
  rast()


#### plot
plot(pac_no_potential_30as, main = "NO pac")
plot(Jung_top_BD_areas_5arcmin, main = "Jung top BD areas")
plot(prim_veg_0toNA_30arcsec, main = "Primary vegetation")
  plot(prim_veg_minuse_0toNA_30arcsec, main = "Primary vegetation with min use")

# use par arguments to plot all three panels in the same plot
par(mfrow = c(2, 2))
plot(pac_no_potential_30as, main = "NO potentially available cl")
plot(Jung_top_BD_areas_5arcmin, main = "Jung top BD areas")
plot(prim_veg_0toNA_30arcsec, main = "Primary vegetation")
plot(prim_veg_minuse_0toNA_30arcsec, main = "Primary vegetation with min use")



```

# Kohdenna data grazing land alueille

--> Onko välttämätöntä? --> EI!
--> voitaisiinko vain poistaa priority alueet country rasterista 5arcmin hilassa?

```{r}
# Hae 1km mcd data
# gl_mask_30arcsec <- here("Data", "Input", "from_harddrive",
#                           "mcd_igbp6to10or16_2001_2020_na.rm_FALSE_30arcsec.tif") |> 
#   rast()
# gl_mask_30arcsec


# Combine 30 as data
r_areas_where_conversion_not_realistic_30as <-
  c(pac_no_potential_30as,
    prim_veg_0toNA_30arcsec,
    prim_veg_minuse_0toNA_30arcsec)



# mask  30arcsec products to grazing lands 2min
# r_areas_where_conversion_not_realistic_30as_glmask <-
#   mask(r_areas_where_conversion_not_realistic_30as,
#        gl_mask_30arcsec)

# mask Jung top BD areas to 5arcmin
#Jung_top_BD_areas_5arcmin_gl <- mask(Jung_top_BD_areas_5arcmin, r_fraction_gl_0toNA)


plot(r_areas_where_conversion_not_realistic_30as)
# plot(r_areas_where_conversion_not_realistic_30as_glmask)
# plot(Jung_top_BD_areas_5arcmin_gl)

```





# Overlay / mask

From full 30arcsec country raster, remove areas with

- top BD areas 5 arcmin resolution!!
- no pac
- primary vegetation
- primary vegetation with min disturbance

```{r}
# create "full" raster
cntry_raster_30arcsec <- 
  rasterize(vect(adm10_simple_faoadded),
            template_rast_30arcsec, field = "FAO_ID") 

# Yhdistä kaikki maskit sum()-funktiolla
r_areas_where_conversion_not_realistic_combined_30as <-
  sum(r_areas_where_conversion_not_realistic_30as,
    #r_areas_where_conversion_not_realistic_30as_glmask, 
      na.rm =T)


# convert values to 1, otherwise NA
# combined_mask_1_or_NA <- 
#   ifel(combined_mask_30arcsec > 0, 1, NA)
# plot(combined_mask_1_or_NA, main = "Combined mask 1 or NA")
plot(r_areas_where_conversion_not_realistic_combined_30as, 
     main = "Combined mask 30arcsec - unrealistic areas for a shift")


### to 5arcmin
# ! na.rm = TRUE 
# ! so if any pixel in 1km cell has value NA, it gets value NA in 5arcmin raster

#this is because now only when all pixels in 1km cells have value 1 (means no go areas), it gets value 1 in 5arcmin raster
# ! if any pixel in 1km cell has value NA, it gets value NA in 5arcmin raster
r_areas_where_conversion_not_realistic_combined_5arcmin <- 
  aggregate(r_areas_where_conversion_not_realistic_combined_30as,
            fact = 10, na.rm = T) 

# add Jung et al top BD areas
r_areas_where_conversion_not_realistic_combined_5arcmin <- 
  sum(c(r_areas_where_conversion_not_realistic_combined_5arcmin,
        Jung_top_BD_areas_5arcmin),
        #Jung_top_BD_areas_5arcmin_gl),
      na.rm = T)

# Convert values > 0 to 1 and otherwise NA
r_areas_where_conversion_not_realistic_combined_5arcmin <- 
  ifel(r_areas_where_conversion_not_realistic_combined_5arcmin > 0, 
       1, 
       NA)

# mask ti 5arcmin to agricultural land grazing lands
r_areas_where_conversion_not_realistic_combined_5arcmin <- 
  mask(r_areas_where_conversion_not_realistic_combined_5arcmin,
       r_fraction_gl_cl_total_0toNA,
       filename = here("Data", "Intermediate_input",
                       "areas_where_conversion_not_realistic_glclareas_5arcmin.tif"),
       overwrite =T)

plot(r_areas_where_conversion_not_realistic_combined_5arcmin,
     main = "Unrealistic areas for shift - strict estimate with na.rm=T")

# this is maybe too conservative?




# r_areas_where_conversion_is_possible_30as <- 
#   mask(cntry_raster_30arcsec, # full country raster is masked outside the regions
#        combined_mask_30arcsec, # where land usage is not realistic
#        inverse = TRUE)
# 
# r_areas_where_conversion_is_possible_30as_1_or_NA <- 
#   ifel(r_areas_where_conversion_is_possible_30as > 0, 1, NA)
# 
# plot(r_areas_where_conversion_is_possible_30as, main = "Reality check - where gl to cl is possible")
# plot(r_areas_where_conversion_is_possible_30as_1_or_NA, main = "Reality check 1 or NA - where gl to cl is possible")
# 

# ####### Aggregate to 5arcmin resolution
# r_reality_check_5arcmin <- 
#   aggregate(r_reality_check_1_or_NA, fact = 10, na.rm = T) |> 
#   mask(r_fraction_gl_cl_total_0toNA,
#        filename = here("Data", "Intermediate_input",
#                        "r_reality_check_5arcmin.tif"),
#        overwrite =T)




################################# get the data
r_areas_where_conversion_not_realistic_combined_5arcmin <-
  rast(here("Data", "Intermediate_input",
              "areas_where_conversion_not_realistic_glclareas_5arcmin.tif"))



plot(r_areas_where_conversion_not_realistic_combined_5arcmin,
     main = "Unrealistic areas for gl to cl - strict estimate with na.rm=T")


```


# Better maps of restricted areas

```{r}
r_areas_where_conversion_not_realistic_combined_5arcmin_NAto0 <- 
  ifel(is.na(r_areas_where_conversion_not_realistic_combined_5arcmin), 
       0,
       r_areas_where_conversion_not_realistic_combined_5arcmin) |> 
  mask(r_fraction_gl_cl_total_0toNA)

# this map to be shown?
(plt_combined_mask_5arcmin <- 
    create_index_map(r_index = r_areas_where_conversion_not_realistic_combined_glmask_5arcmin_NAto0, 
                     tocrs = "ESRI:54030",
                     index_main_title = "Conservation priority regions",
                     index_label = "",
                     colorpal = c("#D3D3D3", "#1F77B4"), 
                     breakvals =  c(-Inf, 0.1, Inf ), 
                     breaknames = c("Not a priority", "Priority area")))




# areas where conversion is possible
# r_reality_check_5arcmin_NAto0 <- 
#   ifel(is.na(r_reality_check_5arcmin), 0, r_reality_check_5arcmin) |> 
#   mask(r_fraction_gl_cl_total_0toNA)
# 
# (plot_reality_check_5arcmin <- 
#     create_index_map(r_index = r_reality_check_5arcmin_NAto0, 
#                      tocrs = "ESRI:54030",
#                      index_main_title = "Areas where conversion is possible",
#                      index_label = "",
#                      colorpal = c("#D3D3D3", "#1F77B4"), 
#                      breakvals =  c(-Inf, 0.1, Inf ), 
#                      breaknames = c("Not possible", "Possible")))

```

#################### OLD

# Test with the final map

```{r}
# 25% to 1000% here
r_protein_change_currenTO_ALLtoCL_recoded <- 
  here("Data", "Intermediate_input",
       "r_protein_change_currenTO_ALLtoCL_recoded_m25to1000.tif") |> # ! Recoded
  rast()

# priority conservation areas
r_areas_where_conversion_not_realistic_combined_glmask_5arcmin <-
  rast(here("Data", "Intermediate_input",
              "areas_where_conversion_not_realistic_glmask_5arcmin.tif"))


# mask outside priority areas
r_protein_change_currenTO_ALLtoCL_recoded_reality_check <-
  mask(r_protein_change_currenTO_ALLtoCL_recoded,
       r_areas_where_conversion_not_realistic_combined_glmask_5arcmin,
       inverse = T)



# Plot
breaknames_updated <- c("Grazing remains (29%)", 
                        "Cropland remains (14%)", 
                        "Mosaic remains (17%)", 
                        "Beneficial shifts (28%)", 
                        "Adverse shifts (12%)") #Cropland to Grazing? we dont know for sure




pal_categories <- 
  c("#A6CEE3", "#C0A2C7", "#1F78B4", "#B2DF8A", "#FFA500")#

(plt_r_protein_change_currenTO_ALLtoCL_recoded_realitycheck <-
create_index_map(r_index = r_protein_change_currenTO_ALLtoCL_recoded_reality_check,
                 tocrs = "ESRI:54030",
                 index_main_title = "Comparing protein outputs: existing land use vs 'all suitable to cropland'",
                 index_label = "Direction of change: if protein shift [-25, 1000]%, unchanged land use",
                 colorpal = pal_categories,
                
                 breakvals = c(-Inf, -9000, -7000, -500, 10, Inf),
                 breaknames = breaknames_updated))





```

# How much the protein yield would increase if all beneficial shifts would be converted to cropland?



```{r}
# Separate "beneficial shift" areas
r_perhaps_realistic_protein_increase_areas <-
  ifel(r_protein_change_currenTO_ALLtoCL_recoded_outside_conservation_areas == 1, # 1 means converting is beneficial
       r_protein_change_currenTO_ALLtoCL_recoded_outside_conservation_areas,
       NA) 

plot(r_perhaps_realistic_protein_increase_areas, col = "#B2DF8A")

# use this as a mask to show:
# how much is the combined protein yield in these areas of beneficial shifts?
r_optimised_combined_protein_yield_kg_perpix_where_beneficial_shifts <- 
  mask(r_optimised_combined_protein_yield_kg_perpix_positive$prot_kgpix_SI1, # ALL to CL where SI > 1
       r_perhaps_realistic_protein_increase_areas)


# Use existing protein yields elsewhere
r_protein_combined_kg_perpix_where_no_changes_in_land_use <-
  mask(r_protein_combined_kg_perpix, 
       r_perhaps_realistic_protein_increase_areas, 
       inverse = T) |> 
  mask(r_fraction_gl_cl_total_0toNA)

# Plot both
plot(r_optimised_combined_protein_yield_kg_perpix_where_beneficial_shifts, col = "#B2DF8A")
plot(r_protein_combined_kg_perpix_where_no_changes_in_land_use, col = "#1F78B4")

# combine these
r_perhaps_realistic_protein_production_kg_perpix <-
  classify(r_protein_combined_kg_perpix_where_no_changes_in_land_use, cbind(NA, 0)) +
  classify(r_optimised_combined_protein_yield_kg_perpix_where_beneficial_shifts, cbind(NA, 0)) |> 
  mask(r_fraction_gl_cl_total_0toNA)

# Compare with existing production
global(r_perhaps_realistic_protein_production_kg_perpix, "sum", na.rm=T)/1e9
global(r_protein_combined_kg_perpix, "sum", na.rm=T)/1e9

(global(r_perhaps_realistic_protein_production_kg_perpix, "sum", na.rm=T) - global(r_protein_combined_kg_perpix, "sum", na.rm=T)) /
  global(r_protein_combined_kg_perpix, "sum", na.rm=T) * 100

# 21.63% more protein could be produced if all beneficial shifts would be converted to cropland!!


```





# Alternative 2 -- much more conservative with na.rm = T

```{r}
# combined_mask_5arcmin_narm_TRUE <-
#   aggregate(combined_mask_1_or_NA, fact = 10, na.rm = T) |>
#   mask(r_fraction_gl_cl_total_0toNA) 
# 
# plot(combined_mask_5arcmin_narm_TRUE, main = "Combined mask 5arcmin - na.rm = TRUE")
# 
# combined_mask_5arcmin_narm_TRUE_NAto0 <-
#   ifel(is.na(combined_mask_5arcmin_narm_TRUE), 0, combined_mask_5arcmin_narm_TRUE) |> 
#   mask(r_fraction_gl_cl_total_0toNA)
# 
# (plt_combined_mask_5arcmin_narm_TRUE <- 
#     create_index_map(r_index = combined_mask_5arcmin_narm_TRUE_NAto0, 
#                      tocrs = "ESRI:54030",
#                      index_main_title = "Conservation priority regions - conservative estimate",
#                      index_label = "",
#                      colorpal = c("#D3D3D3", "#1F77B4"), 
#                      breakvals =  c(-Inf, 0.1, Inf ), 
#                      breaknames = c("Not a priority", "Priority area")))
# 
# 
# 
# 
# # mask final results outside these areas
# breaknames_updated <- c("Grazing remains (%)", 
#                         "Cropland remains (%)", 
#                         "Mosaic remains (%)", 
#                         "Beneficial shifts (%)", 
#                         "Adverse shifts (%)") #Cropland to Grazing
# 
# 
# 
# (plt_r_protein_change_currenTO_ALLtoCL_recoded_realitycheck_narm_TRUE <-
# create_index_map(r_index = mask(r_protein_change_currenTO_ALLtoCL_recoded_reality_check,
#                                 combined_mask_5arcmin_narm_TRUE, inverse = T),
#                  tocrs = "ESRI:54030",
#                  index_main_title = "MORE strict reality check",
#                  index_label = "Direction of change. If protein production changes < ±50%, land use remains same)",
#                  colorpal = pal_categories,
#                 
#                  breakvals = c(-Inf, -9000, -7000, -500, 10, Inf),
#                  breaknames = breaknames_updated))
```

